---
title: "GAMs for 10km data"
author: "Yvonne Barkley"
date: "9/21/2020"
output: html_document
---


Load libraries
```{r,eval=FALSE,echo=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(geoR)
library(tidymv)
```

```{r}
gridsize = 10
loctype = 'allsw'
loctype2 = 'all'
# loctype = 'trackline'
# loctype2 = 'trk'

PmScaled <- readRDS(here::here( paste0('output/envData/complete env data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda') ))
PmRaw <- readRDS(here::here( paste0('output/envData/complete env data/', 'CompletePm_', gridsize, 'km_', loctype2, '_raw.rda') ))

CompletePreds <- dplyr::select(PmRaw, sst:d2smt)
# PmScaled <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_scaledNEW.rda')))


###testing things with DUMMY data
# PmScaledOLD <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_scaledOLD.rda')))

#add column for log effort as offset
PmScaled$log.effort	= log(PmScaled$EffArea)
PmScaled$pa2 = ifelse(PmScaled$pa >1, PmScaled$pa2 == 1, PmScaled$pa)
PmRaw$log.effort	= log(PmRaw$EffArea)
PmRaw$d2smt = PmRaw$d2smt/1000
PmRaw$pa2 = ifelse(PmRaw$pa >1, PmRaw$pa2 == 1, PmRaw$pa)


saveRDS(PmScaled, here::here( paste0('output/envData/complete env data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled2.rda') ))

nrow(filter(PmScaled, pa > 0)) #how many presences? was 118, then 147 for 25km since >1 whale in a cell, now should be 159

# pairs(PmScaled[,15:26])
```

Check correlation of covariates
```{r, echo=FALSE, eval=FALSE}
corrplot.mixed(cor(PmScaled[,18:29]), upper="number", lower="circle")

```

Plot the data, see what's happening
```{r,eval=FALSE}
# par(mfrow = c(1,1))
# with(CompleteTotal, plot(pa~bath_m))
require(Cairo)
require(ggplot2)
source('./code/multiplot.r') 

 
loopVec <- 18:29  #columns from PmScaled to plot
 
# par(mfrow = c(1, 3), mar=c(3,3,1,1), oma=c(0,0,3,1))
 
for (j in loopVec){
   #use original values for plots with PmRaw
   datPlot <- PmScaled[,c(1,j)]  #get dataframe of pa and variable j
   datPlot$bin <- 0
   datPlot$bin <- ifelse(datPlot$pa >0 , 1, 0) #put 1s in color col for plotting later
   datPlot <- select(datPlot, bin, 2, pa)
    #normal plot
   # plot(datPlot[,2], datPlot[,1], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'whales per cell')
   

   ptplt <-  ggplot(datPlot, aes(datPlot[,2],datPlot[,1], colour=as.factor(datPlot[,1]))) +
       geom_point() +   
     theme(legend.position ="none") +
       xlab(paste(colnames(datPlot)[2], 'scaled' )) +
       ylab('whales per cell')+
       # scale_colour_discrete(name='',labels = c('Whales Absent', 'Whales Present'))+
theme_bw() +
  guides(colour=FALSE)
   
   #histogram
   # hist(datPlot[,2], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'frequency', main='' )
   
   histplt <- ggplot(datPlot, aes(datPlot[,2], fill=as.factor(datPlot[,1]))) +
     geom_histogram(colour='black', binwidth = 0.5) +
    xlab(paste(colnames(datPlot)[2],'scaled' )) +
    ylab('Frequency') +
     scale_fill_discrete(name = '', labels = c('Whales Absent', 'Whales Present'))+
     theme_bw() +
     theme(legend.position ="top",
           legend.text=element_text(size=16)) 

   
   #do KS test and plot
   datPlot1 <- filter(datPlot, pa >0)
   datPlot0 <- filter(datPlot, pa ==0)

   datKS <- ks.test(datPlot1[,2], datPlot0[, 2])
     cdf0 <- ecdf(datPlot0[,2]); cdf1 <- ecdf(datPlot1[,2]) #make cdfs for the absences and whales for that variable
   #find min and max stats to draw the line between the greatest distance between the distributions
   minMax <- seq(min(datPlot0[,2], datPlot1[,2]), max(datPlot0[,2], datPlot1[,2]), length.out=length(datPlot0[,2])) 
   x0 <- minMax[which( abs(cdf0(minMax) - cdf1(minMax)) == max(abs(cdf0(minMax) - cdf1(minMax))) )] 
   y0 <- cdf0(x0) #sometimes there are 2 points with the same minmax, take the first one 
   y1 <- cdf1(x0)
   
  ksplt <- ggplot(datPlot, aes(x = datPlot[,2], group = datPlot[,1], color = as.factor(datPlot[,1]))) + #
  stat_ecdf(size=1) +
    theme_bw() +
    theme(legend.position ="none", plot.title = element_text(hjust= 0.5)) +
    xlab(paste(colnames(datPlot)[2],'scaled') ) +
    ylab("ECDF") +
    #geom_line(size=1) +
    geom_segment(aes(x = x0[1], y = y0[1], xend = x0[1], yend = y1[1]),
        linetype = "dashed", color = "red") +
    geom_point(aes(x = x0[1] , y= y0[1]), color="red", size=2) +
    geom_point(aes(x = x0[1] , y= y1[1]), color="red", size=2) +
    ggtitle("K-S Test") +
    # theme(legend.title=element_blank()) +
     scale_colour_discrete(name = 'Distribution Type', labels = c('Whales Absent', 'Whales Present'))+
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.1, hjust=.2,label=paste('D=',round(datKS[1]$statistic,2))) +
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.05, hjust=.2,label=paste('p=',round(datKS$p.value,3)))

Cairo(here::here('figures', paste0('DataPlotsScaled_', gridsize, 'km_', loctype,'_', colnames(datPlot)[2], '.png')), width=14,height=6,
    units='in', dpi=300, bg='white')
  multiplot(ptplt, histplt, ksplt, cols=3)
 dev.off()
  
   # ggsave(paste('DataPlots_',colnames(datPlot)[2],'.png'))
  
  }

```
##More plots
```{r}

 histT600 <- ggplot(PmRaw, aes(temp600, fill=as.factor(pa))) +
     geom_histogram(colour='black', binwidth = 0.5) +
    xlab(paste(colnames(datPlot)[2],'scaled' )) +
    ylab('Frequency') +
     scale_fill_discrete(name = 'Number of Whale Groups') + # labels = c('Whales Absent', 'Whales Present'))+
     theme_bw() +
     theme(legend.position ="top",
           legend.text=element_text(size=16)) 


```


##10km - Binomial

No knot contrain
```{r, eval=FALSE}
# colnames(PmScaled)

#treat as count data using Tweedie?
tw10 <- gam(pa ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + offset(log.effort), data = PmScaled, family = tw)
summary(tw10)

par(mar=c(4,4,3,3), mfrow=c(3,3))
plot(tw10, pages = 3, residuals = TRUE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'deepskyblue', ylim = c(-20, 10))

gam.check(tw10)

# tw25_p <- predict_gam(tw25)

#treat as binary distribution with binomial 
bin10 <- gam(pa2 ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + offset(log.effort), data = PmScaled, family = binomial, link='logit')
summary(bin10)

#treat as binary distribution with binomial 
nb10 <- gam(pa2 ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + offset(log.effort), data = PmScaled, family = nb, link='log')
summary(nb10)

#treat as binary distribution with binomial 
nb10raw <- gam(pa2 ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + offset(log.effort), data = PmScaled, family = nb, link='log')
summary(nb10)
gam.check(nb10)
par(mar=c(4,4,3,3), mfrow=c(3,3))
plot(nb10raw, pages = 3, residuals = TRUE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'orange', ylim = c(-20, 10))
```

Knots contrained to k=3
Automatic term selection is uses an additional penalty term when determining the smoothness of the function.
  We excluded all non-significant variables (alpha=0.05) and refit the models until all variables were significant.
REML is restricted maximum likelihood used to optimize the parameter estimates.
```{r}
#Automatic term selection
nb10k3sel <- gam(pa2 ~ s(bath_m, k=3) + s(sst, k=3) + s(chla, k=3) + s(temp600) + s(slope, k=3) + s(aspect, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(d2smt, k=3) +s(dist, k=3) + offset(log.effort), data = PmScaled, family = nb, link = 'log', select = TRUE, method = "REML")
summary(nb10k3sel)
par(mar=c(4,4,3,3), mfrow=c(2,2))
gam.check(nb10k3sel)

#testing without log.effort offset
# tw10k3sel2 <- gam(pa ~ s(bath_m, k=3) + s(sst, k=3) + s(chla, k=3) + s(temp600) + s(slope, k=3) + s(aspect, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(d2smt, k=3) +s(dist, k=3), data = PmScaled, family = tw, link = 'log', select = TRUE, method = "REML")

#Used raw data for plotting purposes like a kook
nb10k3selraw <- gam(pa2 ~ s(bath_m, k=3) + s(sst, k=3) + s(chla, k=3) + s(temp600) + s(slope, k=3) + s(aspect, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(d2smt, k=3) +s(dist, k=3) + offset(log.effort), data = PmRaw, family = nb, link = 'log', select = TRUE, method = "REML")
summary(nb10k3selraw) #just checking that these are the same so I can plot raw values

par(mar=c(4,4,3,3), mfrow=c(3,3))
plot(nb10k3selraw, pages = 3, residuals = TRUE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'deepskyblue', ylim = c(-20, 10))
```

Taking previous model with k=3, auto term selection and changing some variables from smoothers to linear parameters
```{r}

nb10k3selB <- gam(pa2 ~  sst + chla + s(temp600) + ssh + sshsd + d2smt + offset(log.effort), data = PmScaled, family = nb, link = 'log', select = TRUE, method = "REML")
nb10k3selBraw <- gam(pa2 ~  sst + chla + s(temp600) + ssh + sshsd + d2smt + offset(log.effort), data = PmRaw, family = nb, link = 'log', select = TRUE, method = "REML")

summary(nb10k3selB)
gam.check(nb10k3selB)
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(nb10k3selBraw, pages = 1, residuals = TRUE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'darkseagreen2', all.terms = TRUE)

#eke not significant, remove from model
nb10k3selC <- gam(pa ~ bath_m + sst + chla + s(temp600) + s(wavepow, k=3) + s(d2smt, k=3) + dist + offset(log.effort), data = PmScaled, family = nb, link = 'log', select = TRUE, method = "REML")
nb10k3selCraw <- gam(pa ~ bath_m + sst + chla + s(temp600) + s(wavepow, k=3) + s(d2smt, k=3) + dist + offset(log.effort), data = PmRaw, family = nb, link = 'log', select = TRUE, method = "REML")
summary(nb10k3selC)
gam.check(nb10k3selC)
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(nb10k3selCraw, pages = 1, residuals = TRUE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'salmon1', all.terms = TRUE)

#increase knots for wave power
nb10k3selD <- gam(pa ~ bath_m + sst + chla + s(temp600) + s(wavepow, k=4) + s(d2smt, k=3) + dist + offset(log.effort), data = PmScaled, family = nb, link = 'log', select = TRUE, method = "REML")
summary(nb10k3selD)
gam.check(nb10k3selD)
nb10k3selDraw <- gam(pa ~ bath_m + sst + chla + s(temp600) + s(wavepow, k=4) + s(d2smt, k=3) + dist + offset(log.effort), data = PmRaw, family = nb, link = 'log', select = TRUE, method = "REML")
plot(nb10k3selDraw, pages = 1, residuals = TRUE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'royalblue1', all.terms = TRUE)

#plot with all signif variables and add 2D smoother
nb10k3selE <- gam(pa ~ bath_m + sst + chla + s(temp600) + s(wavepow, k=3) + s(d2smt, k=3) + dist + offset(log.effort) + s(Longitude, Latitude), 
        data = PmScaled, family = nb, link = 'log', select = TRUE, method = "REML")

nb10k3selEraw <- gam(pa ~ bath_m + sst + chla + s(temp600) + s(wavepow, k=3) + s(d2smt, k=3) + dist + offset(log.effort) + s(Longitude, Latitude),    
        data = PmRaw, family = nb, link = 'log', select = TRUE, method = "REML")
summary(nb10k3selE)
gam.check(nb10k3selE)

par(mar=c(4,4,3,3), mfrow=c(3,3))
plot(nb10k3selEraw, residuals = F, pch = 20, cex = 0.25,
 shade = TRUE, shade.col = 'salmon1', all.terms=TRUE)

# hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)

points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20)
points()


```

What happens when a 2D lat/lon smoother is fit only?
```{r}
# require(SDraw) #doesn't work
nb10kLL <- gam(pa2 ~ s(Longitude, Latitude), data = PmScaled, family = nb, link='log', select = TRUE, method = "REML")
summary(nb10kLL)
gam.check(nb10kLL)

saveRDS(hawaiiMap, here::here(paste0('data/hawaiiMap..rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)

whales <- filter(PmScaled, pa >0) #get points for whales to plot

plot(nb10kLL, select = 1, scheme = 2, lwd = 2)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex=0.5, col = 'black')
points(whales$Longitude, whales$Latitude, pch = 1, col='slateblue', lwd=2, cex=0.75)



us<-readOGR(here::here( paste0("data/tl_2016_us_state.shp")))
hawaii<-subset(us,NAME=="Hawaii")
hawaiiMap<-fortify(hawaii)
```


###Variogram 
*How similar are the residuals relative to their distance from each other in space?
  This tests for autocorrelation. If nearby residuals are similar due to unmeasured env effects, for example, the variogram will show a pattern where there are lower values at lower distances and then an increase at farther distances. 
```{r}
#full model WITHOUT 2D smoother
nb10k3selCvar <- variog(coords = PmScaled[,c('Latitude','Longitude')],	data= residuals(nb10k3selC, type = 'pearson'))
plot(nb10k3selCvar, main='Variogram: Best Full Model, No 2D smoother')

#full model WITH 2D smoother
nb10k3selEvar <- variog(coords = PmScaled[,c('Latitude','Longitude')],	data= residuals(nb10k3selE, type = 'pearson'))
plot(nb10k3selEvar, main='Variogram: Best Full Model, WITH 2D smoother')

#full model WITH ONLY 2D smoother
nb10kLLvar <- variog(coords = PmScaled[,c('Latitude','Longitude')],	data= residuals(nb10kLL, type = 'pearson'))
plot(nb10kLLvar, main='Variogram: Best Full Model, WITH ONLY 2D smoother')


```


###Moran's I
```{r}
coords = cbind(PmScaled$Longitude, PmScaled$Latitude)
w = fields:::rdist(coords)
library(ape)
Moran.I(x = PmScaled$pa, w = w)
```
