---
title: "GAMS - Combined Models - 10 km"
author: "Yvonne Barkley"
date: "9/28/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = '', fig.width =8, fig.height = 6, message=FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


Load libraries
```{r, echo=T, eval=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(geoR)
library(tidymv)
library(here)
```


# Research question:
## What environmental variables characterize sperm whale habitat? 
## Hypothesis: Sperm whales are found in deep, productive offshore waters. 

Load universal variables
```{r}
#Values used for file and directory names
survey = 'AllSurveys'
gridsize = 10
loctype = 'allsw'
loctype2 = 'all'

```

Load data
```{r}
PmComb <- readRDS(here::here( paste0('output/models/Combined/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda') ))
# add column for log effort as offset #
PmComb$log.effort	= log(PmComb$EffArea)
PmComb <- subset(PmComb, survey != 1642)
```
```{r echo=FALSE, eval=FALSE}
PmComb <- subset(PmComb, chla <= 10)

PmRaw <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_raw.rda') ))
PmRaw$log.effort	= log(PmRaw$EffArea)
PmRaw$d2smt = PmRaw$d2smt/1000 #for plotting purposes

colnames(PmRaw)[c(18:29,44)] <- c('SST', 'chla', 'temp600m', 'wavepower', 'depth', 'slope', 'aspect', 'distland', 'SSH', 'SSHsd', 'EKE', 'distseamt', 'logeffort')

# head(dplyr::arrange(PmComb, desc(chla)))
# PmComb <- subset(PmComb, chla <= 10)

```

```{r, echo=FALSE, eval=FALSE}
#MISC CODE

CompletePreds <- dplyr::select(PmRaw, sst:d2smt)
# PmComb <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_scaledNEW.rda')))

PmRaw$log.effort	= log(PmRaw$EffArea)
PmRaw$d2smt = PmRaw$d2smt/1000 #for plotting purposes

#save if file again if needed
# saveRDS(PmComb, here::here( paste0('output/envData/complete env data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled2.rda') ))  

# nrow(dplyr::filter(PmComb, pa > 0)) #how many presences? was 118, now should be 147

#rename columns for better plots LATER
# colnames(PmComb)[c(18:29,44)] <- c('SST', 'Chla', 'Temp600m', 'Wave Power', 'Depth', 'Slope', 'Aspect', 'Dist to Land', 'SSH', 'SSHsd', 'EKE', 'Dist to Seamt', 'Log Effort')

```
Check correlation of covariates
```{r fig.width =10, fig.height = 6}
require(corrplot)
corrplot.mixed(cor(PmComb[,18:29]), upper="number", lower="circle")

# Are all correlation coefficients < |0.6|?
abs(cor(PmComb[,18:29])) <= 0.6

```

### KS tests 
I compared the distributions of environmental data between the whales and the absences. 
Plots are attached in separate powerpoint. In summary, temperature at 600 m, SSH, and chlorophyll were the only variables with significantly different distributions (p-value < 0.05). However, the D statistics were close to zero (D ~ 0.1) for each, indicating that although the distributions were different, they were not that far apart. The plots also show how similar the general shape of the distributions are between where the whales were observed and where they were not observed.
```{r echo=FALSE, eval=FALSE}
# par(mfrow = c(1,1))
# with(CompleteTotal, plot(pa~bath_m))
require(Cairo)
require(ggplot2)
source('./code/multiplot.r') 

 
loopVec <- 18:29  #columns from PmComb to plot
 
par(mfrow = c(1, 3), mar=c(3,3,1,1), oma=c(0,0,3,1))
 
for (j in loopVec){
   #use original values for plots with PmRaw
   datPlot <- PmComb[,c(1,j)]  #get dataframe of pa and variable j
   datPlot$bin <- 0
   datPlot$bin <- ifelse(datPlot$pa >0 , 1, 0) #put 1s in color col for plotting later
   datPlot <- select(datPlot, bin, 2, pa)
    #normal plot
   # plot(datPlot[,2], datPlot[,1], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'whales per cell')
   

   ptplt <-  ggplot(datPlot, aes(datPlot[,2],datPlot[,1], colour=as.factor(datPlot[,1]))) +
       geom_point() +   
     theme(legend.position ="none") +
       xlab(paste(colnames(datPlot)[2], 'scaled' )) +
       ylab('whales per cell')+
       # scale_colour_discrete(name='',labels = c('Whales Absent', 'Whales Present'))+
theme_bw() +
  guides(colour=FALSE)
   
   #histogram
   # hist(datPlot[,2], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'frequency', main='' )
   
   histplt <- ggplot(datPlot, aes(datPlot[,2], fill=as.factor(datPlot[,1]))) +
     geom_histogram(colour='black', binwidth = 0.5) +
    xlab(paste(colnames(datPlot)[2],'scaled' )) +
    ylab('Frequency') +
     scale_fill_discrete(name = '', labels = c('Whales Absent', 'Whales Present'))+
     theme_bw() +
     theme(legend.position ="top",
           legend.text=element_text(size=16)) 

   
   #do KS test and plot
   datPlot1 <- filter(datPlot, pa >0)
   datPlot0 <- filter(datPlot, pa ==0)

   datKS <- ks.test(datPlot1[,2], datPlot0[, 2])
     cdf0 <- ecdf(datPlot0[,2]); cdf1 <- ecdf(datPlot1[,2]) #make cdfs for the absences and whales for that variable
   #find min and max stats to draw the line between the greatest distance between the distributions
   minMax <- seq(min(datPlot0[,2], datPlot1[,2]), max(datPlot0[,2], datPlot1[,2]), length.out=length(datPlot0[,2])) 
   x0 <- minMax[which( abs(cdf0(minMax) - cdf1(minMax)) == max(abs(cdf0(minMax) - cdf1(minMax))) )] 
   y0 <- cdf0(x0) #sometimes there are 2 points with the same minmax, take the first one 
   y1 <- cdf1(x0)
   
  ksplt <- ggplot(datPlot, aes(x = datPlot[,2], group = datPlot[,1], color = as.factor(datPlot[,1]))) + #
  stat_ecdf(size=1) +
    theme_bw() +
    theme(legend.position ="none", plot.title = element_text(hjust= 0.5)) +
    xlab(paste(colnames(datPlot)[2],'scaled') ) +
    ylab("ECDF") +
    #geom_line(size=1) +
    geom_segment(aes(x = x0[1], y = y0[1], xend = x0[1], yend = y1[1]),
        linetype = "dashed", color = "red") +
    geom_point(aes(x = x0[1] , y= y0[1]), color="red", size=2) +
    geom_point(aes(x = x0[1] , y= y1[1]), color="red", size=2) +
    ggtitle("K-S Test") +
    # theme(legend.title=element_blank()) +
     scale_colour_discrete(name = 'Distribution Type', labels = c('Whales Absent', 'Whales Present'))+
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.1, hjust=.2,label=paste('D=',round(datKS[1]$statistic,2))) +
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.05, hjust=.2,label=paste('p=',round(datKS$p.value,3)))

  
Cairo(here::here('figures', paste0('DataPlotsScaled_', gridsize, 'km_', loctype,'_', colnames(datPlot)[2], '.png')), width=14,height=6,
    units='in', dpi=300, bg='white')
multiplot(ptplt, histplt, ksplt, cols=3)
dev.off()
  
   # ggsave(paste('DataPlots_',colnames(datPlot)[2],'.png'))
  
  }

```


### Data Splitting 
Split the data into train and test sets
```{r eval=FALSE}
require(dplyr)
splitdf <- function(dataframe, seed=NULL) {
    if (!is.null(seed)) set.seed(seed)
    index <- 1:nrow(dataframe)
    trainindex <- sample(index, trunc(length(index)*0.7))
    trainset <- dataframe[trainindex, ]
    testset <- dataframe[-trainindex, ]
    list(trainset=trainset,testset=testset)
}
trainComb = NULL
testComb = NULL
for (s in c(1641, 1303, 1604, 1705, 1706)){
   
 trSub <- filter(PmComb, survey == s)

 #subset for presences and split 70/30
 pres1 <- filter(trSub, pa > 0 )
 listPres  <- splitdf(pres1, 888) #output is list for train and test

 #subset for absences and split 70/30
 abs0  <- filter(trSub, pa == 0 )
 listAbs <- splitdf(abs0, 888)  #output is list for train and test

 #combine train data for presence and absence
 trainAll <- rbind( listPres$trainset, listAbs$trainset )
 
 #combine test data for presence and absence
 testAll  <- rbind( listPres$testset,  listAbs$testset  )
   
trainComb = rbind( trainComb, trainAll )
testComb  = rbind( testComb,  testAll)

# trainAcOnly$log.effort <- log(trainAcOnly$EffArea)
# testAcOnly$log.effort <- log(testAcOnly$EffArea)
}
saveRDS(trainComb, here::here(  paste0('output/models/Combined/data/Train_', gridsize, 'km_', loctype2, '_Comb.rda')  ))
saveRDS(testComb, here::here(  paste0('output/models/Combined/data/Test_', gridsize, 'km_', loctype2, '_Comb.rda')  ))

# nrow(dplyr::filter(trainAcOnly, trainAcOnly$pa >0))
# nrow(dplyr::filter(testAcOnly, testAcOnly$pa >0))
```

## Generalized Additive Models
The data are treated as count data, number of sperm whale encounters per cell, and we used the Tweedie distribution since it has been shown to work well when fewer positive responses exist within the data.
We used thin-plate regression splines (the default basis) for the smoothers of the environmental predictors. Each smoother was limited to 3 degrees of freedom (k=3) to reduce overfitting parameters per recommendations from other studies building similar types of cetaceans distribution models.The log of the effort was included as an offset to account for the variation in effort per cell.


### 10 km spatial scale, Combined Models using sighting and acoustic data


```{r, eval=FALSE, echo=FALSE}
# No knot contrain

#No 2D smoother for longitude and latitude
tw25 <- gam(pa ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + offset(log.effort), data = PmComb, family = tw)
summary(tw25)
par(mfrow = c(1,2))
plot(tw25)
gam.check(tw25)

tw25LL <- gam(pa ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + s(Longitude, Latitude) + offset(log.effort), data = PmComb, family = tw)
summary(tw25LL)
par(mfrow = c(1,2))
plot(tw25LL)
gam.check(tw25)

# tw25_p <- predict_gam(tw25)

```

* Knots contrained to k=3 according to literature on cetacean distribution models.
* Automatic term selection is uses an additional penalty term when determining the smoothness of the function ('select' argument = TRUE)..
* We excluded all non-significant variables (alpha=0.05) and refit the models until all variables were significant.
* REML is restricted maximum likelihood used to optimize the parameter estimates.

Load training and test data
```{r}
trainComb <- readRDS(here::here(  paste0('output/models/Combined/data/Train_',   gridsize, 'km_', loctype2, '_Comb.rda')  ))
testComb <- readRDS(here::here(  paste0('output/models/Combined/data/Test_',     gridsize, 'km_', loctype2, '_Comb.rda')  ))

```

## __Model Selection__

### FULL MODEL  
+ does not include spatial smoother
```{r}
# * Does NOT include sighted acoustic encounters OR spatial smoother
require(mgcv)
twComb <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3)  + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainComb, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twComb)
        
# + s(slope, k=3) + s(aspect, k=3)       # removed 9/27 
```

No rug plot for 10 km resolution?? Too many datapoints?
```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twComb, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'orange', all.terms = TRUE, main='Full Combined, 10 km')
```


#### MODEL DIAGNOSTICS  
The stripe at the bottom left of the residuals vs. fitted values (linear predictor) corresponds to the zeros.

```{r echo=FALSE, tidy=TRUE}
# model diagnostics
gam.check(twComb)
```



### REDUCE MODEL PARAMETERS
   * Removed non-significant variables:   
      + depth 
      + distance to land  
      + eke  
```{r}
twCombb <- gam(pa ~  s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3)  + s(wavepow, k=3) + offset(log.effort), data = trainComb, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twCombb)

```

```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twCombb, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'orange', all.terms = TRUE, main='Reduced Combined, 10 km')
#### Explained Deviance = `round(((twS999b$null.deviance-twS999b$deviance)/twS999b$null.deviance)*100, 1)`
```
Whale occurrence has a negative relationship with chla, SST, and SSHsd. There's a more complex relationship with distance to seamounts, temp at 600m, SSH, and wave power.


#### MODEL DIAGNOSTICS  

```{r echo=FALSE, tidy=TRUE}
# model diagnostics
gam.check(twCombb)
```


### Include 2D Lat-Lon  
Includes 2D Lat-Lon smoother to account for spatial structure in the data and fit the spatial variation not explained by the other predictors  
  * Notice that the temperature at 600m is no longer significant compared to the previous models  
  * Chlorophyll and SSHsd remain significant  
    + Does this indicate that they aren't spatially structured and are independent of location? 

```{r}
twCombLL <- gam(pa ~ s(Longitude, Latitude) + s(bath_m, k=3) + s(dist, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainComb, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twCombLL)
```


```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twCombLL, residuals = FALSE, pch = 20, cex = 0.25,
shade = T, scheme = 1, theta= 45, phi=40, shade.col = 'orange', all.terms = TRUE, main='Full Combined w/ 2D smoother-10km')
```

```{r echo=FALSE, fig.cap = "Purple dots represent acoustically detected encounters."}  
par(mfrow = c(1,1))
plot(twCombLL, select = 1, scheme = 2, lwd = 2, main='2D Spatial Smoother') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(trainComb, pa >0) #get points for whales to plot

#for 2D smoother plot
points(whales$Longitude, whales$Latitude, pch = 20, col='blueviolet', lwd=2, cex=0.75)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)
```


#### MODEL DIAGNOSTICS  
```{r echo=FALSE, tidy=TRUE}
# model diagnostics
par(mar=c(4,4,3,3),mfrow = c(2,2))
gam.check(twCombLL)
```

### REDUCE MODEL PARAMETERS  
   * Removed non-significant variables:   
      + depth
      + distance to land  
      + distance to seamount  
      + SSH  
      + EKE  
      + wave power  
```{r}
twCombLLb <- gam(pa ~ s(Longitude, Latitude) + s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(sshsd, k=3) + offset(log.effort), data = trainComb, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twCombLLb)
```

```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twCombLLb, pages = , residuals = FALSE, pch = 20, cex = 0.25,
shade = T, scheme = 1, theta= 30, phi=20, shade.col = 'orange', all.terms = TRUE, main='Reduced Combined, 2D smoother-10km')
```

```{r echo=FALSE, fig.cap = "Purple dots represent acoustically detected encounters. Black dots are all data points." }  
par(mfrow = c(1,1))
plot(twCombLLb, select = 1, scheme = 2, lwd = 2, main='2D Spatial Smoother') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(trainComb, pa >0) #get points for whales to plot

#for 2D smoother plot
points(whales$Longitude, whales$Latitude, pch = 20, col='blueviolet', lwd=2, cex=0.75)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)
```



### Conclusions
##### It's interesting that these models include a couple more dynamic variables with few static variables compared to the Acoustics Only models. The Combined models also pick up the same general pattern in whale probability using the spatial smoother, but with more variation in the western portion of the archipelago given the addition of a few more encounters from the sighting data. I think the sighting data add more information, but also don't show us anything drastically different.

$~$

#### NEED HELP HERE: Trying to sort out how to use the test data set for prediction purposes and how to use those results to evaluate model performance.


Predict Test Data
```{r eval=FALSE}

trainComb$predicted <- predict(twS999LLb, data = testS999)
```
```{r eval=FALSE}
lons <- seq(180, 210, 1/4)
lats <- seq(15, 33, 1/4)
zz <- matrix(NA, nrow = 98, ncol = 98) #nrow and ncol are sqrt(max(effCell))
zz[trainComb$effCells] <- trainComb$predicted

image(lons, lats, zz)

```

