---
title: "Acoustics-Only GAMs"
author: "Yvonne Barkley"
date: "9/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = '', fig.width =8, fig.height = 6, message=FALSE)
```


Load libraries
```{r, echo=T, eval=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(geoR)
library(tidymv)
library(here)
```


Load universal variables
```{r}
#Values used for file and directory names
survey = 'AllSurveys'
gridsize = 25
loctype = 'AcOnly'
loctype2 = 'Ac'

```


###Load data
```{r}
PmScaled <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda') ))
# add column for log effort as offset #
PmScaled$log.effort	= log(PmScaled$EffArea)

```
```{r echo=FALSE}
PmScaled <- subset(PmScaled, chla <= 10)

PmRaw <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_raw.rda') ))
PmRaw$log.effort	= log(PmRaw$EffArea)
PmRaw$d2smt = PmRaw$d2smt/1000 #for plotting purposes

colnames(PmRaw)[c(18:29,44)] <- c('SST', 'chla', 'temp600m', 'wavepower', 'depth', 'slope', 'aspect', 'distland', 'SSH', 'SSHsd', 'EKE', 'distseamt', 'logeffort')

# head(dplyr::arrange(PmScaled, desc(chla)))
# PmScaled <- subset(PmScaled, chla <= 10)

```

```{r, echo=FALSE, eval=FALSE}
#MISC CODE

CompletePreds <- dplyr::select(PmRaw, sst:d2smt)
# PmScaled <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_scaledNEW.rda')))

PmRaw$log.effort	= log(PmRaw$EffArea)
PmRaw$d2smt = PmRaw$d2smt/1000 #for plotting purposes

#save if file again if needed
# saveRDS(PmScaled, here::here( paste0('output/envData/complete env data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled2.rda') ))  

# nrow(dplyr::filter(PmScaled, pa > 0)) #how many presences? was 118, now should be 147

#rename columns for better plots LATER
# colnames(PmScaled)[c(18:29,44)] <- c('SST', 'Chla', 'Temp600m', 'Wave Power', 'Depth', 'Slope', 'Aspect', 'Dist to Land', 'SSH', 'SSHsd', 'EKE', 'Dist to Seamt', 'Log Effort')

```
Check correlation of covariates
```{r fig.width =10, fig.height = 6}
require(corrplot)
corrplot.mixed(cor(PmScaled[,18:29]), upper="number", lower="circle")

# Are all correlation coefficients < |0.6|?
abs(cor(PmScaled[,18:29])) <= 0.6

```

### KS tests 
I compared the distributions of environmental data between the whales and the absences. 
Plots are attached in separate powerpoint. In summary, temperature at 600 m, SSH, and chlorophyll were the only variables with significantly different distributions (p-value < 0.05). However, the D statistics were close to zero (D ~ 0.1) for each, indicating that although the distributions were different, they were not that far apart. The plots also show how similar the general shape of the distributions are between where the whales were observed and where they were not observed.
```{r echo=FALSE, eval=FALSE}
# par(mfrow = c(1,1))
# with(CompleteTotal, plot(pa~bath_m))
require(Cairo)
require(ggplot2)
source('./code/multiplot.r') 

 
loopVec <- 18:29  #columns from PmScaled to plot
 
par(mfrow = c(1, 3), mar=c(3,3,1,1), oma=c(0,0,3,1))
 
for (j in loopVec){
   #use original values for plots with PmRaw
   datPlot <- PmScaled[,c(1,j)]  #get dataframe of pa and variable j
   datPlot$bin <- 0
   datPlot$bin <- ifelse(datPlot$pa >0 , 1, 0) #put 1s in color col for plotting later
   datPlot <- select(datPlot, bin, 2, pa)
    #normal plot
   # plot(datPlot[,2], datPlot[,1], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'whales per cell')
   

   ptplt <-  ggplot(datPlot, aes(datPlot[,2],datPlot[,1], colour=as.factor(datPlot[,1]))) +
       geom_point() +   
     theme(legend.position ="none") +
       xlab(paste(colnames(datPlot)[2], 'scaled' )) +
       ylab('whales per cell')+
       # scale_colour_discrete(name='',labels = c('Whales Absent', 'Whales Present'))+
theme_bw() +
  guides(colour=FALSE)
   
   #histogram
   # hist(datPlot[,2], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'frequency', main='' )
   
   histplt <- ggplot(datPlot, aes(datPlot[,2], fill=as.factor(datPlot[,1]))) +
     geom_histogram(colour='black', binwidth = 0.5) +
    xlab(paste(colnames(datPlot)[2],'scaled' )) +
    ylab('Frequency') +
     scale_fill_discrete(name = '', labels = c('Whales Absent', 'Whales Present'))+
     theme_bw() +
     theme(legend.position ="top",
           legend.text=element_text(size=16)) 

   
   #do KS test and plot
   datPlot1 <- filter(datPlot, pa >0)
   datPlot0 <- filter(datPlot, pa ==0)

   datKS <- ks.test(datPlot1[,2], datPlot0[, 2])
     cdf0 <- ecdf(datPlot0[,2]); cdf1 <- ecdf(datPlot1[,2]) #make cdfs for the absences and whales for that variable
   #find min and max stats to draw the line between the greatest distance between the distributions
   minMax <- seq(min(datPlot0[,2], datPlot1[,2]), max(datPlot0[,2], datPlot1[,2]), length.out=length(datPlot0[,2])) 
   x0 <- minMax[which( abs(cdf0(minMax) - cdf1(minMax)) == max(abs(cdf0(minMax) - cdf1(minMax))) )] 
   y0 <- cdf0(x0) #sometimes there are 2 points with the same minmax, take the first one 
   y1 <- cdf1(x0)
   
  ksplt <- ggplot(datPlot, aes(x = datPlot[,2], group = datPlot[,1], color = as.factor(datPlot[,1]))) + #
  stat_ecdf(size=1) +
    theme_bw() +
    theme(legend.position ="none", plot.title = element_text(hjust= 0.5)) +
    xlab(paste(colnames(datPlot)[2],'scaled') ) +
    ylab("ECDF") +
    #geom_line(size=1) +
    geom_segment(aes(x = x0[1], y = y0[1], xend = x0[1], yend = y1[1]),
        linetype = "dashed", color = "red") +
    geom_point(aes(x = x0[1] , y= y0[1]), color="red", size=2) +
    geom_point(aes(x = x0[1] , y= y1[1]), color="red", size=2) +
    ggtitle("K-S Test") +
    # theme(legend.title=element_blank()) +
     scale_colour_discrete(name = 'Distribution Type', labels = c('Whales Absent', 'Whales Present'))+
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.1, hjust=.2,label=paste('D=',round(datKS[1]$statistic,2))) +
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.05, hjust=.2,label=paste('p=',round(datKS$p.value,3)))

  
Cairo(here::here('figures', paste0('DataPlotsScaled_', gridsize, 'km_', loctype,'_', colnames(datPlot)[2], '.png')), width=14,height=6,
    units='in', dpi=300, bg='white')
multiplot(ptplt, histplt, ksplt, cols=3)
dev.off()
  
   # ggsave(paste('DataPlots_',colnames(datPlot)[2],'.png'))
  
  }

```


### Data Splitting 
Split the data into train and test sets
```{r eval=FALSE}
require(dplyr)
splitdf <- function(dataframe, seed=NULL) {
    if (!is.null(seed)) set.seed(seed)
    index <- 1:nrow(dataframe)
    trainindex <- sample(index, trunc(length(index)*0.7))
    trainset <- dataframe[trainindex, ]
    testset <- dataframe[-trainindex, ]
    list(trainset=trainset,testset=testset)
}
trainAcOnly = NULL
testAcOnly = NULL
for (s in c(1641, 1303, 1604, 1705, 1706)){
   
 trSub <- filter(PmScaled, survey == s)

 #subset for presences and split 70/30
 pres1 <- filter(trSub, pa > 0 & sid == 999 & loc == 1)
 listPres  <- splitdf(pres1, 777) #output is list for train and test

 #subset for absences and split 70/30
 abs0  <- filter(trSub, pa == 0 )
 listAbs <- splitdf(abs0, 777)  #output is list for train and test

 #combine train data for presence and absence
 trainAll <- rbind( listPres$trainset, listAbs$trainset )
 
 #combine test data for presence and absence
 testAll  <- rbind( listPres$testset,  listAbs$testset  )
   
trainAcOnly = rbind( trainAcOnly, trainAll )
testAcOnly  = rbind( testAcOnly,  testAll)

# trainAcOnly$log.effort <- log(trainAcOnly$EffArea)
# testAcOnly$log.effort <- log(testAcOnly$EffArea)
}
saveRDS(trainAcOnly, here::here(  paste0('output/models/',loctype, '/data/Train_', gridsize, 'km_', loctype2, '_Loc.rda')  ))
saveRDS(testAcOnly, here::here(  paste0('output/models/',loctype, '/data/Test_', gridsize, 'km_', loctype2, '_Loc.rda')  ))

# nrow(dplyr::filter(trainAcOnly, trainAcOnly$pa >0))
# nrow(dplyr::filter(testAcOnly, testAcOnly$pa >0))
```

## Generalized Additive Models
The data are treated as count data, number of sperm whale encounters per cell, and we used the Tweedie distribution since it has been shown to work well when fewer positive responses exist within the data.
We used thin-plate regression splines (the default basis) for the smoothers of the environmental predictors. Each smoother was limited to 3 degrees of freedom (k=3) to reduce overfitting parameters per recommendations from other studies building similar types of cetaceans distribution models.The log of the effort was included as an offset to account for the variation in effort per cell.


### Tweedie - 25 km spatial scale


```{r, eval=FALSE, echo=FALSE}
# No knot contrain

#No 2D smoother for longitude and latitude
tw25 <- gam(pa ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + offset(log.effort), data = PmScaled, family = tw)
summary(tw25)
par(mfrow = c(1,2))
plot(tw25)
gam.check(tw25)

tw25LL <- gam(pa ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + s(Longitude, Latitude) + offset(log.effort), data = PmScaled, family = tw)
summary(tw25LL)
par(mfrow = c(1,2))
plot(tw25LL)
gam.check(tw25)

# tw25_p <- predict_gam(tw25)

```

* Knots contrained to k=3 according to literature on cetacean distribution models.
* Automatic term selection is uses an additional penalty term when determining the smoothness of the function ('select' argument = TRUE)..
* We excluded all non-significant variables (alpha=0.05) and refit the models until all variables were significant.
* REML is restricted maximum likelihood used to optimize the parameter estimates.

Load training and test data
```{r}
trainAcOnly <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Train_', gridsize, 'km_', loctype2, '.rda')  ))
testAcOnly <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Test_',   gridsize, 'km_', loctype2, '.rda')  ))
trainS999 <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Train_',   gridsize, 'km_', loctype2, '_S999.rda')  ))
testS999 <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Test_',     gridsize, 'km_', loctype2, '_S999.rda')  ))
trainLoc <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Train_',   gridsize, 'km_', loctype2, '_Loc.rda')  ))
testLoc <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Test_',     gridsize, 'km_', loctype2, '_Loc.rda')  ))
```

#### Model Selection
* Includes Sighted Acoustic Encounters
(Filtered chla for values < 10 (scaled))
```{r}
require(mgcv)
#with training dataset
twFull <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(aspect, k=3) + offset(log.effort), data = trainAcOnly, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFull)

```
```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFull, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'lavender', all.terms = TRUE, main='Training Dataset, Sightings Included')

# model diagnostics
gam.check(twFull)
```


```{r eval=FALSE, echo=FALSE}
# * Includes Sighted Acoustic Encounters AND __ASPECT__
# (Filtered chla for values < 10 (scaled))
#with training dataset
twFullasp <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(aspect, k=3) + offset(log.effort), data = trainAcOnly, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFullasp)
# 
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullasp, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'slateblue1', all.terms = TRUE, main='Training Dataset, Sightings Included')

# model diagnostics
gam.check(twFullasp)
```


* Does NOT include sighted acoustic encounters OR __ASPECT__
```{r}
require(mgcv)
#with training dataset
twS99 <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS99)

```
```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twS99, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'turquoise2', all.terms = TRUE, main='Training Dataset, No Sightings')

# model diagnostics
gam.check(twS99)
```

* Does NOT include sighted OR unlocalized acoustic encounters
```{r}
require(mgcv)
#with training dataset
twLoc <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainLoc, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twLoc)

```
```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twLoc, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'yellow', all.terms = TRUE, main='Training Dataset, No Sightings')

# model diagnostics
gam.check(twLoc)
```


```{r echo=FALSE, eval=FALSE}
# * Does NOT include sighted acoustic encounters, DOES include __ASPECT__
#with training dataset
twS99asp <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(aspect, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS99asp)
# 
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twS99asp, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'turquoise4', all.terms = TRUE, main='Training Dataset, No Sightings')

# model diagnostics
gam.check(twS99asp)
```




*With full dataset and __unfiltered chla__
```{r eval=FALSE, echo=FALSE}

twFullb <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = PmScaled, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFullb)
#```

#```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullb, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'hotpink', all.terms = TRUE, main='Full Dataset, All Chla')

# model diagnostics
gam.check(twFullb)
```


*With full dataset and __filtered chla__
```{r eval=FALSE}
PmScaled2 <- subset(PmScaled, chla < 10)

twFullc <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = PmScaled2, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFullc)
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullc, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'turquoise2', all.terms = TRUE, main='Full Dataset, Filtered Chla')

# model diagnostics
gam.check(twFullc)
```

```{r echo=FALSE,eval=FALSE}
#finding the raw values do not give same answer as scaled, just slightly different.
twFullr <- gam(pa ~ s(depth, k=3)+s(distland, k=3) + s(slope, k=3) + s(distseamt, k=3) + s(SST, k=3) + s(chlA, k=3) + s(temp600m, k=3 ) + s(SSH, k=3) + s(SSHsd, k=3) + s(EKE, k=3) + s(wavepower, k=3) + offset(logeffort), data = PmRaw, family = tw, link = 'log', select = TRUE, method = "REML")
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullr, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'lavender', all.terms = TRUE)
```

#### Models including only static variables
What are the effects of the static/geographic variables on sperm whale occurrence?
```{r echo=FALSE,eval=FALSE}
require(mgcv)
twStat <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) + offset(log.effort), data = PmScaled, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twStat)
```
```{r echo=FALSE,eval=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twStatraw, pages = 1, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'lemonchiffon3', all.terms = TRUE)
gam.check(twStat)

```

```{r echo=FALSE, eval=FALSE}
twStatraw <- gam(pa ~ s(depth, k=3)+s(distland, k=3) + s(slope, k=3) + s(distseamt, k=3) + offset(logeffort), data = PmRaw, family = tw, link = 'log', select = TRUE, method = "REML")

par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twStatraw, pages = 1, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'darkseagreen2', all.terms = TRUE)


# + s(sst, k=3) + s(chla, k=3) + s(temp600) + s(slope, k=3) + s(aspect, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(d2smt, k=3) + s(dist, k=3) + 
```


#### Models only including dynamic variables to evaluate how well they explain 
```{r echo=FALSE,eval=FALSE}
twDyn <- gam(pa ~ s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = PmScaled, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twDyn)
```
```{r echo=FALSE, eval=FALSE}

par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twDynraw, pages = 2, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'tan1', all.terms = TRUE, ylim = c(-20, 5))

gam.check(twDyn)
```

```{r echo=FALSE, eval=FALSE}
#raw model for plotting
twDynraw <- gam(pa ~ s(SST, k=3) + s(chlA, k=3) + s(temp600m, k=3 ) + s(SSH, k=3) + s(SSHsd, k=3) + s(EKE, k=3) + s(wavepower, k=3) + offset(logeffort), data = PmRaw, family = tw, link = 'log', select = TRUE, method = "REML")
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twDynraw, pages = 2, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'darkseagreen2', all.terms = TRUE, ylim = c(-20, 5))
```


