---
title: 'Tweedie GAMs - Acoustics Only - 25 km '
author: "Yvonne Barkley"
date: "9/26/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = '', fig.width =8, fig.height = 6, message=FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Technical coding questions:
* what's the easiest way to make the partial residual plots using the raw data when I'm running the models with scaled data?
* I need help with the predict() function for predicting the test data. I thought it was straight-forward but I'm misunderstanding how it works.


Load libraries
```{r, echo=T, eval=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(geoR)
library(tidymv)
library(here)
library(mgcViz)
```


# Research question:
## What environmental variables characterize sperm whale habitat? 
## Hypothesis: Sperm whales are found in deep, productive offshore waters. 


Include more details about what to expect in this document



Load universal variables
```{r}
#Values used for file and directory names
survey = 'AllSurveys'
gridsize = 25
loctype = 'AcOnly'
loctype2 = 'Ac'

```

Load data from 'models/data' folder
```{r}
PmScaled <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda') ))
# add column for log effort as offset #
PmScaled$log.effort	= log(PmScaled$EffArea)
PmScaled <- subset(PmScaled, chla <= 10) #some outliers in a handful of absences

```
```{r eval=FALSE,echo=FALSE}

PmRaw <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_raw.rda') ))
PmRaw$log.effort	= log(PmRaw$EffArea)
PmRaw$d2smt = PmRaw$d2smt/1000 #for plotting purposes

colnames(PmRaw)[c(18:29,44)] <- c('SST.r', 'chla.r', 'temp600m.r', 'wavepower.r', 'depth.r', 'slope.r', 'aspect.r', 'distland.r', 'SSH.r', 'SSHsd.r', 'EKE.r', 'distseamt.r', 'logeffort.r')

PmScaled <- data.frame(PmScaled, PmRaw[, c(18:29,44)])
# head(dplyr::arrange(PmScaled, desc(chla)))
# PmScaled <- subset(PmScaled, chla <= 10)

```

```{r, echo=FALSE, eval=FALSE}
#MISC CODE

CompletePreds <- dplyr::select(PmRaw, sst:d2smt)
# PmScaled <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_scaledNEW.rda')))

PmRaw$log.effort	= log(PmRaw$EffArea)
PmRaw$d2smt = PmRaw$d2smt/1000 #for plotting purposes

#save if file again if needed
# saveRDS(PmScaled, here::here( paste0('output/envData/complete env data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled2.rda') ))  

# nrow(dplyr::filter(PmScaled, pa > 0)) #how many presences? was 118, now should be 147

#rename columns for better plots LATER
# colnames(PmScaled)[c(18:29,44)] <- c('SST', 'Chla', 'Temp600m', 'Wave Power', 'Depth', 'Slope', 'Aspect', 'Dist to Land', 'SSH', 'SSHsd', 'EKE', 'Dist to Seamt', 'Log Effort')

```
Check correlation of covariates
```{r fig.width =10, fig.height = 6}
require(corrplot)
corrplot.mixed(cor(PmScaled[,18:29]), upper="number", lower="circle")

# Are all correlation coefficients < |0.6|?
abs(cor(PmScaled[,18:29])) <= 0.6

```

### KS tests 
I compared the distributions of environmental data between the whales and the absences. 
Plots are attached in separate powerpoint. In summary, temperature at 600 m, SSH, and chlorophyll were the only variables with significantly different distributions (p-value < 0.05). However, the D statistics were close to zero (D ~ 0.1) for each, indicating that although the distributions were different, they were not that far apart. The plots also show how similar the general shape of the distributions are between where the whales were observed and where they were absent.
```{r echo=FALSE, eval=FALSE}
# par(mfrow = c(1,1))
# with(CompleteTotal, plot(pa~bath_m))
require(Cairo)
require(ggplot2)
source('./code/multiplot.r') 

 
loopVec <- 18:29  #columns from PmScaled to plot
 
par(mfrow = c(1, 3), mar=c(3,3,1,1), oma=c(0,0,3,1))
 
for (j in loopVec){
   #use original values for plots with PmRaw
   datPlot <- PmScaled[,c(1,j)]  #get dataframe of pa and variable j
   datPlot$bin <- 0
   datPlot$bin <- ifelse(datPlot$pa >0 , 1, 0) #put 1s in color col for plotting later
   datPlot <- select(datPlot, bin, 2, pa)
    #normal plot
   # plot(datPlot[,2], datPlot[,1], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'whales per cell')
   

   ptplt <-  ggplot(datPlot, aes(datPlot[,2],datPlot[,1], colour=as.factor(datPlot[,1]))) +
       geom_point() +   
     theme(legend.position ="none") +
       xlab(paste(colnames(datPlot)[2], 'scaled' )) +
       ylab('whales per cell')+
       # scale_colour_discrete(name='',labels = c('Whales Absent', 'Whales Present'))+
theme_bw() +
  guides(colour=FALSE)
   
   #histogram
   # hist(datPlot[,2], xlab = paste('scaled',colnames(datPlot)[2]), ylab = 'frequency', main='' )
   
   histplt <- ggplot(datPlot, aes(datPlot[,2], fill=as.factor(datPlot[,1]))) +
     geom_histogram(colour='black', binwidth = 0.5) +
    xlab(paste(colnames(datPlot)[2],'scaled' )) +
    ylab('Frequency') +
     scale_fill_discrete(name = '', labels = c('Whales Absent', 'Whales Present'))+
     theme_bw() +
     theme(legend.position ="top",
           legend.text=element_text(size=16)) 

   
   #do KS test and plot
   datPlot1 <- filter(datPlot, pa >0)
   datPlot0 <- filter(datPlot, pa ==0)

   datKS <- ks.test(datPlot1[,2], datPlot0[, 2])
     cdf0 <- ecdf(datPlot0[,2]); cdf1 <- ecdf(datPlot1[,2]) #make cdfs for the absences and whales for that variable
   #find min and max stats to draw the line between the greatest distance between the distributions
   minMax <- seq(min(datPlot0[,2], datPlot1[,2]), max(datPlot0[,2], datPlot1[,2]), length.out=length(datPlot0[,2])) 
   x0 <- minMax[which( abs(cdf0(minMax) - cdf1(minMax)) == max(abs(cdf0(minMax) - cdf1(minMax))) )] 
   y0 <- cdf0(x0) #sometimes there are 2 points with the same minmax, take the first one 
   y1 <- cdf1(x0)
   
  ksplt <- ggplot(datPlot, aes(x = datPlot[,2], group = datPlot[,1], color = as.factor(datPlot[,1]))) + #
  stat_ecdf(size=1) +
    theme_bw() +
    theme(legend.position ="none", plot.title = element_text(hjust= 0.5)) +
    xlab(paste(colnames(datPlot)[2],'scaled') ) +
    ylab("ECDF") +
    #geom_line(size=1) +
    geom_segment(aes(x = x0[1], y = y0[1], xend = x0[1], yend = y1[1]),
        linetype = "dashed", color = "red") +
    geom_point(aes(x = x0[1] , y= y0[1]), color="red", size=2) +
    geom_point(aes(x = x0[1] , y= y1[1]), color="red", size=2) +
    ggtitle("K-S Test") +
    # theme(legend.title=element_blank()) +
     scale_colour_discrete(name = 'Distribution Type', labels = c('Whales Absent', 'Whales Present'))+
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.1, hjust=.2,label=paste('D=',round(datKS[1]$statistic,2))) +
    annotate("text",x=max(datPlot[,2])-2,y=min(datPlot[,1])+0.05, hjust=.2,label=paste('p=',round(datKS$p.value,3)))

  
Cairo(here::here('figures', paste0('DataPlotsScaled_', gridsize, 'km_', loctype,'_', colnames(datPlot)[2], '.png')), width=14,height=6,
    units='in', dpi=300, bg='white')
multiplot(ptplt, histplt, ksplt, cols=3)
dev.off()
  
   # ggsave(paste('DataPlots_',colnames(datPlot)[2],'.png'))
  
  }

```


### Data Visualization
Histograms showing the general distribution of each environmental predictor for the entire dataset.
```{r}
par(mfrow = c(3, 4), mar=c(3,3,2,1), oma=c(0,0,3,1))
 
dataSet = PmScaled   #raw values

loopVec <-  30:41  #columns from PmScaled to plot
 
 for (j in loopVec){
   
   datPlot <- dataSet[, c(1,j)]
   
   hist(datPlot[,2], main = colnames(datPlot)[2], ylab='frequency', xlab = '')
      # plot(datPlot[,2], datPlot[,1], ylab = 'Whales', xlab = colnames(datPlot)[2])
mtext(paste0("Acoustics Only Data, ", gridsize, 'km grid'), side=3, line=1, outer=TRUE, cex=1, font=1)

 }

# dev.off()
```



```{r, eval=FALSE, fig.cap="Log transformation creates more normal distribution" }
# 
# Let's take the log of the more skewed variables  
#   * chla  
#   * eke  
#   * wave power  
#take the log of some variables that are more skewed
PmScaled$chla.log <- log(PmScaled$chla.r)
PmScaled$eke.log <- log(PmScaled$eke.r)
PmScaled$wavepow.log <- log(PmScaled$wavepower.r)


#plot them
dataSet = PmScaled   #raw values

loopVec <-  57:59  #columns from PmScaled to plot
 
par(mfrow = c(1,3), mar=c(3,3,2,1), oma=c(0,0,3,1))

for (j in loopVec){
   
   datPlot <- dataSet[, c(1,j)]
   
   hist(datPlot[,2], main = colnames(datPlot)[2], ylab='frequency', xlab = '')
      # plot(datPlot[,2], datPlot[,1], ylab = 'Whales', xlab = colnames(datPlot)[2])
mtext(paste0("Acoustics Only Data, ",surveynum,', ', gridsize, 'km grid'), side=3, line=1, outer=TRUE, cex=1, font=1)

 }

```


### Data Splitting 
Split the data into train and test sets
```{r eval=FALSE}
require(dplyr)
splitdf <- function(dataframe, seed=NULL) {
    if (!is.null(seed)) set.seed(seed)
    index <- 1:nrow(dataframe)
    trainindex <- sample(index, trunc(length(index)*0.7))
    trainset <- dataframe[trainindex, ]
    testset <- dataframe[-trainindex, ]
    list(trainset=trainset,testset=testset)
}

trainAcOnly = NULL
testAcOnly = NULL
seed = 1

for (s in c(1641, 1303, 1604, 1705, 1706)){
   
 trSub <- filter(PmScaled, survey == s)

 #subset for presences and split 70/30
 pres1 <- filter(trSub, pa > 0 & sid == 999)# & loc == 1) #for S999 versions
 listPres  <- splitdf(pres1, seed) #output is list for train and test

 #subset for absences and split 70/30
 abs0  <- filter(trSub, pa == 0 )
 listAbs <- splitdf(abs0, seed)  #output is list for train and test

 #combine train data for presence and absence
 trainAll <- rbind( listPres$trainset, listAbs$trainset )
 
 #combine test data for presence and absence
 testAll  <- rbind( listPres$testset,  listAbs$testset  )
   
trainAcOnly = rbind( trainAcOnly, trainAll )
testAcOnly  = rbind( testAcOnly,  testAll)

# trainAcOnly$log.effort <- log(trainAcOnly$EffArea)
# testAcOnly$log.effort <- log(testAcOnly$EffArea)
}
saveRDS(trainAcOnly, here::here(  paste0('output/models/',loctype, '/data/Train_', gridsize, 'km_', loctype2, '_S999b.rda')  ))
saveRDS(testAcOnly, here::here(  paste0('output/models/',loctype, '/data/Test_', gridsize, 'km_', loctype2, '_S999b.rda')  ))

# nrow(dplyr::filter(trainAcOnly, trainAcOnly$pa >0))
# nrow(dplyr::filter(testAcOnly, testAcOnly$pa >0))
```

## Generalized Additive Models
The data are treated as count data, number of sperm whale encounters per cell, and we used the Tweedie distribution since it has been shown to work well when fewer positive responses exist within the data.
We used thin-plate regression splines (the default basis) for the smoothers of the environmental predictors. Each smoother was limited to 3 degrees of freedom (k=3) to reduce overfitting parameters per recommendations from other studies building similar types of cetaceans distribution models.The log of the effort was included as an offset to account for the variation in effort per cell.


### 25 km spatial scale


```{r, eval=FALSE, echo=FALSE}
# No knot contrain

#No 2D smoother for longitude and latitude
tw25 <- gam(pa ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + offset(log.effort), data = PmScaled, family = tw)
summary(tw25)
par(mfrow = c(1,2))
plot(tw25)
gam.check(tw25)

tw25LL <- gam(pa ~ s(bath_m) + s(sst) + s(chla) + s(temp600) + s(slope) + s(aspect) + s(ssh) + s(sshsd) + s(eke) + s(wavepow) + s(d2smt) +s(dist) + s(Longitude, Latitude) + offset(log.effort), data = PmScaled, family = tw)
summary(tw25LL)
par(mfrow = c(1,2))
plot(tw25LL)
gam.check(tw25)

# tw25_p <- predict_gam(tw25)

```

* Knots contrained to k=3 according to literature on cetacean distribution models.
* Automatic term selection is uses an additional penalty term when determining the smoothness of the function ('select' argument = TRUE)..
* We excluded all non-significant variables (alpha=0.05) and refit the models until all variables were significant.
* REML is restricted maximum likelihood used to optimize the parameter estimates.

Load training and test data
```{r}
#seed 1
trainS999 <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Train_',   gridsize, 'km_', loctype2, '_S999b.rda')  ))
testS999 <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Test_',     gridsize, 'km_', loctype2, '_S999b.rda')  ))

```

## __Model Selection__

### FULL MODEL  
Estimate the smoothing parameters for each predictor variable using restricted maximum likelihood (method = 'REML')
+ does not include spatial smoother
+ does not include slope or aspect due to the variation between left and right

Using the scaled values
```{r}
# * Does NOT include sighted acoustic encounters OR spatial smoother
require(mgcv)
twS999 <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3)  + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS999)
        
# + s(slope, k=3) + s(aspect, k=3)       # removed 9/27 
```


```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(4,4))
plot(twS999, pages = 2, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'aquamarine', all.terms = TRUE, main='Full Ac Only-25km')
```


```{r eval=FALSE, echo=FALSE}
# Do log values make a difference? = NO
require(mgcv)
twS999log <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3)  + s(d2smt, k=3) +  s(sst, k=3) + s(chla.log, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke.log, k=3) + s(wavepow.log, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS999log)
        
# + s(slope, k=3) + s(aspect, k=3)       # removed 9/27 
```




#### MODEL DIAGNOSTICS  
The stripe at the bottom left of the residuals vs. fitted values (linear predictor) corresponds to the zeros.

```{r echo=FALSE}
# model diagnostics
gam.check(twS999)
```



### REDUCE MODEL PARAMETERS
   * Removed non-significant variables:   
      + distance to land
      + distance to seamount
      + sst
      + wave power
```{r}
# * Does NOT include sighted acoustic encounters

twS999b <- gam(pa ~ s(bath_m, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS999b)

```

```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(3,2))
plot(twS999b, residuals = FALSE, pch = 20, cex = 0.25, ylim = c(-6,2),
scheme = 1, shade = T, shade.col = 'aquamarine', all.terms = TRUE, main='Reduced Ac Only-25km')
#### Explained Deviance = `round(((twS999b$null.deviance-twS999b$deviance)/twS999b$null.deviance)*100, 1)`

```

* Remove chlorophyll
```{r}
# * Does NOT include sighted acoustic encounters

twS999c <- gam(pa ~ s(bath_m, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS999c)

```

```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(3,2))
plot(twS999c, residuals = FALSE, pch = 20, cex = 0.25, ylim = c(-6,2),
scheme = 1, shade = T, shade.col = 'aquamarine', all.terms = TRUE, main='Reduced Ac Only-25km')
#### Explained Deviance = `round(((twS999b$null.deviance-twS999b$deviance)/twS999b$null.deviance)*100, 1)`

```

#### MODEL DIAGNOSTICS  
  The stripe at the bottom left of the residuals vs. fitted values (linear predictor) corresponds to the zeros.

```{r echo=FALSE}
# model diagnostics
par(mar=c(4,4,3,3),mfrow = c(2,2))
gam.check(twS999b)
```




The final model with no spatial smoother results in an explained deviance of 8.06%. The SSH and EKE are negative effect of SSH and EKE on sperm whale encounters 

$~$

## Include 2D Lat-Lon Smoother  
The 2D Lat-Lon smoother accounts for spatial autocorrelation in the data and fit the spatial variation not explained by the other predictors  
  * Notice that the temperature at 600m is no longer significant compared to the previous models  
  * Chlorophyll and SSHsd remain significant  
    + Does this indicate that they aren't spatially structured and are independent of location? 

```{r}
twS999LL <- gam(pa ~ s(Longitude, Latitude) + s(bath_m, k=3) + s(dist, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS999LL)
```

Full Acoustics Only Model w/ Spatial Smoother  


```{r echo=FALSE}
par(mfrow = c(2, 2), mar=c(4,3,2,1))
# par(mar=c(4,4,3,3),mfrow = c(4,4))
plot(twS999LL,residuals = FALSE, pch = 20, cex = 0.01,
shade = T, scheme =2, theta= 45, phi=40, shade.col = 'aquamarine', all.terms = TRUE)

# mtext(paste0("Full Acoustics Only Model, ", gridsize, 'km grid'), side=3, line=1, outer=TRUE, cex=1, font=1) 
```

```{r echo=FALSE, fig.cap = "Purple dots represent acoustically detected encounters. Black dots are all data points(grid centroids)"}  
par(mfrow = c(1,1))
plot(twS999LL, select = 1, scheme = 2, lwd = 2, main='2D Spatial Smoother') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(trainS999, pa >0) #get points for whales to plot

#for 2D smoother plot
points(whales$Longitude, whales$Latitude, pch = 20, col='blueviolet', lwd=2, cex=0.75)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)
```



#### MODEL DIAGNOSTICS
Check out the residuals and smoother parameters,

```{r }
# model diagnostics
par(mar=c(4,4,3,3),mfrow = c(2,2))
gam.check(twS999LL)
```


### REDUCE MODEL PARAMETERS  
   * Removed non-significant variables:   
      + depth
      + distance to land
      + distance to seamount
      + sst
      + temp at 600 m
      + wave power 
```{r}
twS999LLb <- gam(pa ~ s(Longitude, Latitude) + s(chla, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS999LLb)
```
Reduced Acoustics Only Model w/ Spatial Smoother  
```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twS999LLb, pages = , residuals = FALSE, pch = 20, cex = 0.25,
shade = T, scheme = 2, theta= 30, phi=20, shade.col = 'aquamarine', all.terms = TRUE)
```


* Remove chlorophyll  
```{r}
twS999LLc <- gam(pa ~ s(Longitude, Latitude) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS999LLc)

```
Reduced Acoustics Only Model w/ Spatial Smoother  
```{r echo=FALSE}
# The y-axis is the value taken by the centred smooth. It is the contribution (at a value of the covariate) made to the fitted value for that smooth function.
par(mar=c(4,4,3,3),mfrow = c(2,2))
b <- plot(twS999LLc, residuals = FALSE, pch = 20, cex = 0.25,
shade = T, scheme = 2, theta= 30, phi=20, shade.col = 'aquamarine', all.terms = TRUE)
```


```{r echo=FALSE, fig.cap = "Purple dots represent acoustically detected encounters. Black dots are all data points." }  
par(mfrow = c(1,1))
plot(twS999LLc, select = 1, scheme = 2, lwd = 2, main='2D Spatial Smoother') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(trainS999, pa >0) #get points for whales to plot

#for 2D smoother plot
points(whales$Longitude, whales$Latitude, pch = 20, col='blueviolet', lwd=2, cex=0.75)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)
```

$~$

### Conclusions
##### The first reduced Acoustics Only model included more dynamic variables for the significant smooth terms compared to both models that included a 2D spatial smoother. The former addressed my hypothesis more clearly given that depth, chl a, and SSH/SSHsd were included as important variables. The introduction of the 2D smoother reduces all dynamic variables except for chla and SSHsd. Acoustic encounters of sperm whales indicate sperm whale occurrence is related to productivity in some way whether or not a 2D smoother is included. If the 2D smoother is left out of the model, I should acknowledge how certain important variables may also be spatially autocorrelated. Some papers state that spatial autocorrelation exists, calculate Moran's I to determine the magnitude, and don't do anything more about it, such as Forney et al. 2015. 



## Predict Test Data
```{r tidy=TRUE}

require(magrittr)
require(dplyr)

#### For twS999c, no spatial smoother ####
twTrainFinal <- trainS999 %>% mutate(resid = resid(twS999c), predict = predict(twS999c))
predTrain <- predict.gam(twS999c, type = 'response')  #calculate MSE for these to compare with test set. If they're super different, speaks to the genrality of the model
twTrainFinal$fit <- predTrain

#using scale of 0,1,2 makes this hard to interpret
twMSEtrain <- mean((twTrainFinal$pa - twTrainFinal$fit)^2)  #MSE
# mean(abs((twTrainFinal$pa - twTrainFinal$fit)))  #Mean absolute error
## Calculate MSE AFTER transforming the predictions back to the same scale as the observed data

twPred <- predict.gam(twS999c, newdata = testS999, type = 'response', se.fit = TRUE)
twTestFinal <- data.frame(testS999, fit = twPred$fit, se.fit=twPred$se.fit)
twMSEtest <- mean((twTestFinal$pa - twTestFinal$fit)^2) #MSE

# mean(abs((testFinal$pa - testFinal$fit))) #Mean absolute error


#### For twS999cLL, with spatial smoother ####
#pulling the prediction and residual data from the model
twTrainLL <- trainS999 %>% mutate(resid = resid(twS999LLc), predict = predict(twS999LLc))
predTrainLL <- predict.gam(twS999LLc, type = 'response')  #calculate MSE for these to compare with test set. If they're super different, speaks to the genrality of the model
twTrainLL$fit <- predTrainLL

#using scale of 0,1,2 makes this hard to interpret
twMSEtrainLL <- mean((twTrainLL$pa - twTrainLL$fit)^2)  #MSE
# mean(abs((twTrainFinal$pa - twTrainFinal$fit)))  #Mean absolute error
## Calculate MSE AFTER transforming the predictions back to the same scale as the observed data

twPredLL <- predict.gam(twS999LLc, newdata = testS999, type = 'response', se.fit = TRUE)
twTestLL <- data.frame(testS999, fit = twPredLL$fit, se.fit=twPredLL$se.fit)
twMSEtestLL <- mean((twTestLL$pa - twTestLL$fit)^2) #MSE

# mean(abs((testFinal$pa - testFinal$fit))) #Mean absolute error

# AIC
twAIC <- AIC(twS999c)
twAICLL <- AIC(twS999LLc)

# Explained Deviance
twExpDev = round(((twS999c$null.deviance-twS999c$deviance)/twS999c$null.deviance)*100, 2)
twExpDevLL = round(((twS999LLc$null.deviance-twS999LLc$deviance)/twS999LLc$null.deviance)*100, 2)

```

```{r}
# make summary table of metrics

table = matrix(NA, nrow = 2, ncol = 5)
colnames(table) = c("Model", "ExpDev", "AIC", "MSEtrain", "MSEtest")

# enter info by row

table[1,] <- c("twS999c", paste0(twExpDev, '%'), round(twAIC, 2), round(twMSEtrain,3), round(twMSEtest,3))

table[2,] <- c("twS999LLc", paste0(twExpDevLL, '%'), round(twAICLL,2), round(twMSEtrainLL,3), round(twMSEtestLL,3))
require(knitr)
kable(table, caption = "Tweedie Model Summary Metrics")

```

```{r}
#https://gist.github.com/aperium/9fc737ea311a758328eadf27c2426e47
require(ggplot2)
# require(mgcViz)
# plotting the data.
plot <- ggplot(twTrainFinal, aes(x = depth.r, y = pa)) + 
  geom_point(shape = 19, stroke = 1) +
  xlab("depth (m)") + ylab("whale") +
  theme_bw()
plot %<>% + geom_smooth(aes(y = predict),  method = gam, formula = twS999c, data = trainS999, size = 1, se = FALSE)
# b <- getViz(twS999c)
# o <- plot( sm(b, 1) )
# o + l_fitLine(colour = "red") + l_rug(mapping = aes(x=x, y=y), alpha = 0.8) +
#     l_ciLine(mul = 5, colour = "blue", linetype = 2) + 
#     l_points(shape = 19, size = 1, alpha = 0.1) + theme_classic()



```



### Spatial Autocorrelation in data 
Not model residuals
```{r eval=FALSE}
# https://stats.idre.ucla.edu/r/faq/how-can-i-calculate-morans-i-in-r/

# Create distance matrix, then take the inverse of the matrix values and replace with zeros on diagonal
sw.dists <- as.matrix((dist(cbind(trainS999$Longitude, trainS999$Latitude))))
sw.dists.inv <- 1/sw.dists
diag(sw.dists.inv) <- 0
sw.dists.inv[is.infinite(sw.dists.inv)] <- 0  #turn infinite values to 0

require(ape)
Moran.I(trainS999$pa, sw.dists.inv)

```




```{r echo=FALSE, eval=FALSE}
# * Includes Sighted Acoustic Encounters
# (Filtered chla for values < 10 (scaled))

require(mgcv, eval=FALSE)
#with training dataset
twFull <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(aspect, k=3) + offset(log.effort), data = trainAcOnly, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFull)
# 
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFull, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'lavender', all.terms = TRUE, main='Training Dataset, Sightings Included')

# model diagnostics
gam.check(twFull)
```

```{r eval=FALSE, echo=FALSE}
# * Includes Sighted Acoustic Encounters AND __ASPECT__
# (Filtered chla for values < 10 (scaled))
#with training dataset
twFullasp <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(aspect, k=3) + offset(log.effort), data = trainAcOnly, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFullasp)
# 
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullasp, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'slateblue1', all.terms = TRUE, main='Training Dataset, Sightings Included')

# model diagnostics
gam.check(twFullasp)
```

```{r eval=FALSE, echo=FALSE}
# * Does NOT include sighted acoustic encounters OR __ASPECT__
require(mgcv)
#with training dataset
twS99 <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twS99)
# 
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twS99, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'turquoise2', all.terms = TRUE, main='Training Dataset, No Sightings')

# model diagnostics
gam.check(twS99)
```

```{r eval=FALSE, echo=FALSE}
# * Does NOT include sighted OR unlocalized acoustic encounters

require(mgcv)
#with training dataset
twLoc <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainLoc, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twLoc)
# 
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twLoc, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'yellow', all.terms = TRUE, main='Training Dataset, No Sightings')

# model diagnostics
gam.check(twLoc)
```

```{r eval=FALSE, echo=FALSE}
#*With full dataset and __unfiltered chla__
twFullb <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = PmScaled, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFullb)
#```

#```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullb, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'hotpink', all.terms = TRUE, main='Full Dataset, All Chla')

# model diagnostics
gam.check(twFullb)
```

```{r eval=FALSE,echo=FALSE}
# *With full dataset and __filtered chla__
PmScaled2 <- subset(PmScaled, chla < 10)

twFullc <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) +  s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = PmScaled2, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twFullc)
# ```
# ```{r echo=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullc, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'turquoise2', all.terms = TRUE, main='Full Dataset, Filtered Chla')

# model diagnostics
gam.check(twFullc)
```

```{r echo=FALSE,eval=FALSE}
#finding the raw values do not give same answer as scaled, just slightly different.
twFullr <- gam(pa ~ s(depth, k=3)+s(distland, k=3) + s(slope, k=3) + s(distseamt, k=3) + s(SST, k=3) + s(chlA, k=3) + s(temp600m, k=3 ) + s(SSH, k=3) + s(SSHsd, k=3) + s(EKE, k=3) + s(wavepower, k=3) + offset(logeffort), data = PmRaw, family = tw, link = 'log', select = TRUE, method = "REML")
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twFullr, pages = 3, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'lavender', all.terms = TRUE)
```



```{r echo=FALSE,eval=FALSE}
require(mgcv)
# #### Models including only static variables
# What are the effects of the static/geographic variables on sperm whale occurrence?
#    * Resulted in only ~6% deviance explained for any iteration
#    * Depth, slope, distance to seamount were significant, no aspect included

twStat <- gam(pa ~ s(bath_m, k=3) + s(dist, k=3) + s(slope, k=3) + s(d2smt, k=3) + s(aspect, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twStat)
# 
# ```
# ```{r echo=FALSE,eval=FALSE}
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twStatraw, pages = 1, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'lemonchiffon3', all.terms = TRUE)
gam.check(twStat)

```

```{r echo=FALSE, eval=FALSE}
twStatraw <- gam(pa ~ s(depth, k=3)+s(distland, k=3) + s(slope, k=3) + s(distseamt, k=3) + offset(logeffort), data = PmRaw, family = tw, link = 'log', select = TRUE, method = "REML")

par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twStatraw, pages = 1, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'darkseagreen2', all.terms = TRUE)


# + s(sst, k=3) + s(chla, k=3) + s(temp600) + s(slope, k=3) + s(aspect, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + s(d2smt, k=3) + s(dist, k=3) + 
```



```{r echo=FALSE,eval=FALSE}
# #### Models only including dynamic variables to evaluate their explanatory power alone.
#    * Resulted in 8.6% explained deviance, all variables significant except wave power
twDyn <- gam(pa ~ s(sst, k=3) + s(chla, k=3) + s(temp600, k=3) + s(ssh, k=3) + s(sshsd, k=3) + s(eke, k=3) + s(wavepow, k=3) + offset(log.effort), data = trainS999, family = tw, link = 'log', select = TRUE, method = "REML")
summary(twDyn)
```
```{r echo=FALSE, eval=FALSE}

par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twDynraw, pages = 2, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'tan1', all.terms = TRUE, ylim = c(-20, 5))

gam.check(twDyn)
```

```{r echo=FALSE, eval=FALSE}
#raw model for plotting
twDynraw <- gam(pa ~ s(SST, k=3) + s(chlA, k=3) + s(temp600m, k=3 ) + s(SSH, k=3) + s(SSHsd, k=3) + s(EKE, k=3) + s(wavepower, k=3) + offset(logeffort), data = PmRaw, family = tw, link = 'log', select = TRUE, method = "REML")
par(mar=c(4,4,3,3),mfrow = c(2,2))
plot(twDynraw, pages = 2, residuals = FALSE, pch = 20, cex = 0.25,
scheme = 1, shade = T, shade.col = 'darkseagreen2', all.terms = TRUE, ylim = c(-20, 5))
```


