#####
# Code calculate the amount of trackline that is straight
```{r}


library(here)
library(lubridate) #needs character or numeric to work
library(tidyverse)
library(RcppRoll)

#function to split last 'n' elements in string
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}


```
##################################


#1303 - 20 second intervals
```{r}
#Open gps data for straightPath function

dir_str8path = 'C:\\Users\\yvers\\Documents\\CHP 3\\SpermWhales\\output\\straightPath results\\'
gps1303 <- read.csv(paste0(here('data'), '/GPS_1303.csv'))
```

Survey 1303 & 1641 do not have seconds included in UTC :(
    Have to manipulate the GPSTime column to extract the seconds
#1303 add seconds
```{r}
pattsec <- as.character(substrRight(gps1303$GPSTime, 2))

#some values of GPSTime appear as a single number with no '0' so it effs up the UTC. Gotta add the '0'
pattsec2 <- NULL
for (sec in pattsec){
  if (nchar(sec) == 1) {
    s = paste0('0', sec)
    pattsec2 <- rbind(pattsec2, s)
  } else
      pattsec2 <- rbind(pattsec2, sec)
  }
  
#combine UTC with seconds and convert to POSIXct
gps1303$UTC = mdy_hms(paste0(gps1303$UTC, ':', as.character(pattsec2)))

```

#Straight path for 1303
```{r}
nSmall = 5
nLarge = 26
thld = 20
AllSpeeds <- filter(gps1303, Longitude <= -175.5 & Latitude > 27.2)

AllSpeeds <- filter(gps1303, Longitude >= -162.5 & Longitude <= -161.8 & Latitude < 26)

str8_1303 <- straightPath(gps1303, nSmall, nLarge, thld, plot=TRUE)


write.csv(str8_1303, here::here('output', 'straightPath_1303.csv'), row.names = F)


#filter speeds of 8-10.5 kts from str8AllSpd
str8sub1303 <- filter(str8_1303, Speed >=7 & Speed <= 11)

myPlot <- ggplot(str8subAllSpd, aes(x=Longitude, y=Latitude)) +
  geom_point(aes(group=timeGroup, col=straight), size = 1.3) +
  scale_color_manual(values = c('red', 'darkgreen'))
# scale_size_manual(values = c(2, 1))
print(myPlot)


write.csv(str8sub1303, here::here('output', 'straightPath_1303.csv'), row.names = F)
```

#1641 - 1 min interval
```{r}
# Load data

gps1641 <- read.csv(here::here('data', 'GPS_1641.csv'))

# subset for locations west of -150 W, into the HI EEZ
gps1641 <- filter(gps1641, Longitude <= -150 )


```

#1641 add seconds
```{r}
pattsec <- as.character(substrRight(gps1641$GPSTime, 2))

#some values of GPSTime appear as a single number with no '0' so it effs up the UTC. Gotta add the '0'
pattsec2 <- NULL
for (sec in pattsec){
  if (nchar(sec) == 1) {
    s = paste0('0', sec)
    pattsec2 <- rbind(pattsec2, s)
  } else
      pattsec2 <- rbind(pattsec2, sec)
  }
  
#combine UTC with seconds and convert to POSIXct
gps1641$UTC = mdy_hms(paste0(gps1641$UTC, ':', as.character(pattsec2)))

```


#Straight path for 1641
```{r}
nSmall = 5
nLarge = 15
thld = 20

AllSpeeds <- filter(gps1641, Longitude >= -165 & Longitude <= -163 & Latitude < 24 & Latitude >23.8)

str8_1641 <- straightPath(gps1641, nSmall, nLarge, thld, plot=TRUE)


# write.csv(str8_1641, here::here('output', 'straightPath_1303.csv'), row.names = F)


#filter speeds of 8-10.5 kts from str8AllSpd
str8sub1641 <- filter(str8_1641, Speed >=7 & Speed <= 11)# & Longitude > -165 & Longitude < -162
                      #& Latitude < 24 & Latitude > 21)


myPlot <- ggplot(str8sub1641, aes(x=Longitude, y=Latitude)) +
  geom_point(aes(group=timeGroup, col=straight), size = 1.3) +
  scale_color_manual(values = c('red', 'darkgreen'))
# scale_size_manual(values = c(2, 1))
print(myPlot)

write.csv(str8sub1641, here::here('output', 'straightPath_1641.csv'), row.names = F)
```






#filter gps data for speeds between 8-10 knots
gps1303sub1 <- filter(gps1303, Speed >= 8 & Speed <= 10.5)
gps1303sub1$UTC2<- as.character(gps1303sub1$UTC)

#function to split last 'n' elements in string
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
pattsec <- substrRight(gps1303sub1$GPSTime, 2)

gps1303sub1$UTC = mdy_hms(paste0(gps1303sub1$UTC, ':', pattsec))


#run straightPath from PAMmisc
require(RcppRoll)
str8path <- straightPath(gps1303sub1, 30, 90, 20, plot=TRUE)


#Explore a small subset of GPS data
gps1303sub2 <- filter(gps1303sub1, Longitude <= -175.5 & Latitude > 27.2)
str8path1 <- straightPath(gps1303sub2, 30, 60, 20, plot=TRUE)
str8path2 <- straightPath(gps1303sub2, 15, 30, 20, plot=TRUE)

write.csv(str8path2, here('output', 'TestPath_1303subset.csv'), row.names = F)
write.csv(gps1303sub2, here('output', 'TestPath_1303data.csv'), row.names = F)


big60 <- filter(str8path2, Longitude > -175.67)
big90 <- filter(str8path2, Longitude > -175.67)
big90b <- filter(str8path2, Longitude > -175.67)


#subset for longitudes west of 172
gps1303W172 <- filter(gps1303sub1, Longitude <= -172)
str8pathW172 <- straightPath(gps1303W172, 30, 60, 20, plot=TRUE)


#use all speeds within the same area
gps1303sub3 <- filter(gps1303, Longitude <= -175.5 & Latitude > 27.2)
str8pathALL <- straightPath(gps1303sub3, 30, 90, 20, plot=TRUE)


myPlot2 <- ggplot(AllSpeeds, aes(Longitude, Latitude)) +
  geom_point()
print(myPlot2)

myPlot3 <- ggplot(gps1303, aes(Longitude, Latitude)) +
  geom_path()
print(myPlot3)

patt2 <- substr(gps1303sub1$UTC2, 1, 9) #subset the longest date string
patt2trim <- str_trim(patt2, side = "right") #trim the white space off the shorter strings




dv = patt2trim

hh = floor(gps1303sub1$GPSTime/10000);
mm = floor((gps1303sub1$GPSTime-hh*10000)/100);
ss = floor((gps1303sub1$GPSTime-hh*10000-mm*100));
dv2 = paste0(dv, hh, mm, ss)


require(M3)
patt3 <- combine.date.and.time(date = dv, time=list(hh, mm, ss))
