---
title: "Apply Grid to Data"
author: "Yvonne Barkley"
date: "6/30/2020"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
This code creates the grid and centroids used for calculating the sampling units for the models to incorporate passive acoustic data for any species.

Technical Note: All grids and centroids are calculated separately for each survey to more easily accommodate the different timing and trackline data. However, a more recent change is to combine all of teh centroid data for processing the environmental data for each model type (Combined, Acoustics Only, etc). This may take longer, but it's much easier than splitting up the centroids by survey on separate computers and then having to keep track of everything and consolidate it later, ugh. IT'S MUCH EASIER NOW. 
That said, a loop to load all centroid data is located in the LAST CHUNK of the markdown.


Load functions 
```{r}

source('./code/FromTaiki/gridFunctions_v3.R')
library(PAMmisc)
library(tidyverse)
library(lubridate)
library(Distance)
```

##Load parameters
```{r}
surveynum = 1706
gridsize = 25
loctype = 'Combined'
loctype2 = 'Comb'

```

##Load data if necessary
To revisit a data set:
*load grid or centroid data
*run GPS data again
*filter Pm detections for survey
*get centroids if needed
Then proceed to 'GRID QC' chunk below.
```{r}
#load grid data
gridEffort <- readRDS(here::here( paste0('output/gridEffort/', gridsize, ' km-', loctype, '/gridEffort', surveynum, '_', gridsize, 'km_', loctype2, '_raw.rda') ))

gridEffortNEW <- readRDS(here::here(paste0('output/gridEffort/', gridsize, ' km-', loctype), paste0('gridEffort', surveynum, '_', gridsize, 'km_', loctype2, 'NEW.rda')))
gridEffortOLD <- readRDS(here::here(paste0('output/gridEffort/', gridsize, ' km-', loctype), paste0('gridEffort', surveynum, '_', gridsize, 'km_', loctype2, 'OLD.rda')))
gridEffort <- readRDS(here::here(paste0('output/gridEffort/', gridsize, ' km-', loctype), paste0('gridEffort', surveynum, '_', gridsize, 'km_', loctype2, '_raw.rda')))




#load centroid locations
centroidAll <- readRDS(here::here(paste0('output/centroids/', gridsize, ' km-', loctype), paste0('centroids',  surveynum, '_', gridsize, 'km_', loctype2, '_raw.rda')))

```

##1. GPS data
```{r}
# EXAMPLE WORK FLOW
# 1) First read in gps, then you need to convert longitudes to 0-360,
# convert UTC to POSIXct format, and create a column called "effort" that is TRUE or FALSE
# Column names needed: Longitude, Latitude, UTC, effort
gpsAll <- read.csv(paste0('./output/straightPath results/straightPathWeffort_', surveynum, '.csv'), stringsAsFactors = FALSE) 
unique(gpsAll$join.veffort)
unique(gpsAll$join.aeffort)
#TRIM WHITE SPACE IF NEEDED
gpsAll$join.veffort <- str_trim(gpsAll$join.veffort, side = "right")
gpsAll$join.aeffort <- str_trim(gpsAll$join.aeffort, side = "right")


# gpsAll <- gpsAll %>% mutate_if(is.character, str_trim) #white space after join.veffort!?!?
# gpsAll$effort <- gpsAll$straight == TRUE & (gpsAll$join.aeffort == "on" | gpsAll$join.veffort =="on") # effort used for 'allsw'
gpsAll$effort <- gpsAll$straight == TRUE & gpsAll$join.aeffort == "on" & gpsAll$join.veffort =="on"  # COMBINED EFFORT: Visual and Acoustics BOTH must be on

gpsAll$Longitude <- ifelse(gpsAll$Longitude <= 0, gpsAll$Longitude + 360, gpsAll$Longitude)
gpsAll$UTC <- lubridate::ymd_hms(gpsAll$UTC)
```

```{r}
#EFFORT CHECKING

# Both should be zero for the combined model
#when vis is ON and acoustics is OFF, effort should be FALSE
effVischk=gpsAll[which(gpsAll$join.veffort=='on' & gpsAll$join.aeffort =='off' & gpsAll$straight==TRUE), ]
which(effVischk$effort == TRUE) #should be zero, don't want TRUE effort if ac is off

#when vis is OFF and acoustics is OFF, effort should still be FALSE
effACchk=gpsAll[which(gpsAll$join.veffort=='off' &  gpsAll$join.aeffort=='on'   &  gpsAll$straight==TRUE),]
which(effACchk$effort == TRUE) #should be zero, Don't want this to be false
#Both outputs should be zero

effTotChk=gpsAll[which(gpsAll$straight==TRUE, (gpsAll$join.veffort=='on' &  gpsAll$join.aeffort=='on')), ]


```

###subset GPS for troubleshooting
```{r}

gpsAll <- filter(gpsAll, Id >= 146246 & Id <= 148253)

```



##2. Load sperm whale detection data
Includes localized acoustic encounters and sightings with acoustics only.
```{r}
# 2) Read in detections, convert column names to Longitude, Latitude, UTC,
# and "distance" if you need to create a detection function using Distance package
# Longitude must be matching 0-360 like in the GPS data

pmDets <- read.csv('./output/SpermiesFinalLR_20200919.csv', stringsAsFactors = FALSE)
# pmDets <- filter(SwEnvFinal, loc == 1 | loc == 0 & sid < 999) # localized encs and sightings

#### This is where you change the location from localized position to trackline position ####
pmDets <- rename(pmDets, Longitude =lon2, #lon_trk2
                         Latitude=lat, #lat_trk
                         distance = pdist)

pmDets$Longitude <- ifelse(pmDets$Longitude <= 0, pmDets$Longitude + 360, pmDets$Longitude)
pmDets$UTC <- lubridate::mdy_hm(pmDets$UTC)
pmDets$distance <- abs(pmDets$distance)*1000


# pmDets <- filter(pmDets, type == 'best' | sid < 999, itrk==1)
# pmDetsAc <- filter(pmDets, sid == 999)
# pmDetsub <- filter(pmDets, survey == surveynum, acid >= 71 & acid <= 74)
```

##Make detection function
```{r}
# 3) fit a detection function

#winner! See v2 for other variations
dsmAll_3pcthn <- Distance::ds(pmDetsAc, key = 'hn', truncation = '3%')

plot(dsmAll_3pcthn)

summary(dsmAll_3pcthn)
gof_ds(dsmAll_3pcthn)

detfxn <- dsmAll_3pcthn
summ <- summary(detfxn)

```

##Calculate BIG Grid 
```{r}

# 4) Run grid stuff. This will take a while, and should have a couple progress bars showing you how long the slow parts
# are taking. Set "trunc_m" to whatever truncation distance you might want to use
# It will ask you what pixel size you want  (this is the grid size), in the future you can specify this when you call it

# YB Sept 3, 2020: Make overall larger grid that can be reused for each spatial scale

bigGrid <- doAllGrid(gps = gpsAll,
                     bounds = NULL,
                     dets = pmDetsub,
                     trunc_m = 25e3, #METERS
                     dsmodel = NULL, #NULL or dsm 
                     pixel = NULL,
                     grid = NULL,
                     plot = F)
#Save raw OVERALL grid 
saveRDS(bigGrid, paste0('output/gridEffort/bigGrids/', 'GRID', surveynum, '_', gridsize,'km_', loctype2, '.rda') )
```


##Run grid and centroids
```{r}
#### filter whales by survey 
# pmDetsub <- dplyr::filter(pmDets, survey == surveynum, loc == 1 | loc == 0 & sid < 999 ) # localized encs and sightings
pmDetsub <- dplyr::filter(pmDets, survey == surveynum) #include all whales for the combined models

# LOAD OVERALL GRID IF NEEDED 
# This is the overall grid - use 25km trunc_m
# bigGrid <- readRDS( paste0('output/gridEffort/bigGrids/', 'GRID', surveynum, '_', gridsize,'km_', loctype2, '.rda') )
bigGrid <- readRDS( paste0('output/gridEffort/bigGrids/', 'GRID', surveynum, '_', gridsize,'km.rda') )

gridEffort <- doAllGrid(gps = gpsAll,
                     bounds = NULL,
                     dets = pmDetsub,
                     trunc_m = 16215, #as.numeric(summ$ds$width[1]), #METERS
                     dsmodel = detfxn, #NULL or dsm 
                     pixel = 0.225,
                     grid = bigGrid$grid,  ###!! Use the correct big grid !!!
                     plot = F)

# Result of this is a list with the gps you provided, the grid it created, the effort values for each grid, and the detections you provided (these have the effortArea attached to them, as well as the actualArea of that grid cell - actualArea is mostly there as a reference point for sanity check, prob not used for much)

#The centroid function has to be run twice if there are detections with no effort.
#function calculates the difference between start and end times in minutes
centroidAll <- gridCentroid(gridEffort, gpsAll) #assigns the 'pa' column where 0 is no whale, 1 is whale

# centroidAll2<- gridCentroid(gridEffort2, gpsAll)


#Save raw gridEffort and centroid data

saveRDS(gridEffort, here::here(  paste0('output/gridEffort/', gridsize, ' km-', loctype, '/gridEffort', surveynum, '_', gridsize, 'km_', loctype2, '_raw.rda') ))

saveRDS(centroidAll, here::here(paste0('output/centroids/', gridsize, ' km-', loctype), paste0('centroids', surveynum, '_', gridsize, 'km_', loctype2, '_raw.rda')))
```



#GRID QC
All detections should be included and each have effort > 0. Cells with zero effort cannot have centroids calculated. If all is well here, then proceed to centroids.
Otherwise, make changes to any detections in 'Resolve cell...' chunk.
```{r}
#1. check that all detections are included as expected. 
(pmDetsub$ID %in% gridEffort$detections$ID) #Should be all TRUE
data.frame(pmDetsub$ID , gridEffort$detections$ID) #if mismatch, this won't work

#2. which detections have effortArea = 0? (Which are deemed to have 0 effort? No bueno.)
(effdets <- which(gridEffort$detections$effortArea==0)) #output is the index
(effchg <- gridEffort$detections[effdets,]) #output is all info for each detection with 0 effort

#3. check for duplicate detections in the same cells
gridIndex <- as.data.frame(table(gridEffort$detections$gridIndex))
colnames(gridIndex) <- c('effCells', 'frequency')
gridIndex

if (length(effdets) == 0 & nrow(effchg) == 0){
print('All detections are included and have effort')
} else {
  print('Some detections missing effort')
}

if ((nrow(gridIndex) == length(gridEffort$detections$ID)) == TRUE){
  print('Each detection occurs in its own grid cell')
} else {
  print('There are detections that occur in the same grid cell. Which ones?')
  gridTabl <- gridIndex[which(gridIndex[2] > 1),]
  detInd <- as.integer(gridEffort$detections$gridIndex)
  dupInd <- detInd[duplicated(detInd)] #get value of duplicated grid cell(s)
  gridTabl
}


```


#IF CELLS WITH WHALES EXIST WITH NO EFFORT
Do the following steps.
##1. Resolve whale cell(s) with no effort 
```{r}
#Map out the surrounding cells to find ones with effort nearby
gridmap <- attr(gridEffort$grid, 'map')

#Sometimes >1 detections have 0 effort, can look at all of them
# g = 1 # which row in effchg?
# GridInd = effchg[g,]$gridIndex # what grid cell is it in?
GridInd = effchg[,c(3,43)] # pull acid and gridIndex for all detections with no effort
gridEffort$effort$area[GridInd[,2]] # how much effort is in the cell? (Should be zero)

#Get cell index for the closest surrounding cells. Use stack to make one column for them.
GridSurroundtmp <-data.frame(t(gridmap[GridInd[,2],]))

GridSurround <- stack(GridSurroundtmp) #output 2 columns, 1: grid cell, 2: ac. ID
# GridSurround$acid <- GridInd[,1]       #add column for ac ID
# GridSurround <- GridSurround[,-2]      #remove 'ind' col of X1...X4 col names

GridEff <- data.frame(gridEffort$effort$area[GridSurround$values]) # get effort in surr cells
GridDir <- c('S', 'E', 'N', 'W') #make this vector
GridSurrDF <- data.frame(GridSurround, GridEff, GridDir) #combine surr cells and their effort and direction
colnames(GridSurrDF) <- c("gridID", "acid", "effort", "Dir")
GridSurrDF

#Get the centroid locations of the surrounding cells that have effort around the cell for the detection that has no effort.
CentSurround <- filter(centroidAll, effCells %in% GridSurrDF$gridID) #these cells all have effort

#JOIN centroids with grid map cells info and drop NAs (cells without effort)
GridCentJoin <- GridSurrDF %>% full_join(CentSurround, by = c("gridID" = "effCells")) 
GridCentJoin<-GridCentJoin %>% drop_na()
```

##2. Plot grid of whale cell(s) with no effort
REMEMBER: GridCentJoin only includes the surrounding cells, 'Dir', WITH effort, so not all 'Dir' directions may be included in the plot.
CHOOSE the replacement cells using the direction, GridCentJoin$Dir.
```{r}
x <- gridEffort
actualArea <- as.numeric(st_area(x$grid))
coveragePct <- round(x$effort$area / actualArea, 3)
coveragePct <- ifelse(coveragePct > 1, 1, coveragePct)

pdf(here::here('figures', paste0('Grid', surveynum, '_', gridsize, 'km_', loctype, '_CHECKeff2.pdf')), width=11,height=8)

plot(x$grid, col = gray(1 - coveragePct, alpha = .9), lwd = 0.1)
points(x=x$gps$Longitude, y=x$gps$Latitude, col='blue', pch=15, cex = 0.05)
# points(x=pmDetsub$Longitude, y=pmDetsub$Latitude, col='green', pch=1, cex=0.1, lwd=0.1)
# text(pmDetsub$Longitude, pmDetsub$Latitude,col='magenta', labels=pmDetsub$acid, cex=0.1, lwd = 0.1 )
points(centroidAll$lon, centroidAll$lat, col='red', pch = 19, cex=0.1, lwd=0.1)
# points(x=pmDetsub$Longitude, y=pmDetsub$Latitude, col='green', pch=0, cex=0.2, lwd=0.1)
# points(	 178.0128, 22.57317, col='darkorchid', pch=8, cex=0.1, lwd=0.1) #grid locations of 
# # text(	191.1134, 24.82075,col='black', labels=pmDetsGrid1705$acid, cex=0.1, lwd = 0.1 )
# points(	192.0295, 21.76568, col='darkorange', pch=8, cex=0.1, lwd=0.1) #grid locations of 
# text(	192.0295, 	21.76568, col='black', labels=pmDetsGrid1705$acid, cex=0.1, lwd = 0.1 )
# points(192.7076, 24.31634, col='magenta', pch = 0, cex=0.2, lwd=0.1) #where is this centroid compared to the actual location? for 1303 sighting to change depth to -697
# points(192.7200, 24.28700, col='goldenrod', pch = 0, cex=0.2, lwd=0.1) #where is this centroid compared to the actual location?

#plot detections and their centroids missing effort
text(x=effchg$Longitude, y=effchg$Latitude	, col='magenta', labels=effchg$sid, cex=0.1, lwd = 0.1 )
points(x=effchg$Longitude, y=effchg$Latitude	, col='orange', pch = 19, cex=0.2, lwd = 0.07 )
# text(GridCentJoin$lon, GridCentJoin$lat, col='darkviolet', labels=GridCentJoin$Dir, cex=0.1, lwd=0.1)

#plot detections missing in the centroid data
# text(x=CentMiss$Longitude, y=CentMiss$Latitude	, col='magenta', labels=CentMiss$acid, cex=0.1, lwd = 0.1 )
#  points(x=CentMiss$Longitude, y=CentMiss$Latitude	, col='orange', pch = 19, cex=0.07, lwd = 0.07 )

dev.off()


```


##3. Replace whale cells with no effort
```{r}
#IF MORE THAN ONE PRESENCE HAS NO EFFORT, DO EACH ONE SEPARATELY, WAAHH
#change values for gridIndex, effortArea, and actualArea in 'gridEffort$detections' while keeping everything else the same
CentNew = 1 #c(1,4)  #represents the row number(s) in GridCentJoin that corresponds with the correct dir, SO S=1, E=2, N=3, W=4
#!! 'newCell' will change depending on the survey and spatial scale #
newCell <- GridCentJoin[CentNew,1] # Select the gridIndex for the new cells based on the plot

newEffArea <- gridEffort$effort$area[newCell] # get the effort for each new cell from original grid
newActArea <- actualArea[newCell] # get the new cell's actual area

#This replaces the old cell without effort with info for the new cell in these 3 cols
effchg2 <- effchg[3,] %>% mutate(gridIndex = newCell, effortArea=newEffArea, actualArea=newActArea)

#remove old effchg and replace with new cell info in the detections list
gridEffort$detections <- gridEffort$detections[-effdets,] #effdets is the index from grid detections
gridEffort$detections <- rbind(gridEffort$detections, effchg2) #bind new cell info with rest of grid detections

#original rows from 'effdets' index should be missing with the new cells at the bottom 
 View(gridEffort$detections)
 
```

#4. Save new grid data
!!! Check that all whale cells have Effort > 0
```{r}
saveRDS(gridEffort, here::here(paste0('output/gridEffort/', gridsize, ' km-', loctype), paste0('gridEffort', surveynum, '_', gridsize, 'km_', loctype2, '.rda')))
```

#5. Run centroids again
```{r}
centroidAll <- gridCentroid(gridEffort, gpsAll)
```


#Verify Centroids
##1. Check that all centroids with presences have effort as expected.

At this point, all presences AND absences should have effort > 0.
Presences can be filtered using detections$gridIndex 
```{r}

#do another check to make sure all detections have centroid locations WITH EFFORT
(all_prz <- dplyr::filter(centroidAll, pa>0))  

print('Do the number of presences in the centroids match the number of encounters?')

if((nrow(all_prz) == nrow(pmDetsub))==TRUE){
  print('Hooray! Each enc is located in its own cell')
} else {
  print('Oopsy! May have >1 enc per cell. Lets sort it out, itll be fun, I swear...' )
  
  centroidAlltmp <- filter(centroidAll, pa != 1)     #remove the rows with presences
  
  dupRow <- subset(all_prz, effCells %in% dupInd)      #subset the duplicated rows, use == if %in% doesn't work
  all_prz = all_prz[!(all_prz$effCells %in% dupInd), ] #remove duplicated rows based on the effCell value

  for (dup in 1:length(dupInd)){
    dupRow$pa[dup] <- dupRow$pa[dup] + (gridTabl[dup,2]-1) #For each duplicate, take the (total enc-1) and add it to the pa column
  }
  print(dupRow)                                      #check that the pa column is correct
all_prz = rbind(all_prz, dupRow)                     #add the new row to the rest of the presences
centroidAll <- rbind(centroidAlltmp, all_prz)       #Should equal the same rows as before since just changing the value in pa

} ## We want TRUE, but FALSE could mean 2 detections in one cell
# dupRow2 <- dupRow
# colnames(dupRow2)[3] <- 'UTC'
# dupRow2 <- dupRow2[ ,-4]
# dupRow2$survey <- surveynum
#   saveRDS(dupRow2, here::here(paste0('output/centroids/', gridsize, ' km-', loctype), paste0('centroids', surveynum, '_', gridsize, 'km_', loctype2, 'dup.rda')))

```

##2. Do if centroids with presence is missing effort.

```{r}
### IF DETECTIONS MISSING IN CENTROIDS ###
#Are there any missing detections or mismatches in detections between the grid data and the centroid data? These should all be 0
nrow(all_prz)
nrow(pmDetsub)
allPrz <- all_prz$effCells
allPrz[!( allPrz %in% gridDets)]

table(gridDets)
table(allPrz)

(notINgrid <- (all_prz$effCells[!all_prz$effCells %in% gridEffort$detections$gridIndex])) 
(notINcent <- (gridEffort$detections$gridIndex[!gridEffort$detections$gridIndex %in% all_prz$effCells]))         




#which detection is missing in the centroid data?
CentInd = which(gridEffort$detections$gridIndex == notINcent) #outputs index
CentMiss = gridEffort$detections[CentInd,]
# Do more if things don't match, but they should and all should be well.
# (id_not <- which(!(gridIndex$Var1 %in% all_prz$effCells == TRUE))) #find the detections that don't match up
# (not_det <- gridEffort$detections[id_not,]) #this is hopefully empty. These are the detections NOT IN the centroids

```


##3. More Checks
Some cells with presences were passed more than once. These need to have the correct times associated with them to retrieve the correct environmental data. This chunk finds the detections that occur > 1, picks the time that is within a day of the detection UTC, and assigns it to that detection.
THEN, it takes the other cells that don't match the detection UTC and assign it to pa = 0, so it's still counted but not as a presence, pa = 1. 
The same process is done for the absences, where any effort spanning more than 2-3 days are identified and included as separate 
```{r}
#### PRESENCE DATA ####
# `dbl_prz` and `dbl_det` should have the same number of rows
#Need to know which time (start or end) is associated with a detection if pa = 1 and diff_mins > 1440 (mins in a day)
dbl_prz <- dplyr::filter(centroidAll, abs(diff_mins) > 1440 & pa > 0)

#In cells with PRESENCE, which detections occur in cells that were passed > 1 time
dbl_idx <- which(gridEffort$detections$gridIndex %in% dbl_prz$effCells == TRUE)
dbl_det <- gridEffort$detections[dbl_idx,]
dbl_det <- dbl_det[order(dbl_det$gridIndex),] #detections in order of gridIndex to match dbl_prz

if ((nrow(dbl_prz) != nrow(dbl_det))==TRUE){ #should be true. IF IT'S NOT, PROBABLY DUE TO >1 whale in the same cell
#remove duplicated gridIndexes
dbl_det = dbl_det[!duplicated(dbl_det[ , c("gridIndex")]),]
nrow(dbl_prz) == nrow(dbl_det)
} else {
  print('Same nrow')
}
## For some cells with pa = 1, the start and end times are not within a day of the detection time for the same location (cell centroid). The detection times are compared to the start and end times. Those times that differ by less than a day are used. 
idt1 = which(abs(difftime(dbl_det$UTC, dbl_prz$start, units = "mins")) < 1440) #find rows where time difference btwn detection and start is less than a day
idt2 = which(abs(difftime(dbl_det$UTC, dbl_prz$end, units = "mins")) < 1440)   #find rows where time difference btwn detection and end is less than a day
idt3 = which(abs(difftime(dbl_det$UTC, dbl_prz$start, units = 'mins')) > 1440 & abs(difftime(dbl_det$UTC, dbl_prz$end, units = 'mins')) > 1440) #find rows where time difference btwn detection and both start and end is more than a day

gr8end <- dplyr::select(dbl_prz[idt1,], `UTC`= end, effCells, EffArea, diff_mins, lon, lat, pa) #df for end times that occur on diff day than detection
gr8start <- dplyr::select(dbl_prz[idt2,], `UTC`= start, effCells, EffArea, diff_mins, lon, lat, pa) #df for start times that occur on diff day than detection

# Take start and end times when both occur on different day than the detection and create a row for them in a df
gr8both1 <- dplyr::select(dbl_prz[idt3,], `UTC`= end, effCells, EffArea, diff_mins, lon, lat, pa) 
gr8both2 <- dplyr::select(dbl_prz[idt3,], `UTC`= start, effCells, EffArea, diff_mins, lon, lat, pa)
GR8all <- rbind(gr8end, gr8start, gr8both1, gr8both2)
GR8all$pa <- 0 #these are not presences because they are the times that occur on a different day than the detection

# take the actual detection times and match them with the centroid lat lon. 
dbl_det1 <- data.frame(dplyr::select(dbl_det[idt1,], UTC), dplyr::select(dbl_prz[idt1,], effCells, EffArea, diff_mins, lon, lat, pa))
dbl_det2 <- data.frame(dplyr::select(dbl_det[idt2,], UTC), dplyr::select(dbl_prz[idt2,], effCells, EffArea, diff_mins, lon, lat, pa))
dbl_det3 <- data.frame(dplyr::select(dbl_det[idt3,], UTC), dplyr::select(dbl_prz[idt3,], effCells, EffArea, diff_mins, lon, lat, pa))
DBL_prz <- rbind(dbl_det1, dbl_det2, dbl_det3)   ## create df of centroids for all presences with >1 time

#### ABSENCE DATA ####
#In cells with ABSENCE, manipulate to have both times in one column
dbl_abz <- filter(centroidAll, abs(diff_mins) > 3200 & pa == 0) #double effort within 2-3 days? Threshold based on grid data
dbl_start <- dplyr::select(dbl_abz, `UTC`=start)#first time
dbl_end   <- dplyr::select(dbl_abz, `UTC`=end)  #second time
dbl_else <- dplyr::select(dbl_abz, -start, -end)#the rest of the data
dbl_startdf <- data.frame(dbl_start, dbl_else)   #combine first time with the rest of the data
dbl_enddf <- data.frame(dbl_end, dbl_else)       #combine second time with the rest of the data
DBL_abz <- rbind(dbl_startdf, dbl_enddf, GR8all) ## create df of centroids for all absences that had double effort


#Configure cells with single effort to have same format as double effort to combine them
sgl_abz <- dplyr::filter(centroidAll, abs(diff_mins) <= 3200 & pa == 0) # all times are within 2-3 days, which is fine bc datasets are monthly
SGL_abz <- dplyr::select(sgl_abz, `UTC` = start, everything(), -end)

sgl_prz <- dplyr::filter(centroidAll, abs(diff_mins) <= 1440 & pa > 0)
SGL_prz <- dplyr::select(sgl_prz, `UTC` = start, everything(), -end)


####COMBINE ALL!!! ####

ALLcentroids <- rbind(DBL_prz, SGL_prz, DBL_abz, SGL_abz)
ALLcentroids$survey <- surveynum
```
##CHECK FOR NA
```{r}

sum(is.na(ALLcentroids))     # Total NA for presences
(colsum <- colSums(is.na(ALLcentroids))) # Total NA per column
ALLcentroids<- ALLcentroids[complete.cases(ALLcentroids), ]
```
##Save
```{r}
nrow(pmDetsub) #number of detections
nrow(dplyr::filter(ALLcentroids, pa > 0)) #should equal number of detections
print('Do the number of encounters match the number of centroids with encounters?')
if (nrow(pmDetsub) == nrow(dplyr::filter(ALLcentroids, pa == 1))){
  print('Yes, all encounters match to a centroid. YOU ARE A GENIUS. Now save that shit.')
 saveRDS(ALLcentroids, here::here(paste0('output/centroids/', gridsize, ' km-', loctype), paste0('centroids', surveynum, '_', gridsize, 'km_', loctype2, '.rda')))
  print('Shit is saved')
} else {
  print('No, but are there > 1 encs in a cell? Check that and it should be all good.')   
  saveRDS(ALLcentroids, here::here(paste0('output/centroids/', gridsize, ' km-', loctype), paste0('centroids', surveynum, '_', gridsize, 'km_', loctype2, '.rda')))
  print('Shit is saved')
}
```


#Plot for Sanity Check
```{r}
# Theres also a plotty function, but it will look pretty messy for all your data. Trackline is (roughly) in blue,
# detections are red dots, amount of effort is shaded in gray, but for a big grid you wont really be able to see the shading
# plotGridResult(gridEffort)

#non-function form of plotting to plot one detection or other data points
x <- gridEffort
actualArea <- as.numeric(st_area(x$grid))
coveragePct <- round(x$effort$area / actualArea, 3)
coveragePct <- ifelse(coveragePct > 1, 1, coveragePct)

pdf(here::here('figures', paste0('Grid', surveynum, '_', gridsize, 'km_', loctype, '_COMB.pdf')), width=11,height=8)

plot(x$grid, col = gray(1 - coveragePct, alpha = .9), lwd = 0.2)
points(x=x$gps$Longitude, y=x$gps$Latitude, col='blue', pch=16, cex = 0.1)
points(centroidAll$lon, centroidAll$lat, col='red', pch = 19, cex=0.1, lwd=0.1)

# points(x=centroidAll$lon, y=centroidAll$lat, col='red', pch = 16, cex=0.5 ) #point for A121.S44
#203.4318	21.36744	three times through here?

# points(x=dbl_eff2$lon, y=dbl_eff2$lat, col='red', pch = 0, cex=0.2, lwd = 0.2 ) 
points(x=pmDetsub$Longitude, y=pmDetsub$Latitude, col='green', pch=1, cex=0.3, lwd=0.3) #Longitude is trackline location
text(pmDetsub$Longitude, pmDetsub$Latitude,col='magenta', labels=pmDetsub$acid, cex=0.3, lwd = 0.3 )
# points(x=187.6831, y=26.60179	, col='darkturquoise', pch = 15, cex=0.5, lwd = 0.5 ) #square
# points(x=187.5513, y=26.66397	, col='deeppink3', pch = 17, cex=0.5, lwd = 0.5 ) #triangle
# points(effchg$Longitude, effchg$Latitude, col='orange', pch=3, cex=0.05)

dev.off()
# points(effchg$lon2, effchg$lat, col='red', pch = 0, cex=0.2, lwd = 0.2 )
# points(all_prz$lon, all_prz$lat,col='goldenrod', pch = 16, cex=0.2, lwd = 0.2 )


# plot(pmDetsub$lon, pmDetsub$Latitude,col='magenta', pch = 17, cex=0.2, lwd = 0.2 )


# points(201.8474, 22.30299,col='green', pch = 16, cex=0.2, lwd = 0.2 )
# points(202.0274, 22.30299,col='darkmagenta', pch = 16, cex=0.2, lwd = 0.2 )
# points(201.7574, 22.30299,col='deepskyblue', pch = 16, cex=0.2, lwd = 0.2 )
# points(201.9824, 22.34799,col='darkorange', pch = 0, cex=0.2, lwd = 0.2 ) #15053 centroid ->CHOOSE THIS ONE
# points(202.0724, 22.34799,col='darkmagenta', pch = 0, cex=0.2, lwd = 0.2 )#15054 centroid

# points(effchg$Longitude, effchg$Latitude, col='darkred', pch = 16, cex=0.2, lwd = 0.2 ) #dot, 1303_A20
# points(effchg$lon_trk2, effchg$lat_trk, col='darkorange2', pch = 16, cex=0.2, lwd = 0.2 ) #trackline location
# points(surroundCells$lon[1], surroundCells$lat[1], col='aquamarine2', pch = 16, cex=0.2, lwd = 0.2 )
# points(surroundCells$lon[2], surroundCells$lat[2], col='coral2', pch = 16, cex=0.2, lwd = 0.2 )
# points(surroundCells$lon[3], surroundCells$lat[3], col='darkorchid2', pch = 16, cex=0.2, lwd = 0.2 )
# 
# 
# 
# #From straightpath_yb
# ggplot(str8_1303, aes(x=Longitude, y=Latitude)) +
#             geom_point(aes(group=timeGroup, col=straight), size = 1.3) +
#             scale_color_manual(values = c('red', 'darkgreen')) +
#             ggtitle(paste0('nSmall=', nSmall, ", nLarge=", nLarge)) +
#   geom_point(x=-167.28, y=24.287, col='magenta')

```

## Centroid Consolidate!
```{r}
CombCents = NULL
# AcOnlyEnv = NULL
# pa1Check = NULL
for (s in c(1641, 1303, 1604, 1705, 1706)){
  combi <- readRDS(here::here( paste0('output/centroids/', gridsize, ' km-', loctype, '/',  #new AcOnly centroids
                                        'centroids', s, '_', gridsize, 'km_', loctype2, '.rda') ))

  CombCents = rbind(CombCents, combi)  
  
}
saveRDS(CombCents, here::here( paste0('output/centroids/', gridsize, ' km-', loctype, '/',  #new Combined centroids
                                        'centroids_AllSurveys', '_', gridsize, 'km_', loctype2, '.rda') ))

```

