---
title: "Environmental Data Inspection"
author: "Yvonne Barkley"
date: "7/27/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r}
library(tidyverse)

#to remove variables, replace **
rm(list = ls()[grep("env", ls())])
```



# Check environmental data

#LOAD DATA
```{r}
surveynum = 1641
gridsize = 25
loctype = 'allsw'
loctype2 = 'all'

#to remove variables, replace **
# rm(list = ls()[grep("diff", ls())])

# Get gridEffort for the correct survey and spatial scale. 
gridEffortCleanup <- readRDS(paste0('output/gridEffort/', gridsize, ' km-', loctype, '/', 'gridEffort', surveynum, '_', gridsize, 'km_', loctype2, '.rda'))
pmDetsGrid <- gridEffortCleanup$detections
pmDetsGrid <- pmDetsGrid[, c(1:26,43,45)]

```


#consolidate grid detections
```{r}


pmDetsGrid1641 = pmDetsGrid
pmDetsGrid1642 = pmDetsGrid
pmDetsGrid1303 = pmDetsGrid
pmDetsGrid1604 = pmDetsGrid
pmDetsGrid1705 = pmDetsGrid
pmDetsGrid1706 = pmDetsGrid


pmDetsGridTot = rbind.data.frame(pmDetsGrid1641,pmDetsGrid1642, pmDetsGrid1303, pmDetsGrid1604, pmDetsGrid1705, pmDetsGrid1706)

```



```{r}
env1_sst      <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_sst.rda')))
env2_chla     <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_chla.rda')))
env3_godas600 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_temp600.rda')))
env4_ww3      <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_wavepow.rda')))
# env4_ww3      <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_wp.rda')))
env5_bath     <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_bathy.rda')))

# env5_bath     <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_bathyAll700.rda')))

env67_slpasp  <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_slpasp.rda')))
env8_d2land   <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_d2land.rda')))
env9_ssh      <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_ssh.rda')))
env10_sshsd   <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_sshsd.rda')))
env11_eke     <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_eke.rda')))
env12_d2smt   <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_dist2smt.rda')))
# env12_d2smt   <- read.csv(here::here(paste0('output/data_output/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_dist2smt.csv')))

#check out all depths
# env5_bathAll     <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0(surveynum, '_', gridsize, 'km_', loctype2, '_bathyAll.rda')))

#1303 rearrange and resave env5_bath data as tibble df
# env5_bath <-  dplyr::select(env5_bath, last_col(), everything()) #and saved
```


Change the column names to UTC
```{r}
colnames(env3_godas600)[colnames(env3_godas600) == "date"] <- "UTC"
colnames(env4_ww3)[colnames(env4_ww3) == "date"] <- "UTC"
colnames(env9_ssh)[colnames(env9_ssh) == "date"] <- "UTC"
colnames(env10_sshsd)[colnames(env10_sshsd) == "date"] <- "UTC"
colnames(env11_eke)[colnames(env11_eke) == "date"] <- "UTC"

```


Fill NAs to see where they are
```{r}
matchedData[matchedData[,8]=='NA'] <- NA                  #put in NaNs to variable column

env1_sst[env1_sst[,8]=='NA'] <- NA      
env2_chla[env2_chla[,8]=='NA'] <- NA       
env3_godas600[env3_godas600[,5]=='NA'] <- NA  
env4_ww3[env4_ww3[,9]=='NA'] <- NA        
env5_bath[env5_bath[,1]=='NA'] <- NA
env67_slpasp[env67_slpasp[,3]=='NA'] <- NA    
env67_slpasp[env67_slpasp[,4]=='NA'] <- NA    
env8_d2land[env8_d2land[,1]=='NA'] <- NA     
env9_ssh$ssh[env9_ssh[,4]=='NA'] <- NA     
env10_sshsd$sshsd[env10_sshsd[,4]=='NA'] <- NA     
env11_eke$eke[env11_eke[,4]=='NA'] <- NA       
env12_d2smt[env12_d2smt[,2]=='NA'] <- NA 


env11_eke$eke[is.na(env11_eke[, 4]),]


```


#Bind data
Bind all data together with the presence/absence information
```{r}
#presence/absence info comes from env1_sst

#TO make sure that the combining the dataframes is correct, sort each one based on effCells. Use arrange or sort, NOT order (f's things up)
# Don't use joins because the duplicate effCells also f things up.

env1_sstSORT  <- arrange(env1_sst, effCells)
env2_chlaSORT <- arrange(env2_chla, effCells)
env3_godasSORT<- arrange(env3_godas600, effCells)
env4_ww3SORT  <- arrange(env4_ww3, effCells)
env5_bathSORT <- arrange(env5_bath, effCells)
env67_slpaspSORT<- arrange(env67_slpasp, effCells)
env8_d2landSORT<- arrange(env8_d2land, effCells)
# env8_d2landSORT2<- arrange(env8_d2land, effCells)
env9_sshSORT  <- arrange(env9_ssh, effCells)
env10_sshsdSORT<- arrange(env10_sshsd, effCells)
env11_ekeSORT <- arrange(env11_eke, effCells)
env12_d2smtSORT<- arrange(env12_d2smt, effCells)

PmEnvTMP = data.frame(env1_sstSORT[, c(1:3, 5:7,9)], `chla`=env2_chlaSORT[, 9], `temp600`=env3_godasSORT[,5], `wavepow`=env4_ww3SORT[,9], `bath_m`= env5_bathSORT$bath_m, `slope`=env67_slpaspSORT[,3], `aspect`=env67_slpaspSORT[,4], 'd2land'=env8_d2landSORT[,1], `ssh`=env9_sshSORT[,4], `sshsd`=env10_sshsdSORT[,4], `eke`=env11_ekeSORT[,4], `d2smt`=env12_d2smtSORT[,1])
colnames(PmEnvTMP)[colnames(PmEnvTMP) == "sst_mean"] <- "sst"


# PmEnvDum = data.frame(env1_sstSORT[, c(1:3, 5:7,9)],env8_d2landSORT[,1], env5_bathSORT[,1])

# PmEnvTMP2 = data.frame(env1_sstSORT[, c(1:3, 5:7,9)], env2_chlaSORT[, c(2,9)], env3_godasSORT[,c(5,8)], env4_ww3SORT[,c(8,9)], env5_bathSORT[,c(1,6)], env67_slpaspSORT[,c(3,7)], env67_slpaspSORT[,c(4,7)], env8_d2landSORT[,c(1,6)], env9_sshSORT[,c(4,7)], env10_sshsdSORT[,c(4,7)], env11_ekeSORT[,c(4,7)], env12_d2smtSORT[,c(2,5)])
# 
# #try again
# PmEnvTMP3=list(env1_sst[, c(2,9)], env2_chla[, c(2,9)], env3_godas600[,c(5,8)], env4_ww3[,c(8,9)], env5_bath[,c(1,6)], env67_slpasp[,c(3,7)], env67_slpasp[,c(4,7)], env8_d2land[,c(1,6)], env9_ssh[,c(4,7)], env10_sshsd[,c(4,7)], env11_eke[,c(4,7)], env12_d2smt[,c(2,5)]) %>% reduce(inner_join, by = 'effCells')
# list for whatevers
# env1_sst
# env2_chla
# env3_godas600
# env4_ww3
# env5_bath
# env67_slpasp
# env8_d2land
# env9_ssh
# env10_sshsd
# env11_eke
# env12_d2smt 
```


#Remove NAs from total Env Data
Check for NAs with pa>0 and pa=0
```{r}

#find any NAs separately in presences or absences before removing them
pa1 <- filter(PmEnvTMP, pa > 0)  
nrow(pa1) == nrow(pmDetsGrid)  ## !!!!!! Check number of presences: Do they match btwn centroid data and grid data?
pa0 <- filter(PmEnvTMP, pa == 0)

# plot(pa1$dist)
which(pa1$dist>0)  #any positive depths I need to murder?


sum(is.na(pa1))     # Total NA for presences
(colsum <- colSums(is.na(pa1))) # Total NA per column


```

FILL IN MISSING DATA FOR PRESENCES
```{r}
####### Replace NAs ######
#If NA's exist in the env data for presences, sort that out here:
#This is specifically for SSH, SSHsd, EKE

if(colsum > 0){
pa1NAssh <- pa1[is.na(pa1$ssh) > 0, ]  #find NAs based on columns in colsum
pa1NAsshsd <- pa1[is.na(pa1$sshsd) > 0, ]  #find NAs based on columns in colsum
pa1NAeke <- pa1[is.na(pa1$eke) > 0, ]  #find NAs based on columns in colsum

# pa1NAb <- pa1NA[,1:6] #used for 2017...?

# Since the resolution of the ssh data is ~8km, increasing the search out to 8 km ~ 0.055
#REPLACE MISSING SSH VALUES
for (r in 1:nrow(pa1NAssh)) {
  
  lonBad <- pa1NAssh[r, 4] #pull out longitude
  lonPlus <- (lonBad + 0.055) +360
  lonMinus <- (lonBad - 0.055) +360
  
  latBad <- pa1NAssh[r, 5] #pull out latitude
  latPlus <- (latBad + 0.055) 
  latMinus <- (latBad - 0.055)
  
  #create new vector with cell values around original cell
  UpLft <- c(lonPlus,  latPlus)
  UpRt  <- c(lonMinus, latPlus)
  LoLft <- c(lonPlus,  latMinus)
  LoRt  <- c(lonMinus, latMinus)
  
  #create dataframe with new buffered positions with UTC of original position
  newPosdf <- as.data.frame(rbind(UpLft, UpRt, LoLft, LoRt))
  newPosdf$UTC <- pa1NAssh[r,1]
  colnames(newPosdf) <- c('Longitude', 'Latitude', 'UTC')
  
  
  env9_sshbuff = NULL
for (i in 1:nrow(newPosdf)){
 
  bufftmp <- which(abs(sshtidya$lon-newPosdf$Longitude[i]) == min(abs(sshtidya$lon-newPosdf$Longitude[i])) & 
                    abs(sshtidya$lat-newPosdf$Latitude[i]) == min(abs(sshtidya$lat-newPosdf$Latitude[i])) &
                    abs(sshtidya$date-newPosdf$UTC[i]) == min(abs(sshtidya$date-newPosdf$UTC[i])))
  sshbuff <- sshtidya[bufftmp[1],] #some had duplicates, so take the first one
  sshbuff$truelon <- newPosdf$Longitude[i]
  sshbuff$truelat <- newPosdf$Latitude[i]
  
  
  env9_sshbuff <- rbind(env9_sshbuff, sshbuff)
}

env9_sshbuff$effCells <- pa1NAssh$effCells[r] #add column with effCell value for the original location

sshbuffMean <- mean(env9_sshbuff$ssh) #take average of new buffered ssh
#replace missing ssh value for the rth row in pa1NAssh
pa1NAssh[r, 15] <- mean(env9_sshbuff$ssh)
}


#REPLACE MISSING SSHSD VALUES
for (r in 1:nrow(pa1NAsshsd)) {
  
  lonBad <- pa1NAsshsd[r, 4] #pull out longitude
  lonPlus <- (lonBad + 0.055) +360
  lonMinus <- (lonBad - 0.055) +360
  
  latBad <- pa1NAsshsd[r, 5] #pull out latitude
  latPlus <- (latBad + 0.055) 
  latMinus <- (latBad - 0.055)
  
  #create new vector with cell values around original cell
  UpLft <- c(lonPlus,  latPlus)
  UpRt  <- c(lonMinus, latPlus)
  LoLft <- c(lonPlus,  latMinus)
  LoRt  <- c(lonMinus, latMinus)
  
  #create dataframe with new buffered positions with UTC of original position
  newPosdf <- as.data.frame(rbind(UpLft, UpRt, LoLft, LoRt))
  newPosdf$UTC <- pa1NAsshsd[r,1]
  colnames(newPosdf) <- c('Longitude', 'Latitude', 'UTC')
  
  
  env9_sshZbuff = NULL
for (i in 1:nrow(newPosdf)){
 
  bufftmp <- which(abs(sshZtidya$lon-newPosdf$Longitude[i]) == min(abs(sshZtidya$lon-newPosdf$Longitude[i])) & 
                    abs(sshZtidya$lat-newPosdf$Latitude[i]) == min(abs(sshZtidya$lat-newPosdf$Latitude[i])) &
                    abs(sshZtidya$date-newPosdf$UTC[i]) == min(abs(sshZtidya$date-newPosdf$UTC[i])))
  sshsdbuff <- sshZtidya[bufftmp[1],] #some had duplicates, so take the first one
  sshsdbuff$truelon <- newPosdf$Longitude[i]
  sshsdbuff$truelat <- newPosdf$Latitude[i]
  
  
  env9_sshZbuff <- rbind(env9_sshZbuff, sshsdbuff)
}

env9_sshZbuff$effCells <- pa1NAsshsd$effCells[r] #add column with effCell value for the original location

#replace missing sshsd value with mean for the rth row in pa1NAsshsd
pa1NAsshsd[r, 16] <- mean(env9_sshZbuff$sshsd)
}

#REPLACE MISSING EKE VALUES
for (r in 1:nrow(pa1NAeke)) {
  
  lonBad <- pa1NAeke[r, 4] #pull out longitude
  lonPlus <- (lonBad + 0.055) +360
  lonMinus <- (lonBad - 0.055) +360
  
  latBad <- pa1NAeke[r, 5] #pull out latitude
  latPlus <- (latBad + 0.055) 
  latMinus <- (latBad - 0.055)
  
  #create new vector with cell values around original cell
  UpLft <- c(lonPlus,  latPlus)
  UpRt  <- c(lonMinus, latPlus)
  LoLft <- c(lonPlus,  latMinus)
  LoRt  <- c(lonMinus, latMinus)
  
  #create dataframe with new buffered positions with UTC of original position
  newPosdf <- as.data.frame(rbind(UpLft, UpRt, LoLft, LoRt))
  newPosdf$UTC <- pa1NAeke[r,1]
  colnames(newPosdf) <- c('Longitude', 'Latitude', 'UTC')
  
  
  env9_ekebuff = NULL
for (i in 1:nrow(newPosdf)){
 
  bufftmp <- which(abs(ekeZtidya$lon-newPosdf$Longitude[i]) == min(abs(ekeZtidya$lon-newPosdf$Longitude[i])) & 
                    abs(ekeZtidya$lat-newPosdf$Latitude[i]) == min(abs(ekeZtidya$lat-newPosdf$Latitude[i])) &
                    abs(ekeZtidya$date-newPosdf$UTC[i]) == min(abs(ekeZtidya$date-newPosdf$UTC[i])))
  ekebuff <- ekeZtidya[bufftmp[1],] #some had duplicates, so take the first one
  ekebuff$truelon <- newPosdf$Longitude[i]
  ekebuff$truelat <- newPosdf$Latitude[i]
  
  
  env9_ekebuff <- rbind(env9_ekebuff, ekebuff)
}

env9_ekebuff$effCells <- pa1NAeke$effCells[r] #add column with effCell value for the original location

#replace missing eke value with the mean for the rth row in pa1NAeke
pa1NAeke[r, 17] <- mean(env9_ekebuff$eke)
}


}




```


#Merge data
Get detection numbers and click info for presences. Everything was saved in the '$detections' part of the list for gridEffort. 
Match up the click info with PmEnv using the gridIndex and effCells (named effCells in 8_GetEnvData.Rmd)
```{r}
###
sum(is.na(pa0))     # Total NA for absences
colSums(is.na(pa0)) # Total NA per column


PmEnv <- PmEnvTMP %>% drop_na()
PmEnvAbs <- filter(PmEnv, pa == 0)
PmEnvPres <- filter(PmEnv, pa > 0)
# PmEnvPres <- rename(PmEnvPres, Longitude = lon, Latitude=lat)
# PmEnvAbs <- rename(PmEnvAbs, Longitude = lon, Latitude=lat)


colSums(is.na(PmEnv)) #CHECK NAs AGAIN, champ
```

#DATA INSPECTION PLOTS
```{r}

dataSet = PmEnv

pdf(here::here('figures', paste0('DataInspc1_', surveynum, '_', gridsize, 'km_', loctype, '_raw.pdf')), width=11,height=8)


par(mfrow = c(3, 4), mar=c(3,4,2,1), oma=c(0,0,3,1))
 
loopVec <- 7:18  #columns from PmScaled to plot
 
 for (j in loopVec){
   
   datPlot <- dataSet[,c(6,j)]
   
   plot(datPlot[,2], main = colnames(datPlot)[2], xlab = 'index', ylab='')
      # plot(datPlot[,2], datPlot[,1], ylab = 'Whales', xlab = colnames(datPlot)[2])
mtext(paste0("Survey: ", surveynum,', ', loctype,' data' ), side=3, line=1, outer=TRUE, cex=1, font=1)

 }

dev.off()


pdf(here::here('figures', paste0('DataInspc2_', surveynum, '_', gridsize, 'km_', loctype, '_raw.pdf')), width=11,height=8)

par(mfrow = c(3, 4), mar=c(3,3,2,1), oma=c(0,0,3,1))
 
loopVec <- 7:18  #columns from PmScaled to plot
 
 for (j in loopVec){
   
   datPlot <- dataSet[,c(6,j)]
   
   plot(datPlot[,2], datPlot[,1], main = colnames(datPlot)[2],  ylab='', xlab = '')
      # plot(datPlot[,2], datPlot[,1], ylab = 'Whales', xlab = colnames(datPlot)[2])
mtext(paste0("Survey: ", surveynum,', ', loctype,' data'), side=3, line=1, outer=TRUE, cex=1, font=1)

}

dev.off()



dataSet = pa1

pdf(here::here('figures', paste0('DataInspc3_', surveynum, '_', gridsize, 'km_', loctype, '_raw.pdf')), width=11,height=8)

par(mfrow = c(3, 4), mar=c(3,3,2,1), oma=c(0,0,3,1))
 
loopVec <- 7:18  #columns from PmScaled to plot
 
 for (j in loopVec){
   
   datPlot <- dataSet[,c(6,j)]
   
   plot(datPlot[,2], main = colnames(datPlot)[2], xlab = 'index', ylab='')
      # plot(datPlot[,2], datPlot[,1], ylab = 'Whales', xlab = colnames(datPlot)[2])
mtext(paste0("Survey: ",surveynum,', ', loctype,' data - Positive Whale'), side=3, line=1, outer=TRUE, cex=1, font=1)

 }

dev.off()

```

#MERGE & SAVE DATA
Combine data from centroids with aux data from grid detections by cell number
```{r}
# connect the centroids for Pm detections and env data with click info from pmDetsGrid
merge_tmp <- merge(PmEnvPres, pmDetsGrid, by.x = 'effCells', by.y = 'gridIndex')

merge_tmp2 <- dplyr::select(merge_tmp, c(6,2,23,19:22,1,3,45,4,5,43,27,28,36,35,44,29, everything())) #This contains the presences with the env data from the centroid locations

# merge_tmp2 <- dplyr::select(merge_tmp, c(2,23,19:22,1,3,45,4,5,28,27,everything())) #This contains the presences with the env data from the centroid locations
# spruce this up some more to bind better with the absences

merge_tmp3 <- merge_tmp2 %>% rename(UTC=UTC.x, UTCp = UTC.y, Longitude=Longitude.x, Latitude = Latitude.x, Latitudep = Latitude.y, Longitudep = Longitude.y)

#merge absence data with acoustic encounter data to include all envData with the aux acoustic information
trythis <- bind_rows(merge_tmp3, PmEnvAbs)
# trythis$survey = surveynum

(dataname <- paste0(surveynum, '_', gridsize, 'km_', loctype2, '_complete'))
assign(dataname, trythis)

#save completed file
saveRDS(assign(dataname, trythis), here::here(paste0('output/envData/', gridsize, ' km-', loctype), dataname))


```


#Complete Data
```{r}
Complete1641 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1641_', gridsize, 'km_', loctype2, '_complete')))
Complete1641$survey = 1641
Complete1641$year = 2010

Complete1642 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1642_', gridsize, 'km_', loctype2, '_complete')))
Complete1642$survey = 1642
Complete1642$year = 2010

Complete1303 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1303_', gridsize, 'km_', loctype2, '_complete')))
Complete1303$survey = 1303
Complete1303$year = 2013

Complete1604 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1604_', gridsize, 'km_', loctype2, '_complete')))
Complete1604$survey = 1604
Complete1604$year = 2016

Complete1705 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1705_', gridsize, 'km_', loctype2, '_complete')))
Complete1705$survey = 1705
Complete1705$year = 2017

Complete1706 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1706_', gridsize, 'km_', loctype2, '_complete')))
Complete1706$survey = 1706
Complete1706$year = 2017

#Added in presence row with pa=2 that got dropped somewhere...
# combine1705 <- bind_rows(Complete1705b, Complete1705)
# saveRDS(assign(dataname, combine1705), here::here(paste0('output/envData/', gridsize, ' km-', loctype), dataname))

CompleteTotal <- bind_rows(Complete1641, Complete1642, Complete1303, Complete1604, Complete1705, Complete1706)   

#convert Longitude of centroids to 0-360
CompleteTotal$Longitude <- ifelse(CompleteTotal$Longitude<0, CompleteTotal$Longitude+360, CompleteTotal$Longitude)
CompleteTotal$lon_trk <- ifelse(CompleteTotal$lon_trk<0, CompleteTotal$lon_trk+360, CompleteTotal$lon_trk)

CompleteTotal <- CompleteTotal[ , -c(15,18)]  #extra lat lon cols don't need
 
#filter out distances based on closest distance of a whale
d2land2 <- data.frame(d2land$ID, d2land$survey, d2land$acid,d2land$bath_m, d2land$dist, d2land$peak)
max(d2land2$d2land.dist)
CompleteTotal <- filter(CompleteTotal, dist <= -15)

sum(is.na(CompleteTotal[,18:29]))     # Total NA for env data columns
colSums(is.na(CompleteTotal[,18:29])) # Total NA per column


saveRDS(CompleteTotal, here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_raw.rda')))

```

#Scale data
```{r}
#FILTER, SCALE and CENTER PREDICTORS

#9/13/20: filter out distances based on closest distance of a whale (DONE ABOVE)
# d2land2 <- data.frame(d2land$ID, d2land$survey, d2land$acid,d2land$bath_m, d2land$dist, d2land$peak)
# max(d2land2$d2land.dist)
# CompleteTotal <- filter(CompleteTotal, dist <= -15)
# saveRDS(CompleteTotal, here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '.rda')))


CompletePreds <- dplyr::select(CompleteTotal, sst:d2smt)
CompleteScaledtmp <- scale(CompletePreds, scale = TRUE)

CompleteScaled <- data.frame(dplyr::select(CompleteTotal, pa:distance), CompleteScaledtmp, dplyr::select(CompleteTotal, peak:year))

write.csv(CompleteScaled, here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_scaled.csv')), row.names = FALSE)

saveRDS(CompleteScaled, here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda')))


```



#OLD THINGS -----------

#FOR 700m EDIT:
Load and bind data from all surveys
```{r}
Complete1641 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1641_', gridsize, 'km_', loctype2, '_complete700')))

Complete1303 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1303_', gridsize, 'km_', loctype2, '_complete700')))

Complete1604 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1604_', gridsize, 'km_', loctype2, '_complete700')))

Complete1705 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1705_', gridsize, 'km_', loctype2, '_complete700')))

Complete1706 <- readRDS(here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('1706_', gridsize, 'km_', loctype2, '_complete700')))


CompleteTotal700 <- bind_rows(Complete1641, Complete1303, Complete1604, Complete1705, Complete1706)   

sum(is.na(CompleteTotal700[,15:26]))     # Total NA for env data columns
colSums(is.na(CompleteTotal700[,15:26])) # Total NA per column


saveRDS(CompleteTotal700, here::here(paste0('output/envData/', gridsize, ' km-', loctype), paste0('CompletePm_', gridsize, 'km_', loctype2, '_raw700.rda')))

```




A little extra clean-up
```{r}
env5_bath = filter(env5_bath, bath_m < -100)
env12_d2smt   <- filter(env12_d2smt, effCells %in% env5_bath$effCells)
env1_sst   <- filter(env1_sst, effCells %in% env5_bath$effCells)
env2_chla   <- filter(env2_chla, effCells %in% env5_bath$effCells)
env3_godas600 <- filter(env3_godas600, effCells %in% env5_bath$effCells)
env4_ww3 <- filter(env4_ww3, effCells %in% env5_bath$effCells)
env67_slpasp <- filter(env67_slpasp, effCells %in% env5_bath$effCells)
env8_d2land <- filter(env8_d2land, effCells %in% env5_bath$effCells)
env9_ssh <- filter(env9_ssh, effCells %in% env5_bath$effCells)
env10_sshsd <- filter(env10_sshsd, effCells %in% env5_bath$effCells)
env11_eke <- filter(env11_eke, effCells %in% env5_bath$effCells)


saveRDS(env12_d2smt, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_dist2smt.rda')))
saveRDS(env1_sst, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_sst.rda')))
saveRDS(env2_chla, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_chla.rda')))
saveRDS(env3_godas600, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_temp600.rda')))
saveRDS(env4_ww3, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_wavepow.rda')))
saveRDS(env67_slpasp, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_slpasp.rda')))
saveRDS(env8_d2land, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_d2land.rda')))
saveRDS(env9_ssh, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_ssh.rda')))
saveRDS(env10_sshsd, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_sshsd.rda')))
saveRDS(env11_eke, here::here('output/envData/10 km-localized', paste0(surveynum, '_', gridsize, 'km_eke.rda')))

```

