---
title: "Kruskal-Wallis Test"
author: "Yvonne Barkley"
date: "6/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Kruskal-Wallis Test for nonparametric comparisons
don't use, apparently this is the wrong test 
The null hypothesis = populations are identical. If p<0.05, then populations are nonidentical/different.
1. Test if there's a significant difference between peaks A, B, and the trackline location for localized encs.
2. Also test if there's a difference between the best, min, max of each peak.
```{r}
#Load tidy sperm whale data and filter for localized encounters only
SwEnvDataLocTrk <- read_csv(here::here('output', 'SpermiesWithEnvDataLocTrk_20200527-scaled.csv'))
SwEnvData <- read_csv(here::here('output', 'SpermiesWithEnvData_20200527.csv'))

#1.
#for trackline location data
swtrk <- select(SwEnvDataLocTrk, ID, lon_trk, lat_trk, sstAQ_m:eke)
swtrk_sub <- distinct(swtrk, ID, .keep_all = TRUE)
colnames(swtrk_sub)[colnames(swtrk_sub) %in% c('lon_trk', 'lat_trk')] <- c('lon', 'lat')

swtrklong <- swtrk_sub %>% pivot_longer(-(ID:lat_trk),names_to = 'variable', values_to = 'values')
swtrklong$peak <- 'trk'
swtrk_sub$peak <- 'trk'
swtrk_sub$d2smt_km <- 'NA'
swtrk_sub <- select(swtrk_sub, ID:d2land_km, d2smt_km, everything()) #reorder columns


sw.bestA <- filter(SwEnvData, SwEnvData$type == 'best' & SwEnvData$peak == 'A')
swAsub <- select(sw.bestA, ID, lon, lat, sstAQ_m:eke)
sw.bestAlong <- swAsub %>% pivot_longer(-(ID:lat),names_to = 'variable', values_to = 'values')
sw.bestAlong$peak <- 'A'
swAsub$peak <- 'A'

sw.bestB <- filter(SwEnvData, SwEnvData$type == 'best' & SwEnvData$peak == 'B')
swBsub <- select(sw.bestB, ID, lon, lat, sstAQ_m:eke)
sw.bestBlong <- swBsub %>% pivot_longer(-(ID:lat),names_to = 'variable', values_to = 'values')
sw.bestBlong$peak <- 'B'
swBsub$peak <- 'B'

# swanova <- rbind(swtrklong, sw.bestAlong, sw.bestBlong)
######Combine trackline and estimated locations for localized whales 

swKW <- rbind(swtrk_sub, swAsub, swBsub)
swKW$peak <- as.factor(swKW$peak)
swKW <- swKW[, -c(2,3,5:6)]

# https://stat-methods.com/home/kruskal-wallis-r/
library("ggplot2")
library("qqplotr")
library("dplyr")
library("DescTools")
library("FSA")
library("PMCMRplus")

swKW %>% filter(peak != 'trk') %>% mutate_at(vars(d2smt_km), as.numeric) %>% select(peak, d2smt_km) %>% group_by(peak) %>% 
  dplyr::summarise(n=n(), 
            mean=mean(d2smt_km, na.rm = TRUE), 
            sd=sd(d2smt_km, na.rm = TRUE),
            stderr=sd/sqrt(n),
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(d2smt_km, na.rm = TRUE),
            min=min(d2smt_km, na.rm = TRUE), 
            max=max(d2smt_km, na.rm = TRUE),
            IQR=IQR(d2smt_km, na.rm = TRUE),
            LCLmed = MedianCI(d2smt_km, na.rm=TRUE)[2],
            UCLmed = MedianCI(d2smt_km, na.rm=TRUE)[3])
```

```{r}
#Test each group for normality. A p-value < 0.05 would indicate that we should reject the assumption of normality (so it's not normally distributed if p<0.05. 
shaptest <- lapply(swKW[,c(4,7:13,15:19)], shapiro.test)
shapres <- sapply(shaptest, `[`, c("statistic", "p.value"))
``` 
The wind variable, wind_ms, is the only normally distributed variable where p.value > 0.05.


```{r}
#Perform QQ plots by group
ggplot(data = swKW, mapping = aes(sample = par_m, color = peak, fill = peak)) +
  # stat_qq_band()#(alpha=0.5, conf=0.95, bandType = "pointwise") +
  stat_qq_line() +
  stat_qq_point(col="black") +
  facet_wrap(~ peak, scales = "free") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") + theme_bw() 
```

```{r}
#Perform the Kruskal-Wallis test
kwresults = list()
kwdf = NULL
for (i in names(swKW[,c(4,7:13,16:19)])){ #skips d2smt for now
  
  kwresults[[i]] <- kruskal.test(formula(paste(i, "~peak")), data = swKW)
  kwt <- cbind(kwresults[[i]]$data.name, kwresults[[i]]$p.value)
  kwdf <- as.data.frame(rbind(kwdf, kwt))
}
colnames(kwdf) <- c('test', 'pval')

#for single variable     
swKW2 <- swKW %>% filter(peak != 'trk') %>% droplevels() %>% mutate_at(vars(d2smt_km), as.numeric)
kwsmt <- kruskal.test(d2smt_km ~ peak, data=swKW2)
kwdf <- rbind(kwdf, cbind(`test`= kwsmt$data.name, `pval`= kwsmt$p.value))
```
No non-normal variable is significantly different between peaks or trackline locations 

```{r}
#test wind variable using anova

sw_wind <- aov(wind_ms~peak, data=swKW)
aov <-anova(sw_wind)

windnam <- paste('wind_ms', 'by', 'peak')
windnam <- as.data.frame(cbind(`test` = windnam, `pval`=aov$`Pr(>F)`[1]))
kwdf <- rbind(kwdf, windnam)

write.csv(kwdf, here::here('output', 'Spermies_KruskalWallis_BtwnPeaks.csv'), row.names = F)

```
The normally distributed wind variable also does not significantly differ between peaks and trackline locations.



