---
title: "Examine Variables for Foraging Groups"
author: "Yvonne Barkley"
date: "10/22/2020"
output:
 pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["flafter"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = '', fig.width =8, fig.height = 6, message=FALSE, tidy.opts=list(width.cutoff=60))
```


Looking at the contour plots of temperature at 584m, there is some spatial variation in the contours in the northern portion of the NWHI region near the foraging whale locations. 
Different numbers of knots were tested in the models to see how much these contours changed with increasing 'wiggliness.'

  * gam(Temp584m ~  s(Longitude, Latitude, k=10)
  * gam(Temp584m ~  s(Longitude, Latitude, k=50) 
  * gam(Temp584m ~  s(Longitude, Latitude, k=100)
    
Contour plots suggest that variability in temperature exists at this depth near the northern portion and could be affecting sperm whale prey in some way. Therefore, including the standard deviation of Temp at 584m and the EKE at 584m in the models could provide more information and help explain some of the processes occuring at depth that may be creating more suitable foraging habitat compared to other parts of the archipelago.


```{r eval=TRUE, echo=FALSE, fig.cap='Contour plot of temperature at 584 m.' }
require(mgcv)
survey = 'AllSurveys'
gridsize = 25
loctype = 'Combined'
loctype2 = 'Comb'
PmFor <- readRDS(here::here( paste0('data/Pm_ForagingOnly.rda') ))
trainHunt2 <- readRDS(here::here(  paste0('output/models/',loctype, '/data/Train_',gridsize, 'km_', loctype2, '_HuntNEW.rda')  ))

modTemp584k100 <- gam(Temp584m ~  s(Longitude, Latitude, k=100) , data = PmFor, select = TRUE, method = "REML") 
modTemp584k10 <- gam(Temp584m ~  s(Longitude, Latitude, k=10) , data = PmFor, select = TRUE, method = "REML") 
modTemp584k50 <- gam(Temp584m ~  s(Longitude, Latitude, k=50) , data = PmFor, select = TRUE, method = "REML") 


par(mfrow = c(1,1))

plot(modTemp584k10, select = 1, scheme = 2, lwd = 2, main = 'Temp at 584m ~ LAT:LON, k=10') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(PmFor, pa >0) #get points for whales to plot
points(whales$Longitude, whales$Latitude, pch = 10, col='mediumslateblue', lwd=1.25, cex=1)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)


plot(modTemp584k50, select = 1, scheme = 2, lwd = 2, main = 'Temp at 584m ~ LAT:LON, k=50') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(PmFor, pa >0) #get points for whales to plot
points(whales$Longitude, whales$Latitude, pch = 10, col='mediumslateblue', lwd=1.25, cex=1)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)



plot(modTemp584k100, select = 1, scheme = 2, lwd = 2, main = 'Temp at 584m ~ LAT:LON, k=100') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(PmFor, pa >0) #get points for whales to plot
points(whales$Longitude, whales$Latitude, pch = 10, col='mediumslateblue', lwd=1.25, cex=1)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)

```


### Variables at 584 m

The standard deviation of temperature at 584m and EKE at 584m were computed to test whether these variable at depth explained the presence of foraging whales.

Each variable was not significant when included separately in the models nor did they contribute to explaining the data in the best models with Temp at 584m, log(Chla), and SSHsd.


```{r echo=FALSE, fig.cap='Standard Deviation of Temp at 584m. The spatial plot is the result of fitting the a spatial smoother to the standard deviation of Temp584 to examine the spatial variability.'}
PmFor2 <- subset(PmFor, pa >0)
# par(mfrow= c(2,2))
# hist(PmFor2$temp600_sd.r, main = '')
# plot(PmFor$temp600_sd.r, PmFor$pa, ylab="Foraging Groups per cell", xlab="SD Temp at 584m")
# hist(PmFor2$eke600m.r, main ='')
# plot(PmFor$eke600m.r, PmFor$pa, ylab="Foraging Groups per cell", xlab="EKE at 584m")

mlay <- rbind(c(2,3),c(1,1))
# print(mlay)
# layout.show(1)
layout(mlay)
par(mar=c(4,4,1,1))
modTemp584mSD <- gam(pa ~  s(Longitude, Latitude, k=10) +
                    s(temp600_sd.r, k=3) + offset(log.effort) , data = trainHunt2, family = nb,
                    link = 'log', select = TRUE, method = "REML")
summary(modTemp584mSD)
modTemp584sd <- gam(temp600_sd.r ~ s(Longitude, Latitude, k=10) , data = PmFor, select = TRUE, method = "REML") 
# summary(modTemp584sd)
# gam.check(modTemp584sd)

### TempSD 584m Plots####
plot(modTemp584sd, select = 1, scheme = 2, lwd = 2, main = 'Temp at 584mSD ~ LAT:LON, k=10') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(PmFor, pa >0) #get points for whales to plot
points(whales$Longitude, whales$Latitude, pch = 10, col='mediumslateblue', lwd=1.25, cex=1)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)
hist(PmFor2$temp600_sd.r, main = '', xlab="SD Temp at 584m")
plot(PmFor$temp600_sd.r, PmFor$pa, ylab="Foraging Groups per cell", xlab="SD Temp at 584m")
```


```{r echo=FALSE, fig.cap='EKE at 584 m. The spatial plot is the result of fitting the a spatial smoother to the standard deviation of EKE584 to examine the spatial variability.'}
modEKE584m <- gam(pa ~  s(Longitude, Latitude, k=10) +
                   s(eke600m.r, k=3) + offset(log.effort), data = trainHunt2, family = nb,
                   link = 'log', select = TRUE, method = "REML")
summary(modEKE584m)


modEKE584 <- gam(eke600m.r ~  s(Longitude, Latitude, k=10) , data = trainHunt2, select = TRUE, method = "REML") 
# summary(modEKE584)
# gam.check(modEKE584)

### EKE 584m Plots####
mlay <- rbind(c(2,3),c(1,1))
# print(mlay)
# layout.show(1)
layout(mlay)
par(mar=c(4,4,1,1))
plot(modEKE584, select = 1, scheme = 2, lwd = 2, main = 'EKE at 584m ~ LAT:LON, k=10') #select the first smoother, select = 1
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)
whales <- subset(PmFor, pa >0) #get points for whales to plot
points(whales$Longitude, whales$Latitude, pch = 10, col='mediumslateblue', lwd=1.25, cex=1)
points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.5)
hist(PmFor2$eke600m.r, main ='',  xlab="EKE at 584m")
plot(PmFor$eke600m.r, PmFor$pa, ylab="Foraging Groups per cell", xlab="EKE at 584m")

```

#Nov 1: Include SD vars in full model
```{r}
modTest <- gam(pa ~  s(Longitude, Latitude, k=10) + s(depth.r, k=3) + s(distland.r, k=3)  + s(distseamt.r, k=3) +  s(SST, k=3) + s(log.Chla, k=3) + s(Temp584m, k=3) + s(SSH, k=3) + s(SSHsd, k=3) + s(log.eke, k=3) + s(log.wp, k=3) +  s(eke600m.r, k=3) + s(temp600_sd.r, k=3) + s(SSTsd.r, k=3) + offset(log.effort)  , data = trainHunt2, family = nb, link = 'log', select = TRUE, method = "REML")

summary(modTest)

modTest2 <- gam(pa ~  s(Longitude, Latitude, k=10) + s(SST, k=3) + s(log.Chla, k=3)  + s(Temp584m, k=3) + s(SSHsd, k=3) + s(SSTsd.r, k=3) + offset(log.effort) , data = trainHunt2, family = nb, link = 'log', select = TRUE, method = "REML")
summary(modTest2)


modTest3 <- gam(pa ~  s(Longitude, Latitude, k=10) + s(SST, k=3) + s(log.Chla, k=3)  + s(Temp584m, k=3) + s(SSHsd, k=3)  + offset(log.effort) , data = trainHunt2, family = nb, link = 'log', select = TRUE, method = "REML")
summary(modTest3)

modTest4 <- gam(pa ~  s(Longitude, Latitude, k=10) + s(SSTsd.r, k=3) + offset(log.effort) , data = trainHunt2, family = nb, link = 'log', select = TRUE, method = "REML")
summary(modTest4)

```



```{r echo=FALSE, eval=FALSE}
# Models with just the new variables

modTemp584mSD <- gam(pa ~  s(Longitude, Latitude, k=10) +
                    s(temp600_sd.r, k=3) + offset(log.effort) , data = trainHunt2, family = nb,
                    link = 'log', select = TRUE, method = "REML")
summary(modTemp584mSD) 

modEKE584m <- gam(pa ~  s(Longitude, Latitude, k=10) +
                   s(eke600m.r, k=3) + offset(log.effort) , data = trainHunt2, family = nb,
                   link = 'log', select = TRUE, method = "REML")
summary(modEKE584m) 

modTest <- gam(pa ~  s(Longitude, Latitude, k=10) + s(Temp584m, k=3) + 
                       s(temp600_sd.r, k=3) + s(eke600m.r, k=3) +
                        offset(log.effort) , data = trainHunt2, family = nb,
                    link = 'log', select = TRUE, method = "REML")
summary(modTest) 
```


## Find FORAGING Sperm Whales with high SD of SSH and Temp584m
This code is mainly for looking at specific examples of whales found in areas of higher SSHsd or Temp584sd per Margaret's suggestion. 
The rasters for each variable were saved in the 11_AddCovs_ForJeff.R script, which is also where the EKEsd at 584 m and the SSTsd was calculated. 
```{r}
setwd("C:\\Users\\yvers\\Documents\\CHP 3\\SpermWhales\\data\\For Zack-data/")
covar_dir = "C:/Users/yvers/Documents/CHP 3/SpermWhales/data/For Zack-data/"
mhi <- readOGR("whales/mhi_coast/Coastline.shp")
nwhi <- readOGR("whales/nwhi_coast/Coastline.shp")

SSH <- brick(paste0(covar_dir, "SSH.nc"), 
             varname = "zos")
mhi2 <- raster::shift(x = mhi,
                      dx = +360)
nwhi2 <- raster::shift(x = nwhi,
                      dx = +360)
#there are better for the NWHI
hawaiiMap <- readRDS(here::here(paste0('data/hawaiiMap.rda')))
hawaiiMap$Longitude2 = ifelse(hawaiiMap$Longitude<1, hawaiiMap$Longitude+360, hawaiiMap$Longitude)

## Use original SSHsd data - 'Z' ####
# Using focal on the SSH raster gives different values than original data...So used orig df
dateSplit = data.frame(1, time=as.character(as.Date(sshZtidya$date)))  #some data finagling
dateSplit2 = separate(dateSplit, "time",  c("year", "month", "day")) #splits date into year, month, day
sshZtidya$year = as.numeric(dateSplit2$year)
sshZtidya$month = as.numeric(dateSplit2$month)

PmFor2$LongP <- PmFor2$Longitudep-360


# Explore Data for gradients in foraging whales
plot(PmFor2$Longitude, PmFor2$SSHsd)
plot(PmFor2$Latitude, PmFor2$SSHsd)
plot(PmFor$Longitude, PmFor$SSHsd)

plot(PmFor2$Longitude, PmFor2$SSH)

#Temp 584m ######
plot(PmFor2$Longitude, PmFor2$Temp584m)
PmHiT584lo <- subset(PmFor2, Longitude > 200 & Longitude < 205 & Temp584m < 6 & survey == 1604) # low temp
PmHiT584hi <- subset(PmFor2, Longitude < 180 & Temp584m > 7.5 ) # hi temp
PmHiT584hib <- subset(PmFor2, Longitude > 185 & Longitude < 190 & Temp584m > 7.5 & survey == 1641) # hi temp

# SSH straight up ####
PmHiSSH <- subset(PmFor2, Longitude > 198 & Longitude < 202 & SSH > 0.7 & survey == 1604)
PmHiSSHb <- subset(PmFor2, Longitude > 195 & Longitude < 210 & survey == 1604)
PmHiSSHb <- subset(PmHiSSHb, !(effCells %in% PmHiSSH$effCells))

#put in the extra detection for effCell = 740
PmHiSSH <- rbind(PmHiSSH, PmHiSSH[3,])
getPm <- pmDetsGrid1604[which(pmDetsGrid1604$acid == 14), ]

PmHiSSH[4,3] <- getPm$UTC
PmHiSSH[4,4] <- getPm$ID
PmHiSSH[4,6] <- getPm$acid
PmHiSSH[4,7] <- getPm$sid
PmHiSSH[4,13] <- getPm$Longitude
PmHiSSH[4,14] <- getPm$Latitude
#########
```

#Nov 1 USE: MHI High SSHsd Foragers - Sept 2010
```{r}
# found foraging groups that were sighted or localized for example
sshZtidy2017 = subset(sshZtidya, year == 2017)
sshZtidy2010 = subset(sshZtidya, year == 2010)

#JUST from FORAGING
PmSSHsdhi <- subset(PmFor2, SSHsd > 0.010 & pa >0  )#& survey == 1706)

PmSSHsdhi2 <- subset( PmSSHsdhi, SSHsd > 0.01 & pa > 0 & Longitude > 200)
PmSSHsdhi2 <- dplyr::select(PmSSHsdhi2, pa:Latitudep, SSHsd, peak, loc)
PmForSSHsdHi2 <- PmSSHsdhi2[PmSSHsdhi2$acid %in% c(155,171),    ]

sshsd201009 = subset(sshZtidy2010, month == 09)
sshsd201009$sshsd = sshsd201009$sshsd*100 
SSH201009sd <- raster::rasterFromXYZ(sshsd201009[, c(1,2,4)])
sshsdHI_extent <- extent(199, 206, 18, 25)
SSH201009sdcrop <- raster::crop(SSH201009sd, y = sshsdHI_extent)
SSH201009sdcrop <- raster::shift(x = SSH201009sdcrop,
                      dx = -360)
mhi3 <- raster::crop(mhi2, y = sshsdHI_extent)
mhi3 <- raster::shift(x = mhi3,
                      dx = -360)
pal2 <- terrain.colors(10)

png(here::here('output/models/ForMcManus/Sept2010_PmSSHsd.png'), width = 590, height = 600, pointsize = 20)

plot(SSH201009sdcrop, col = pal2, main='Standard Deviation of SSH, Sept 2010')
points( Latitudep ~ Longitudep, data = PmForSSHsdHi2,
       pch = 16, col='red', lwd=1, cex=1 )
plot(mhi3, add = T, col = 'black', lwd = 3)  
north.arrow(xb= 199.6 ,  yb= 24 ,  len=  0.25,  lab = "N")
scalebar(250, xy=c(199.3, 18.5), type='bar', lonlat = TRUE, below='km')

dev.off()



## plot SSH for same Sept 2010 whales ####
time_idx <- which( names(SSH) %in% paste0("X2010.09.16.00.00.00"))
SSH201009 <- subset(x = SSH, 
              subset = time_idx)
SSH201009crop <- raster::crop(SSH201009, y = sshsdHI_extent)
SSH201009crop <- SSH201009crop*100

png(here::here('output/models/ForMcManus/Sept2010_PmSSH.png'), width = 590, height = 600, pointsize = 20)

plot(SSH201009crop, col = pal2 , main = 'SSH, Sept 2010')
plot(mhi3, add = T, col = 'black', lwd=3)  # can al
points( Latitudep ~ Longitudep, data = PmForSSHsdHi2,
       pch = 16, col='red', lwd=1, cex=1 )
north.arrow(xb= 199.6 ,  yb= 24 ,  len=  0.25,  lab = "N")
scalebar(250, xy=c(199.3, 18.5), type='bar', lonlat = TRUE, below='km')

dev.off()
```

#Nov 1 USE: High Temp at 584m and Stand Dev
```{r}
T584 <- readRDS(paste0(covar_dir, "2010-temp584m.RDS")) #loads raster stack of temp at 584
T584sd <- readRDS(paste0(covar_dir, "201008-temp584msd.RDS")) #loads raster layer 08/2010 of temp at 584
# plot(PmFor2$Longitude, PmFor2$SSTsd.r)
# PmSSTsdhi <- subset( PmFor2, SSTsd.r > 0.10 & pa >0  )#& survey == 1706)
# plot(PmSSTsdhi$Longitude, PmSSTsdhi$SSTsd.r)
# PmSSTsdhi2 <- subset(PmSSTsdhi, Longitude < 180  )#& survey == 1706)
# 
# plot(PmFor2$Longitude, PmFor2$SST)
# plot(PmFor2$Longitude, PmFor2$Temp584m)
plot(PmFor2$Longitude, PmFor2$temp600_sd.r)

PmT584sdhi <- subset(PmFor2, Longitude > 205)#& survey == 1706)

#find some in the NWHI
PmT584sdhi <- subset(PmFor2, Longitude > 187 & Longitude <= 195 & temp600_sd.r > 0.008)#& survey == 1706)
PmT584sdhiB <- subset(PmFor2, Longitude > 187 & Longitude <= 195 & survey == 1706)
PmT584sdhiB <- PmT584sdhiB[1:4,]

plot(PmT584sdhi$Longitude, PmT584sdhi$temp600_sd.r)
plot(PmT584sdhi$Longitude, PmT584sdhi$Temp584m)

## Temp and SD Temp 584 High - NWHI July 2017####

#these are the sperm whale groups in the paper
PmT584sdhi3 <- PmT584sdhi[PmT584sdhi$acid %in% c(73,98),    ] #pull out encs based on ACid

# Saved in 11_AddCovs_ForJeff.R
T584b <- readRDS(paste0(covar_dir, "2017-temp584m.RDS")) #loads raster stack of temp at 584
T584sd2 <- readRDS(paste0(covar_dir, "201707-temp584msd.RDS")) #loads raster layer 08/2010 of temp at 584


## Plot Temp584 - July 2017 ####
time_idx <- which( names(T584b) %in% paste0("X2017_07"))
T584201707 <- subset(x = T584b, 
              subset = time_idx)
T584201707 <- raster::resample(x = T584201707,
                                 y = SSH201009)
T584201707 <- T584201707-273.15
T584_extent <- extent(185, 195, 23, 28)
T584201707crop <- raster::crop(T584201707, y = T584_extent)
# T584201707crop <- T584201707crop-273.15
nwhi4 <- raster::crop(nwhi2, y = T584_extent)

# save plot
png(here::here('output/models/ForMcManus/July2017_PmT584.png'), width = 800, height = 530, pointsize = 20)

plot(T584201707crop, col = pal2, main = 'Temp at 584 m, July 2017')
# plot(nwhi4, add = T, col = 'black', lwd=3)  # can al
points( Latitudep ~ Longitudep, data = PmT584sdhi3,
       pch = 16, col='red', lwd=1, cex=1 )

hawaiiMap2 = subset(hawaiiMap, Longitude2 >= 185 & Longitude2 <= 195) 
points(hawaiiMap2$Longitude2, hawaiiMap2$Latitude, pch=20, cex = 0.75)
north.arrow(xb= 185.5 ,  yb= 27 ,  len=  0.25,  lab = "N")
scalebar(250, xy=c(185.3, 23.5), type='bar', lonlat = TRUE, below='km')

dev.off()

## Plot Stand Dev of Temp584 - July 2017 ####
T584201707sd <- raster::resample(x = T584sd2,
                                 y = SSH201009)
T584201707sdcrop <- raster::crop(T584201707sd, y = T584_extent)

#save plot
png(here::here('output/models/ForMcManus/July2017_PmT584sd.png'), width = 800, height = 530, pointsize = 20)

plot(T584201707sdcrop, col = pal2, main = 'Standard Deviaton of Temp at 584m, July 2017')
points( Latitudep ~ Longitudep, data = PmT584sdhi3,
       pch = 16, col='red', lwd=1, cex=1 )

# points( Latitudep ~ Longitudep, data = PmT584sdhiB,  #looking at whales in surrounding area with lower Tempsd
#        pch = 16, col='orange', lwd=1, cex=1 )
# hawaiiMap2 = subset(hawaiiMap, Longitude2 >= 185 & Longitude2 <= 195) 
points(hawaiiMap2$Longitude2, hawaiiMap2$Latitude, pch=20, cex = 0.75)
north.arrow(xb= 185.5 ,  yb= 27 ,  len=  0.25,  lab = "N")
scalebar(250, xy=c(185.3, 23.5), type='bar', lonlat = TRUE, below='km')

dev.off()

## Temp and SD Temp 584 High- MHI Aug 2010 ####

PmT584sdhi2 <- subset(PmFor2, Longitude > 206.5)#& survey == 1706)

# Subset the raster for Aug 2010
# Resample to same as SSH raster resolution
time_idx <- which( names(T584) %in% paste0("X2010_08"))
T584201008 <- subset(x = T584, 
              subset = time_idx)
T584201008 <- raster::resample(x = T584201008,
                                 y = SSH201009)

T584_extent <- extent(202, 210, 17, 22)
T584201008crop <- raster::crop(T584201008, y = T584_extent)
T584201008crop <- T584201008crop-273.15
mhi4 <- raster::crop(mhi2, y = T584_extent)

plot(T584201008crop, col = pal2, main = 'Temp at 584m, Aug 2010')
plot(mhi4, add = T, col = 'black', lwd=3)  # can al
points( Latitudep ~ Longitudep, data = PmT584sdhi2,
       pch = 16, col='red', lwd=1, cex=1 )
north.arrow(xb= 209.6 ,  yb= 21 ,  len=  0.25,  lab = "N")
scalebar(250, xy=c(207.4, 17.5), type='bar', lonlat = TRUE, below='km')


# crop and plot raster of Temp584 sd
T584201008sd <- raster::resample(x = T584sd,
                                 y = SSH201009)
T584201008sdcrop <- raster::crop(T584201008sd, y = T584_extent)

plot(T584201008sdcrop, col = pal2, main = 'Temp at 584m Stand Dev, Aug 2010')
plot(mhi4, add = T, col = 'black', lwd=3) 
points( Latitudep ~ Longitudep, data = PmT584sdhi2,
       pch = 16, col='red', lwd=1, cex=1 )
north.arrow(xb= 209.6 ,  yb= 21 ,  len=  0.25,  lab = "N")
scalebar(250, xy=c(207.4, 17.5), type='bar', lonlat = TRUE, below='km')


```


#DIDN'T Use
```{r}
PmSSHsdhi3 <- subset( PmSSHsdhi, SSHsd > 0.01 & pa > 0 & Longitude < 190)
PmSSHsdhi3 <- dplyr::select(PmSSHsdhi3, pa:Latitude, SSHsd, peak, loc)



plot(PmSSHsdhi$Longitude, PmSSHsdhi$SSHsd)
PmForSSHsdHi <- PmSSHsdhi[PmSSHsdhi$acid %in% c(292,293),    ]

 #add to PmCombo2 to test
sshsd201709 = subset(sshZtidy2017, month == 09)
SSH201709sd <- raster::rasterFromXYZ(sshsd201709[, c(1,2,4)])

#PLOT

pal2 <- terrain.colors(8)
plot(SSH201709sd,col = pal2, ylim = c(15,31))
plot(nwhi2, add = T, col = 'black')
plot(mhi2, add = T, col = 'black', lwd = 4, bg = 'red')  # can also specify border
points( Latitudep ~ Longitudep, data = PmForSSHsdHi,
       pch = 16, col='red', lwd=1, cex=0.8 )

sshsdHI_extent <- extent(180, 190, 22, 30)
SSH201709sdcrop <- raster::crop(SSH201709sd, y = sshsdHI_extent)
mhi3 <- raster::crop(mhi2, y = sshsdHI_extent)
nwhi3 <- raster::crop(nwhi2, y = sshsdHI_extent)

png('PmSSHsdHigh_Sept2017.png', width = 800, height = 500, pointsize = 20)

# spplot(SSH2016sd)  # in sp, needs axes labels
plot(SSH201709sdcrop, col = pal2 , main = 'High SSHsd, Sept 2017')
hawaiiMap2 = subset(hawaiiMap, Longitude2 >= 180 & Longitude2 <= 190) 
points(hawaiiMap2$Longitude2, hawaiiMap2$Latitude, pch=20, cex = 0.75)
points( Latitudep ~ Longitudep, data = PmForSSHsdHi,
       pch = 16, col='purple', lwd=1, cex=1 )

north.arrow(xb= 197 ,  yb= 28.5  ,  len=  0.25,  lab = "N")
scalebar(250, xy=click(), type='bar', lonlat = TRUE, below='km')

# raster::contour(SSHsd2017crop, add=T, labcex = 0.5, nlevels=5)  #contours are busy
# plot(mhi2, add = T, col = 'black')  # can also specify border
plot(nwhi3, add = T, col = 'black')

points(hawaiiMap$Longitude2, hawaiiMap$Latitude, pch=20, cex = 0.75)

points( Latitudep ~ Longitudep, data = PmSSHsd1706,
       pch = 16, col='red', lwd=1, cex=1 )


dev.off()

```

#DIDN'T USE: plot for high SSHsd - August 2017 
```{r}
require(rasterVis)
require(GISTools)
setwd("C:/Users/yvers/Documents/CHP 3/SpermWhales/output/models/ForMcManus")

PmSSHsdAllHi <- subset(PmCombo, sshsd.r > 0.015 & pa >0 & loc==1 | sid <999 & sshsd.r > 0.015)
PmSSHsdAllHi <- subset(PmSSHsdAllHi, sshsd.r > 0.015 & pa >0 & survey == 1706)
PmSSHsd1706 <- PmSSHsdAllHi[PmSSHsdAllHi$acid %in% c(149, 186, 208),    ]

sshZtidy2017 = subset(sshZtidya, year == 2017)
sshsd201708 = subset(sshZtidy2017, month == 08)

SSH201708sd <- raster::rasterFromXYZ(sshsd201708[, c(1,2,4)])


pal2 <- terrain.colors(8)
png('PmSSHsdHigh_Aug2017_Big.png', width = 800, height = 500, pointsize = 20)

plot(SSH201708sd,col = pal2, ylim = c(15,31))
plot(nwhi2, add = T, col = 'black')
plot(mhi2, add = T, col = 'black', lwd = 4, bg = 'red')  # can also specify border

points( Latitudep ~ Longitudep, data = PmSSHsd1706,
       pch = 16, col='red', lwd=1, cex=0.8 )
dev.off()


sshsdHI_extent <- extent(183, 199, 22, 30)
SSHsd2017crop <- raster::crop(SSH2017sd, y = sshsdHI_extent)
mhi3 <- raster::crop(mhi2, y = sshsdHI_extent)
nwhi3 <- raster::crop(nwhi2, y = sshsdHI_extent)

png('PmSSHsdHigh_Aug2017.png', width = 800, height = 500, pointsize = 20)

# spplot(SSH2016sd)  # in sp, needs axes labels
plot(SSHsd2017crop, col = pal2 , main = 'High SSHsd, August 2017')
north.arrow(xb= 197 ,  yb= 28.5  ,  len=  0.25,  lab = "N")
scalebar(250, xy=click(), type='bar', lonlat = TRUE, below='km')

# raster::contour(SSHsd2017crop, add=T, labcex = 0.5, nlevels=5)  #contours are busy
# plot(mhi2, add = T, col = 'black')  # can also specify border
plot(nwhi3, add = T, col = 'black')
hawaiiMap2 = subset(hawaiiMap, Latitude >= 22)
points(hawaiiMap2$Longitude2, hawaiiMap2$Latitude, pch=20, cex = 0.75)

points( Latitudep ~ Longitudep, data = PmSSHsd1706,
       pch = 16, col='red', lwd=1, cex=1 )


dev.off()


## plot SSH for same whales ####

png('PmSSH_Aug2017.png', width = 800, height = 500, pointsize = 20)
time_idx <- which( names(SSH) %in% paste0("X2017.08.16.12.00.00"))
SSH2017 <- subset(x = SSH, 
              subset = time_idx)
SSH2017crop <- raster::crop(SSH2017, y = sshsdHI_extent)
plot(SSH2017crop, col = pal2 , main = 'SSH, August 2017')
raster::contour(SSH2017crop, add=T, labcex = 0.5, nlevels=5)
plot(mhi2, add = T, col = 'black', lwd=8)  # can al
points( Latitudep ~ Longitudep, data = PmSSHsd1706,
       pch = 16, col='red', lwd=1, cex=1 )
dev.off()


```

###plot for low SSHsd - Aug 2010
```{r}

PmSSHsdlo <- subset(PmFor2, Longitude > 205 & SSHsd < 0.0025)

sshZtidy2010 = subset(sshZtidya, year == 2010)
 #add to PmCombo2 to test
sshZtidy2010 = subset(sshZtidy2010, month == 08)

SSH2010sd <- raster::rasterFromXYZ(sshZtidy2010[, c(1,2,4)])

sshsdLO_extent <- extent(202, 210, 17, 22)
SSHsd2010crop <- raster::crop(SSH2010sd, y = sshsdLO_extent)
mhi3 <- raster::crop(mhi2, y = sshsdLO_extent)
nwhi3 <- raster::crop(nwhi2, y = sshsdHI_extent)

png('PmSSHsdLow_Aug2010.png', width = 800, height = 600, pointsize = 20)
pal2 <- terrain.colors(8)
# spplot(SSH2016sd)  # in sp, needs axes labels

plot(SSHsd2010crop, col = pal2 , main = 'SSHsd, August 2010')
raster::contour(SSHsd2010crop, add=T, labcex = 0.5, nlevels=5)
plot(mhi3, add = T, col = 'black', lwd=8)  # can also specify border
# plot(nwhi2, add = T, col = 'black')

points( Latitudep ~ Longitudep, data = PmSSHsdlo,
       pch = 16, col='red', lwd=1, cex=1 )
dev.off()


png('PmSSH_Aug2010.png', width = 800, height = 600, pointsize = 20)

## plot SSH for same whales ####
time_idx <- which( names(SSH) %in% paste0("X2010.08.16.12.00.00"))
SSH2010 <- subset(x = SSH, 
              subset = time_idx)
SSH2010crop <- raster::crop(SSH2010, y = sshsdLO_extent)
plot(SSH2010crop, col = pal2 , main = 'SSH, August 2010')
raster::contour(SSH2010crop, add=T, labcex = 0.5, nlevels=5)
plot(mhi3, add = T, col = 'black', lwd=8)  # can al
points( Latitudep ~ Longitudep, data = PmSSHsdlo,
       pch = 16, col='red', lwd=1, cex=1 )

dev.off()
```

```{r}

# Plot LOW SSH = upwelling  ####

#From foraging whales
plot(PmFor2$Longitude, PmFor2$SSH)
PmSSHlow <- subset(PmFor2, Longitude > 203 & SSH < 0.6) 
PmSSHlow <- dplyr::select(PmSSHlow, UTC, UTCp, ID, Longitude, Latitude, SSH)

SSH <- brick(paste0(covar_dir, "SSH.nc"),
            varname = "zos")
time_idx <- which( names(SSH) %in% paste0("X", "2010", ".09.16.00.00.00") )


SSH2010 <- subset(x = SSH,
               subset = time_idx)  #Only looking at SSH for July 2016


pal2 <- terrain.colors(8)
# spplot(SSH2016sd)  # in sp, needs axes labels
plot(SSH2016sd, col = pal2 )
raster::contour(SSH2016sd, add=T, labcex = 0.5, nlevels=7)
plot(mhi2, add = T, col = 'black')  # can also specify border
plot(nwhi2, add = T, col = 'black')

points( Latitudep ~ Longitudep, data = PmSSHsdhi,
       pch = 16, col='red', lwd=1, cex=0.5 )
points( Latitudep ~ Longitudep, data = PmHiSSHb,
       pch = 16, col='purple', lwd=1, cex=0.5 )


```

