---
title: "SDMs Revisited"
author: "Yvonne Barkley, PhD"
date: "2/7/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
surveynum = 1641
gridsize = 25
loctype = 'Combined'
loctype2 = 'Comb'

```

```{r}
# LOAD EDITED gridEffort and centroidALL - already checked for things in 1_GridData_v4

gridEffort <- readRDS(here::here(  paste0('output/gridEffort/', gridsize, ' km-', loctype, '/gridEffort', surveynum, '_', gridsize, 'km_', loctype2, '.rda') ))

centroidAll <- readRDS(here::here(paste0('output/centroids/', gridsize, ' km-', loctype), paste0('centroids', surveynum, '_', gridsize, 'km_', loctype2, '.rda')))

```




# The Complete Data Set
```{r, echo=F, eval=FALSE, message=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(geoR)
library(tidymv)
library(here)
library(lubridate)
```

```{r echo=FALSE}
#Values for COMBINED MODELS
survey = 'AllSurveys'
gridsize = 25
loctype = 'Combined'
loctype2 = 'Comb'
#data has logged vars and temporal variables
## SAVED from line 251 in 10_PmSDM_AcousticCombined.rmd - Does not include forage column
PmCombo <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_EDIT.rda') )) #no A999's
PmCombo2 <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda') ))

PmCombined <- readRDS( here::here('data/Pm_Combined.rda')) #contains A999's
```


# Sperm whale data used for centroids
```{r}
sw <- read.csv(here::here('output/SpermiesFinalLR_20200919.csv'))

length(unique(sw$ID))


```



#From 10_PmSDM_Foraging

### 2022: Redo Click Codes
```{r}
PmCombo$click_code[PmCombo$ID == "1303_42_21"] <- "cdckrs"  #(was ckrs, added coda)
PmCombo$click_code[PmCombo$ID == "1604_30_999"] <- "rs"     #(was s)
PmCombo$click_code[PmCombo$ID == "1604_67_999"] <- "rs"     #(was s)
PmCombo$click_code[PmCombo$ID == "1705_251_999"] <- "rs"    #(was s)
PmCombo$click_code[PmCombo$ID == "1706_2_999"] <- "rs"      #(was r)
PmCombo$click_code[PmCombo$ID == "1706_41_999"] <- "rs"     #(was s)
# PmCombo$click_code[PmCombo$ID == "1706_135_999"] <- "ckrs" #not included as separate encounter, was combined with 1706A49 in same cell, which is also ckrs
PmCombo$click_code[PmCombo$ID == "1303_71_999"] <- "rs"     #(was s)


```

### Run some totals
Sightings, Sightings/Acoustics, Acoustics Only
```{r}
#sighted only
PmVisOnly <- filter(PmCombined, acid == 999)
aggregate(x = PmVisOnly$pa,                # Specify data column
          by = list(PmVisOnly$survey),              # Specify group indicator
          FUN = sum)

PmVisOnly2 <- readRDS( here::here('data/Pm_Sighting.rda'))

PmVisOnly2_A999 <- filter(PmVisOnly2, acid == 999)
aggregate(x = PmVisOnly2_A999$pa,                # Specify data column
          by = list(PmVisOnly2_A999$survey),              # Specify group indicator
          FUN = sum)

PmVisEd <- readRDS(here::here(  paste0('output/models/Combined/data/SighingOnly_25km_EditDup.rda') ))
PmVisEd_A999 <- filter(PmVisEd, acid == 999)

aggregate(x = PmVisEd_A999$pa,                # Specify data column
          by = list(PmVisEd_A999$survey),              # Specify group indicator
          FUN = sum)

PmVis <- readRDS(here::here(  paste0('output/models/',loctype, '/data/SighingOnly_25km_Combined.rda') ))
PmVisOnly3 <- filter(PmVis, acid == 999)




#Sighted and acoustic acid <999, sid <999
PmVisAc <- filter(PmCombo, acid <999 & sid< 999)
aggregate(x = PmVisAc$pa,                # Specify data column
          by = list(PmVisAc$survey),              # Specify group indicator
          FUN = sum)

#sighted and acoustic acid <999, sid <999 - 2 less 1706 than table
PmAcLoc <- filter(PmCombo, sid == 999 & loc == 1)
aggregate(x = PmAcLoc$pa,                # Specify data column
          by = list(PmAcLoc$survey),              # Specify group indicator
          FUN = sum)

#Acoustics - Localized / Trackline encounters
PmAcLoc2 <- readRDS( here::here('data/Pm_Acoustics.rda'))

#has 2 more 1706 than PmAcLoc (1706_327_999, 1706_325_999) - ALL GOOD with table
PmAcLoc_2 <- filter(PmAcLoc2, sid == 999 & loc == 1)
aggregate(x = PmAcLoc_2$pa,                
          by = list(PmAcLoc_2$survey),    
          FUN = sum)

#has 2 less 1706 than the table & 2 more 1641! WTF
PmAcUnloc <- filter(PmCombo, sid == 999 & loc == 0)
aggregate(x = PmAcUnloc$pa,                
          by = list(PmAcUnloc$survey),    
          FUN = sum)
#has 2 less 1706 than the table!
PmAcUnloc_2 <- filter(PmAcLoc2, sid == 999 & loc == 0)
aggregate(x = PmAcUnloc_2$pa,                
          by = list(PmAcUnloc$survey),    
          FUN = sum)


# Check # of encounters by survey
Pm1303 <- filter(PmCombo, survey == 1303); sum(Pm1303$pa)
Pm2017 <- filter(PmCombo, survey == 1706 | survey ==1705)
sum(Pm2017$pa)

Pm1706b <- filter(CompleteScaled, survey == 1706, acid < 999) #CompleteScaled from 9_DataCleanup

Pm2017_pre <- filter(CompleteScaled, survey == 1705 | survey == 1706)
sum(Pm2017_pre$pa)

```


### 2022: Sanity Checks

## Foraging and Non-Foraging Group Models

#### Foraging models incorporated a subset of the sperm whale encounters from the Combined Models (sighted with acoustics, and acoustics only) that only included encounters with **regular clicks** and **creaks** to indicate foraging.


```{r eval=FALSE,echo=FALSE}
## Forage and Social Group Data Subset ####
#1. Denote forage or not. 
PmFor1 <- filter(PmCombo, grepl('r|k', click_code))   # the 'k' implies ck for creaks because c also for codas
PmFor0 <- filter(PmCombo, !grepl('r|k', click_code))
PmFor1$forage <- 1  #forage
PmFor0$forage <- 0  #non-forage
PmForage <- bind_rows(PmFor1, PmFor0)

nrow(PmFor0)
nrow(PmFor1)

#2. Denote non-foraging
PmPositive <- filter( PmCombo, pa > 0 )
PmAbsent <- filter( PmCombo, pa == 0 )

PmNonFor1 <- filter(PmPositive, !grepl('r|k', click_code))  #subset all encs without r or ck in click_code
PmNonFor0 <- filter(PmPositive, grepl('r|k', click_code))

#designate whether the encounters included codas (1) or not (0)
PmNonFor1$nonfor <- 1
PmNonFor0$nonfor <- 0
PmAbsent$nonfor <- 0

PmNonFor <- rbind(PmNonFor1,PmNonFor0, PmAbsent)


PmNonFor2 <- PmNonFor[, c(1,2,4,54,62)]

nrow(PmNonFor1)  # represent all non-foraging 
nrow(PmNonFor0)  # represent all foraging encounters, which are now 0
```


```{r}
# How many encounters are foraging and non-foraging? 196?! Not 209...
sum(PmFor1$pa) # 109 foraging, not include pa >1, 122
sum(PmNonFor1$pa) #71 non-foraging, not include pa >1, 74

# How many cells have multiple groups, pa > 1? How many total groups? 
PmFor_multi <- PmFor1[which(PmFor1$pa>1),]; length(PmFor_multi$pa);sum(PmFor_multi$pa)
PmNonFor_multi <- PmNonFor1[which(PmNonFor1$pa>1),]; length(PmNonFor_multi$pa); sum(PmNonFor_multi$pa)


# Are encounters used >1 between foraging and non-foraging models? NO
PmFor1$ID %in% PmNonFor1 
PmNonFor1$ID %in% PmFor1
```

```{r}


```


