---
title: "SDMs Revisited"
author: "Yvonne Barkley, PhD"
date: "2/7/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

This is mostly for summarizing the data after implementing the new 16km-grid.
FIG 3 Barplot is made here using 'fig3' dataframe.

```{r}
library(tidyverse)
library(lattice)

```

#Summarize Sperm Whale Encounters
From DataCleanupAllSurveys_2022.Rmd, use merge_tmp3
```{r}
merge_tmp3 %>%
  group_by(survey) %>%
  summarise(length(unique(ID))) 

```

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
From 10_pmSDM_AcousticCombined.Rmd

!!!!!!!! YB did these steps to get PmCombo used to make the models. ALL GOOD!
```{r eval=FALSE, echo=FALSE}

PmEdit <- CompleteTotal  #From DataCleanupAllSurveys_2022


##### COMBO: log transform, temporal variables, click code redo #####

# Nov 3: Redo the PmCombined dataset to 

PmEdit$log.effort	= log(PmEdit$EffArea)
PmEdit$log.chla <- log(PmEdit$chla)
PmEdit$log.eke <- log(PmEdit$eke)
PmEdit$log.wp <- log(PmEdit$wavepow)

#From Jeff, Oct 14, 2020
#add year and quarterly time variables
#year
PmEdit$year <- 2010
PmEdit$year[PmEdit$survey==1303] <- 2013 
PmEdit$year[PmEdit$survey==1604] <- 2016
PmEdit$year[PmEdit$survey==1705] <- 2017
PmEdit$year[PmEdit$survey==1706] <- 2017
PmEdit2 <- dplyr::select(PmEdit, last_col(), everything())

#quarter: test on PmEdit2, then add to PmEdit
# dateSplit = data.frame(1, time=as.character(as.Date(PmEdit$UTC)))  #some data finagling
# dateSplit2 = separate(dateSplit, "time",  c("year", "month", "day")) #splits date into year, month, day
# PmEdit2$month = as.numeric(dateSplit2$month) #add to PmEdit2 to test
# PmEdit2$qtr = 1
# PmEdit2$qtr[PmEdit2$month >= 4 & PmEdit2$month <=6] <- 2 
# PmEdit2$qtr[PmEdit2$month >= 7 & PmEdit2$month <=9] <- 3
# PmEdit2$qtr[PmEdit2$month >= 10 & PmEdit2$month <=12] <- 4
# PmEdit2 <- dplyr::select(PmEdit2, last_col(), everything())
# PmEdit$qtr <- PmEdit2$qtr

# saveRDS(PmEdit, here::here(  paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda') ))
```

```{r echo=FALSE, eval=FALSE}
### MUST DO!!!!! CLICK CODE REDO ####
# For DUPLICATES: need to validate the click code and make sure the cells with > 1 group have all clicks accounted for.
# did this somewhat manually by looking at each 'manyGrps' subset separately to combine click codes into one row to represent the cell. Didn't care about the individual data for each detection itself at this point. but that info is still available.

 # 1) Combined: find which rows have > 1 group in a cell 
manyGrps1641 =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1641), ]
manyGrps1303 =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1303), ]
manyGrps1604 =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1604), ]
manyGrps1705 =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1705), ]
manyGrps1706 =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1706), ]

#  1) Sightings Only:
manyGrps1641s =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1641 & PmEdit$sid !=999), ]
manyGrps1303s =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1303 & PmEdit$sid !=999), ]
manyGrps1604s =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1604 & PmEdit$sid !=999), ]
manyGrps1705s =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1705 & PmEdit$sid !=999), ]
manyGrps1706s =  PmEdit[ which(PmEdit$pa  > 1 & PmEdit$survey == 1706 & PmEdit$sid !=999), ]



 # 2) add in the missing click type to the click code VERY CAREFULLY: no need for 1303 in this step for either
 #    Data for 1641 and 1303 differ slightly between data sets, double-check rows and click codes saved!
# manyGrps1641[1,41] <- 'cds'
# manyGrps1641[3,54] <- 'ckcd'
#
manyGrps1303[2:3,41] <- 'ckrs'
manyGrps1303[5,41] <- 'rs'
#
manyGrps1705[2,41] <- 'cdr'

#acoustic**** 1705 *********
# manyGrps1705[2,41] <- 'cdrs'

#
manyGrps1706[1:2,41] <- 'cdr'

# ## Since not using click codes for sighting only model, use this to change pa so it doesn't incorporate any acoustic encounters
# manyGrps1641s$pa <- 1    # pa==2 because an acoustic enc was included, change to 1 for sighting
# #
# manyGrps1303s$pa[1:3] <- 1
# manyGrps1303s$pa[4:5] <- 2
# #
# manyGrps1604s$pa <- 1
# manyGrps1705s$pa <- 2
# manyGrps1706s$pa <- 1  #keep both 

 # 3) keep the rows with all encompassing click code
#for COMBO
manyGrps1641ed <- manyGrps1641[c(1,3), ]  #4/2022: no change, no duplicates including A999
manyGrps1303ed <- manyGrps1303[c(1:3,5), ]  #4/2022: includes A999, so take last row for clk code
manyGrps1705ed <- manyGrps1705[1, ]           #4/2022: includes A999, so take last row for clk code
manyGrps1706ed <- manyGrps1706[1, ]  #4/2022: includes A999, so take last row for clk code

# manyGrps1641eds <- manyGrps1641s                #Nov3: no change, no duplicates including A999
# manyGrps1303eds <- manyGrps1303s[c(1:3,5), ]    #Nov3: includes A999
# manyGrps1604eds <- manyGrps1604s
# manyGrps1705eds <- manyGrps1705s[2, ]           #Nov3: includes A999
# manyGrps1706eds <- manyGrps1706s                #Nov3: includes A999



#for Acoustics Only
manyGrps1641ed <- manyGrps1641[c(1,3,5,8), ]
manyGrps1303ed <- manyGrps1303[c(1,4,5,7), ]
manyGrps1705ed <- manyGrps1705[2, ]
manyGrps1706ed <- manyGrps1706[c(1,3,6), ]

#same for both
# manyGrps1604ed <- manyGrps1604[1, ]


# gotta remove these rows with >1 group from PmEdit...so keep all rows for absences and single groups
PmEditsubby <- subset( PmEdit, pa==0 | pa == 1 ) #Combined and Acoustic
PmEditsubS  <-subset( PmEdit, pa==0 | (pa == 1 & sid !=999)) #Sighting


#save for Combined Models
PmCombo <- rbind(PmEditsubby, manyGrps1641ed, manyGrps1303ed, manyGrps1705ed, manyGrps1706ed)
saveRDS(PmCombo, here::here(paste0('data/Pm_Combined_2022.rda')))

```

~~~~~~~~~~~~~~~~~~~~~~~~~~

# FIGURE 3 BARPLOT
## Use PmCombo to tally up the different encounters for each survey

```{r}
require(tidyr)

# simplify PmCombo and add column for Forage/NonForage

PmTally <- select(PmForage, pa, ID, year, acid, sid, click_code, peak, loc, forage)
PmTally <- plyr::ddply(PmTally, .(ID), randomRows, 1) #get unique IDs
PmTally <- PmTally[-194,]
PmTally <- dplyr::filter(PmTally, acid < 999)

PmTally$name <- NA

PmTally$name <- ifelse(PmTally$loc == 1, 'Localized' , 
                       ifelse(PmTally$sid < 999, 'Sighted',
                             ifelse(PmTally$loc == 0, 'Trackline', 'NONE'))
                )

PmTally$name <- ifelse(PmTally$loc == 0 & PmTally$sid == 999, 'Trackline', PmTally$name)
PmTally$forage <- ifelse(PmTally$forage == 1, 'Foraging', 'Non-Foraging')

Pm_uniq <- plyr::ddply(PmTally, .(ID), randomRows, 1) #unique IDs

PmTally2 <- select(Pm_uniq, year, name, forage)

# ggplot(PmTally2) +
  # geom_bar(aes(x=year, y=name, fill = forage))



#make long format??
PmTallyLong <- pivot_longer(Pm_uniq, cols= name)#c(forage, name))

fig3 <-aggregate(x = PmTallyLong$pa,                # Specify data column
          by = list(PmTallyLong$year,PmTallyLong$value, PmTallyLong$forage),              # Specify group indicator
          FUN = sum)
colnames(fig3) <- c('year', 'type', 'forage', 'total')

table5 <-aggregate(x = PmTallyLong$pa,                # Specify data column
          by = list(PmTallyLong$value),              # Specify group indicator
          FUN = sum)
colnames(fig3) <- c('year', 'type', 'forage', 'total')



require(viridis)

vir_pal <- viridis_pal(option = 'A')(20)


png(file = "C:/Users/yvers/OneDrive/Documents/Publications/2022_FiRS_SDMs/figures/Fig3_BarPlot.png", width = 175, height = 120, res = 300, units = "mm") # 8x6"

# pdf(file = "C:/Users/yvers/OneDrive/Documents/Publications/2022_FiRS_SDMs/figures/Fig3_BarPlot.pdf", width = 6.6, height = 4.7)

ggplot(fig3, aes(x=type, y=total, fill=forage)) + 
  geom_bar(stat="identity") + 
  facet_grid(~year) +
  theme_bw() +
   scale_fill_manual(values = c('#3b0f70','#f7705c'), name = "") + # '#3b0f70' use if 3 fills needed
  theme(axis.text.x = element_text(angle = 86, hjust = 1),
        strip.text.x = element_text(size = 18),
        axis.text = element_text(size=12, colour = 'black'),
        legend.text = element_text(size= 10),
        # legend.position = 'top',
        axis.title.x = element_text(size=14, margin = unit(c(4,0,0,0), 'mm')),
        axis.title.y = element_text(size=14, margin = unit(c(0,4,0,0), 'mm'))) +
  labs(x='Encounter Type', y = 'Total Groups')
  
dev.off()  

# ggplot(d3, aes(x=as.factor(year), y=total, fill=type)) + 
#   geom_bar(stat="identity", position = position_dodge()) +
#   theme_bw() + 
#   scale_fill_manual(values = vir_pal,
#                     #breaks = c('Localized Acoustic', 'Trackline Acoustic', 'Sighted Acoustic', 'Foraging', 'NonForaging'),
#                     name = "Encounter Type")
#   scale_x_discrete(limits=d3$year)
 



```


#Fig XX
New plot to show types of clicks in the different encounter types
```{r}
Pm_uniq$click_code2 <- str_replace_all(Pm_uniq$click_code, "r|ck", "f")
Pm_uniq$click_code2 <- str_replace_all(Pm_uniq$click_code2, "ff", "f")


PmTallyLong2 <- Pm_uniq %>% select(pa, ID, forage, name, click_code2) %>%
  pivot_longer(cols = name)

figX <-aggregate(x = PmTallyLong2$pa,                # Specify data column
          by = list(PmTallyLong2$forage,PmTallyLong2$value, PmTallyLong2$click_code2),              # Specify group indicator
          FUN = sum)
colnames(figX) <- c('forage', 'type', 'click_code2', 'total')

require(viridis)

vir_pal <- viridis_pal(option = 'A')(10)
vir_pal2 <- c(vir_pal[c(2:4, 6, 7)], "#FEC98DFF" ,"#FCFDBFFF", "#FEAF77FF")

png(file = "C:/Users/yvers/OneDrive/Documents/Publications/2022_FiRS_SDMs/figures/FigX_BarPlot_clickcode.png", width = 175, height = 120, res = 300, units = "mm") # 8x6"

# pdf(file = "C:/Users/yvers/OneDrive/Documents/Publications/2022_FiRS_SDMs/figures/Fig3_BarPlot.pdf", width = 6.6, height = 4.7)

ggplot(figX, aes(x=forage, y=total, fill=click_code2)) + 
  geom_bar(stat="identity") + 
  facet_grid(~type) +
  theme_bw() +
   scale_fill_manual(values = vir_pal2, name = "Click Types",
                     labels=c("codas", "codas/forage", "codas/forage/slow","codas/slow","forage", "forage/slow", "slow")) + # '#3b0f70' use if 3 fills needed
  theme(axis.text.x = element_text(angle = 86, hjust = 1),
        strip.text.x = element_text(size = 16),
        axis.text = element_text(size=12, colour = 'black'),
        legend.text = element_text(size= 10),
        # legend.position = 'top',
        axis.title.x = element_text(size=14, margin = unit(c(0,0,0,0), 'mm')),
        axis.title.y = element_text(size=14, margin = unit(c(0,4,0,0), 'mm'))) +
  labs(x='', y = 'Total Groups')
  
dev.off()  


```





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
surveynum = 1641
gridsize = 25
loctype = 'Combined'
loctype2 = 'Comb'

```

```{r}
# LOAD EDITED gridEffort and centroidALL - already checked for things in 1_GridData_v4

gridEffort <- readRDS(here::here(  paste0('output/gridEffort/', gridsize, ' km-', loctype, '/gridEffort', surveynum, '_', gridsize, 'km_', loctype2, '.rda') ))

centroidAll <- readRDS(here::here(paste0('output/centroids/', gridsize, ' km-', loctype), paste0('centroids', surveynum, '_', gridsize, 'km_', loctype2, '.rda')))

```




# The Complete Data Set
```{r, echo=F, eval=FALSE, message=FALSE}
library(tidyverse)
library(mgcv)
library(corrplot)
library(geoR)
library(tidymv)
library(here)
library(lubridate)
```

```{r echo=FALSE}
#Values for COMBINED MODELS
survey = 'AllSurveys'
gridsize = 25
loctype = 'Combined'
loctype2 = 'Comb'
#data has logged vars and temporal variables
## SAVED from line 251 in 10_PmSDM_AcousticCombined.rmd - Does not include forage column
PmCombo <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_EDIT.rda') )) #no A999's
PmCombo2 <- readRDS(here::here( paste0('output/models/', loctype, '/data/', 'CompletePm_', gridsize, 'km_', loctype2, '_scaled.rda') ))

PmCombined <- readRDS( here::here('data/Pm_Combined.rda')) #contains A999's


#From 16km data saved in 9_DataCleanup_ALlSurveys_2022
PmCombo <- CompleteTotal
```


# Sperm whale data used for centroids
```{r}
sw <- read.csv(here::here('output/SpermiesFinalLR_20200919.csv'))

length(unique(sw$ID))


```



#From 10_PmSDM_Foraging

### 2022: Redo Click Codes
```{r}
PmCombo$click_code[PmCombo$ID == "1303_42_21"] <- "cdckrs"  #(was ckrs, added coda)
PmCombo$click_code[PmCombo$ID == "1604_30_999"] <- "rs"     #(was s)
PmCombo$click_code[PmCombo$ID == "1604_67_999"] <- "rs"     #(was s)
PmCombo$click_code[PmCombo$ID == "1705_251_999"] <- "rs"    #(was s)
PmCombo$click_code[PmCombo$ID == "1706_2_999"] <- "rs"      #(was r)
PmCombo$click_code[PmCombo$ID == "1706_41_999"] <- "rs"     #(was s)
# PmCombo$click_code[PmCombo$ID == "1706_135_999"] <- "ckrs" #not included as separate encounter, was combined with 1706A49 in same cell, which is also ckrs
PmCombo$click_code[PmCombo$ID == "1303_71_999"] <- "rs"     #(was s)


```

### Run some totals
Sightings, Sightings/Acoustics, Acoustics Only
```{r}
#sighted only
PmVisOnly <- filter(PmCombined, acid == 999)
aggregate(x = PmVisOnly$pa,                # Specify data column
          by = list(PmVisOnly$survey),              # Specify group indicator
          FUN = sum)

PmVisOnly2 <- readRDS( here::here('data/Pm_Sighting.rda'))

PmVisOnly2_A999 <- filter(PmVisOnly2, acid == 999)
aggregate(x = PmVisOnly2_A999$pa,                # Specify data column
          by = list(PmVisOnly2_A999$survey),              # Specify group indicator
          FUN = sum)

PmVisEd <- readRDS(here::here(  paste0('output/models/Combined/data/SighingOnly_25km_EditDup.rda') ))
PmVisEd_A999 <- filter(PmVisEd, acid == 999)

aggregate(x = PmVisEd_A999$pa,                # Specify data column
          by = list(PmVisEd_A999$survey),              # Specify group indicator
          FUN = sum)

PmVis <- readRDS(here::here(  paste0('output/models/',loctype, '/data/SighingOnly_25km_Combined.rda') ))
PmVisOnly3 <- filter(PmVis, acid == 999)




#Sighted and acoustic acid <999, sid <999
PmVisAc <- filter(PmCombo, acid <999 & sid< 999)
aggregate(x = PmVisAc$pa,                # Specify data column
          by = list(PmVisAc$survey),              # Specify group indicator
          FUN = sum)

#sighted and acoustic acid <999, sid <999 - 2 less 1706 than table
PmAcLoc <- filter(PmCombo, sid == 999 & loc == 1)
aggregate(x = PmAcLoc$pa,                # Specify data column
          by = list(PmAcLoc$survey),              # Specify group indicator
          FUN = sum)

#Acoustics - Localized / Trackline encounters
PmAcLoc2 <- readRDS( here::here('data/Pm_Acoustics.rda'))

#has 2 more 1706 than PmAcLoc (1706_327_999, 1706_325_999) - ALL GOOD with table
PmAcLoc_2 <- filter(PmAcLoc2, sid == 999 & loc == 1)
aggregate(x = PmAcLoc_2$pa,                
          by = list(PmAcLoc_2$survey),    
          FUN = sum)

#has 2 less 1706 than the table & 2 more 1641! WTF
PmAcUnloc <- filter(PmCombo, sid == 999 & loc == 0)
aggregate(x = PmAcUnloc$pa,                
          by = list(PmAcUnloc$survey),    
          FUN = sum)
#has 2 less 1706 than the table!
PmAcUnloc_2 <- filter(PmAcLoc2, sid == 999 & loc == 0)
aggregate(x = PmAcUnloc_2$pa,                
          by = list(PmAcUnloc$survey),    
          FUN = sum)


# Check # of encounters by survey
Pm1303 <- filter(PmCombo, survey == 1303); sum(Pm1303$pa)
Pm2017 <- filter(PmCombo, survey == 1706 | survey ==1705)
sum(Pm2017$pa)

Pm1706b <- filter(CompleteScaled, survey == 1706, acid < 999) #CompleteScaled from 9_DataCleanup

Pm2017_pre <- filter(CompleteScaled, survey == 1705 | survey == 1706)
sum(Pm2017_pre$pa)

```


### 2022: Sanity Checks

## Foraging and Non-Foraging Group Models

#### Foraging models incorporated a subset of the sperm whale encounters from the Combined Models (sighted with acoustics, and acoustics only) that only included encounters with **regular clicks** and **creaks** to indicate foraging.


```{r eval=FALSE,echo=FALSE}
## Forage and Social Group Data Subset ####
#1. Denote forage or not. 
PmFor1 <- filter(PmCombo, grepl('r|k', click_code))   # the 'k' implies ck for creaks because c also for codas
PmFor0 <- filter(PmCombo, !grepl('r|k', click_code))
PmFor1$forage <- 1  #forage
PmFor0$forage <- 0  #non-forage
PmForage <- bind_rows(PmFor1, PmFor0)


nrow(PmFor1)
nrow(PmFor0)

#2. Denote non-foraging
PmPositive <- filter( PmCombo, pa > 0 )
PmAbsent <- filter( PmCombo, pa == 0 )

PmNonFor1 <- filter(PmPositive, !grepl('r|k', click_code))  #subset all encs without r or ck in click_code
PmNonFor0 <- filter(PmPositive, grepl('r|k', click_code))  #this is the same as PmFor1

#designate whether the encounters included codas (1) or not (0)
PmNonFor1$nonfor <- 1
PmNonFor0$nonfor <- 0
PmAbsent$nonfor <- 0

PmNonFor <- rbind(PmNonFor1,PmNonFor0, PmAbsent)


# PmNonFor2 <- PmNonFor[, c(1,2,4,54,62)]  #don't know why this is here

nrow(PmNonFor1)  # represent all non-foraging 
nrow(PmNonFor0)  # represent all foraging encounters, which are now 0
nrow(PmAbsent)
```


```{r}
# How many encounters are foraging and non-foraging? 196?! Not 209...
sum(PmFor1$pa) # 109 foraging, not include pa >1, 122
sum(PmNonFor1$pa) #71 non-foraging, not include pa >1, 74

# How many cells have multiple groups, pa > 1? How many total groups? 
PmFor_multi <- PmFor1[which(PmFor1$pa>1),]; length(PmFor_multi$pa);sum(PmFor_multi$pa)
PmNonFor_multi <- PmNonFor1[which(PmNonFor1$pa>1),]; length(PmNonFor_multi$pa); sum(PmNonFor_multi$pa)


# Are encounters used >1 between foraging and non-foraging models? NO
PmFor1$ID %in% PmNonFor1 
PmNonFor1$ID %in% PmFor1
```

```{r}


```


